### 2021-11-11 href.tex
Update comments

\newcommand{\discreteCosineRF}{ \href{https://docs.scipy.org/doc/scipy/tutorial/fft.html\#discrete-cosine-transforms}{discrete\_cosine\_transform} }
\newcommand{\fftOneDRf}{
\href{https://docs.scipy.org/doc/scipy/tutorial/fft.html\#d-discrete-fourier-transforms}{fft\_1D}
}

\newcommand{\fftTwoDRf}{
\href{https://docs.scipy.org/doc/scipy/tutorial/fft.html\#and-n-d-discrete-fourier-transforms}{fft\_2D}
}

\newcommand{\interOneDRf}{
\href{https://docs.scipy.org/doc/scipy/tutorial/interpolate.html\#d-interpolation-interp1d}{interpolation\_1D} 
}

\newcommand{\multiRf}{
\href{https://docs.scipy.org/doc/scipy/tutorial/interpolate.html\#multivariate-data-interpolation-griddata}{multivariate\_data} 
}

\newcommand{\bsplineRf}{
\href{https://docs.scipy.org/doc/scipy/tutorial/signal.html\#b-splines}{bspline}  
}

\newcommand{\splineOneDRf}{
    \href{https://docs.scipy.org/doc/scipy/tutorial/interpolate.html\#spline-interpolation-in-1-d-procedural-interpolate-splxxx}{
    spline\_1D} 
}

\newcommand{\splineTwoDRf}{
    \href{https://docs.scipy.org/doc/scipy/tutorial/interpolate.html\#d-spline-representation-procedural-bisplrep}{
    spline\_2D} 
}

\newcommand{\bfgsRf}{
    \href{https://docs.scipy.org/doc/scipy/tutorial/optimize.html\#broyden-fletcher-goldfarb-shanno-algorithm-method-bfgs}{broyden\_fletcher\_goldfarb\_shanno} 
}

\newcommand{\globalRf}{
    \href{https://docs.scipy.org/doc/scipy/tutorial/optimize.html\#global-optimization}{
    global\_optimization} 
}

\newcommand{\slsqpRf}{
    \href{https://docs.scipy.org/doc/scipy/tutorial/optimize.html\#sequential-least-squares-programming-slsqp-algorithm-method-slsqp}{
    slsqp} 
}

\newcommand{\lsmRf}{
    \href{https://docs.scipy.org/doc/scipy/tutorial/optimize.html\#least-squares-minimization-least-squares}{
    least\_square\_minimization} 
}

\newcommand{\nelderRf}{
    \href{https://docs.scipy.org/doc/scipy/tutorial/optimize.html\#nelder-mead-simplex-algorithm-method-nelder-mead}{nelder\_mead\_simplex}
}

\newcommand{\ncgRf}{
    \href{https://docs.scipy.org/doc/scipy/tutorial/optimize.html\#newton-conjugate-gradient-algorithm-method-newton-cg}{
    newton\_conjugate\_gradient}  
}

\newcommand{\rootRf}{
    \href{https://docs.scipy.org/doc/scipy/tutorial/optimize.html\#sets-of-equations}{
    root\_findings}  
}

\newcommand{\rootLargRf}{
    \href{https://docs.scipy.org/doc/scipy/tutorial/optimize.html\#root-finding-for-large-problems}{
    root\_finding\_large} 
}

\newcommand{\rootLargePredRf}{
    \href{https://docs.scipy.org/doc/scipy/tutorial/optimize.html\#still-too-slow-preconditioning}{
    root\_finding\_large\_preconditioned} 
}

\newcommand{\trustExactRf}{
    \href{https://docs.scipy.org/doc/scipy/tutorial/optimize.html\#trust-region-nearly-exact-algorithm-method-trust-exact}{
    trust\_region\_exact} 
}

\newcommand{\trustTruncRf}{
    \href{https://docs.scipy.org/doc/scipy/tutorial/optimize.html\#trust-region-truncated-generalized-lanczos-conjugate-gradient-algorithm-method-trust-krylov}{
    trust\_region\_truncated\_generalized\_lanczos} 
}

\newcommand{\trustNCGRf}{
    \href{https://docs.scipy.org/doc/scipy/tutorial/optimize.html\#trust-region-newton-conjugate-gradient-algorithm-method-trust-ncg}{
    trust\_region\_newton\_conjugate\_gradient} 
}

\newcommand{\AdaboostRf}{
    \href{https://scikit-learn.org/stable/auto_examples/ensemble/plot_adaboost_regression.html}{AdaBoost} 
}

\newcommand{\brrRf}{
    \href{https://scikit-learn.org/stable/auto_examples/linear_model/plot_bayesian_ridge.html}{Bayesian Ridge Regression} 
}

\newcommand{\onlineClassifierComparisonRf}{
    \href{https://scikit-learn.org/stable/auto_examples/linear_model/plot_sgd_comparison.html}{Online classifier comparison} 
}

\newcommand{\kmeansRf}{
    \href{https://scikit-learn.org/stable/auto_examples/cluster/plot_cluster_iris.html}{K-Means Clustering}  
}

\newcommand{\covarianceRf}{
    \href{https://scikit-learn.org/stable/auto_examples/covariance/plot_mahalanobis_distances.html}{Covariance estimation}  
}

\newcommand{\decisionRf}{
    \href{https://scikit-learn.org/stable/auto_examples/tree/plot_tree_regression.html}{Decision Tree Regression}
}

\newcommand{\digitsRf}{
    \href{https://scikit-learn.org/stable/auto_examples/linear_model/plot_sgd_comparison.html}{Digits Classification} 
}

\newcommand{\faceRf}{
    \href{https://scikit-learn.org/stable/auto_examples/decomposition/plot_faces_decomposition.html}{Face Recognition} 
}

\newcommand{\penaltyRf}{
    \href{https://scikit-learn.org/stable/auto_examples/linear_model/plot_logistic_l1_l2_sparsity.html}{L1 Penalty} 
}

\newcommand{\lassoRf}{
    \href{https://scikit-learn.org/stable/auto_examples/linear_model/plot_lasso_and_elasticnet.html}{Lasso and Elastic Net} 
}

\newcommand{\hyperplaneRf}{
    \href{https://scikit-learn.org/stable/auto_examples/linear_model/plot_sgd_separating_hyperplane.html}{Separating hyperplane} 
}

\newcommand{\mnistRf}{
    \href{https://scikit-learn.org/stable/auto_examples/linear_model/plot_sparse_logistic_regression_mnist.html}{MNIST classification} 
}

\newcommand{\multitaskRf}{
    \href{https://scikit-learn.org/stable/auto_examples/linear_model/plot_multi_task_lasso_support.html}{Multitask Lasso}  
}

\newcommand{\ompRf}{
    \href{https://scikit-learn.org/stable/auto_examples/linear_model/plot_omp.html}{Orthogonal Matching} 
}

\newcommand{\pcaRf}{
    \href{https://scikit-learn.org/stable/auto_examples/decomposition/plot_pca_iris.html}{PCA decomposition}  
}

\newcommand{\robustRf}{
    \href{https://scikit-learn.org/stable/auto_examples/linear_model/plot_ransac.html}{Robust Linear Regression} 
}

\newcommand{\toyRf}{
    \href{https://scikit-learn.org/stable/auto_examples/cluster/plot_segmentation_toy.html\#sphx-glr-auto-examples-cluster-plot-segmentation-toy-py}{Segmentation toy}  
}

\newcommand{\svmRf}{
    \href{https://scikit-learn.org/stable/auto_examples/svm/plot_separating_hyperplane_unbalanced.html\#sphx-glr-auto-examples-svm-plot-separating-hyperplane-unbalanced-py}{SVM}  
}

\newcommand{\tomographyRf}{
    \href{https://scikit-learn.org/stable/auto_examples/applications/plot_tomography_l1_reconstruction.html\#sphx-glr-auto-examples-applications-plot-tomography-l1-reconstruction-py}{
    Tomography} 
}

\newcommand{\weightedRf}{
    \href{https://scikit-learn.org/stable/auto_examples/linear_model/plot_sgd_weighted_samples.html\#sphx-glr-auto-examples-linear-model-plot-sgd-weighted-samples-py}{SGD weighted samples} 
}