### 2022-04-14 IEEEtran/paper-annotated.tex
Revision

%% bare_jrnl_compsoc.tex
%% V1.4b
%% 2015/08/26
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
%% Computer Society journal paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/pkg/ieeetran
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall the IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. The IEEE's font choices and paper sizes can   ***
% *** trigger bugs that do not appear when using other class files.       ***                          ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

\documentclass[10pt,journal,compsoc]{IEEEtran}

\usepackage{todonotes}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{color}
\usepackage{hyperref}
\usepackage{multirow}
\usepackage{pifont}
\usepackage{xcolor,colortbl}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=blue!50!black,citecolor=blue!75!black,filecolor=black,urlcolor=blue!75!black}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{xspace}
\usepackage{makecell}
\usepackage{ulem} % provides sout
\usepackage[frozencache=true,cachedir=minted-cache]{minted} 
\usepackage[htt]{hyphenat}      
\usepackage{subcaption}
\usepackage{graphicx,import}
\usepackage{booktabs}
\usepackage{soul}

%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[10pt,journal,compsoc]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/pkg/ifpdf
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
\ifCLASSOPTIONcompsoc
  % IEEE Computer Society needs nocompress option
  % requires cite.sty v4.0 or later (November 2003)
  \usepackage[nocompress]{cite}
\else
  % normal IEEE
  \usepackage{cite}
\fi
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.
%
% Note that some packages require special options to format as the Computer
% Society requires. In particular, Computer Society  papers do not use
% compressed citation ranges as is done in typical IEEE papers
% (e.g., [1]-[4]). Instead, they list every citation separately in order
% (e.g., [1], [2], [3], [4]). To get the latter we need to load the cite
% package with the nocompress option which is supported by cite.sty v4.0
% and later. Note also the use of a CLASSOPTION conditional provided by
% IEEEtran.cls V1.7 and later.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at: 
% http://www.ctan.org/pkg/graphicx
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
%\usepackage{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics.
%
% Note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/amsmath

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/pkg/algorithms
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/pkg/algorithmicx

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/array

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=footnotesize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a sans serif font rather
% than the serif font used in traditional IEEE formatting and thus the need
% to invoke different subfig.sty package options depending on whether
% compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/pkg/subfig

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure.
% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
% corrections already built into the system in which case a warning will
% be issued if an attempt is made to load fixltx2e.sty as it is no longer
% needed.
% The latest version and documentation can be found at:
% http://www.ctan.org/pkg/fixltx2e

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/stfloats
% Do not use the stfloats baselinefloat ability as the IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that the IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/pkg/dblfloatfix

%\ifCLASSOPTIONcaptionsoff
%  \usepackage[nomarkers]{endfloat}
% \let\MYoriglatexcaption\caption
% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
%\fi
% endfloat.sty was written by James Darrell McCauley, Jeff Goldberg and 
% Axel Sommerfeldt. This package may be useful when used in conjunction with 
% IEEEtran.cls'  captionsoff option. Some IEEE journals/societies require that
% submissions have lists of figures/tables at the end of the paper and that
% figures/tables without any captions are placed on a page by themselves at
% the end of the document. If needed, the draftcls IEEEtran class option or
% \CLASSINPUTbaselinestretch interface can be used to increase the line
% spacing as well. Be sure and use the nomarkers option of endfloat to
% prevent endfloat from "marking" where the figures would have been placed
% in the text. The two hack lines of code above are a slight modification of
% that suggested by in the endfloat docs (section 8.4.1) to ensure that
% the full captions always appear in the list of figures/tables - even if
% the user used the short optional argument of \caption[]{}.
% IEEE papers do not typically make use of \caption[]'s optional argument,
% so this should not be an issue. A similar trick can be used to disable
% captions of packages such as subfig.sty that lack options to turn off
% the subcaptions:
% For subfig.sty:
% \let\MYorigsubfloat\subfloat
% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
% However, the above trick will not work if both optional arguments of
% the \subfloat command are used. Furthermore, there needs to be a
% description of each subfigure *somewhere* and endfloat does not add
% subfigure captions to its list of figures. Thus, the best approach is to
% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
% and instead reference/explain all the subfigures within the main caption.
% The latest version of endfloat.sty and its documentation can obtained at:
% http://www.ctan.org/pkg/endfloat
%
% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
% later in the document, say, to conditionally put the References on a 
% page by themselves.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\newcommand{\cross}[0]{\cellcolor{red!65}\ding{53}}
\newcommand{\valid}[0]{\cellcolor{green!75!black}\ding{51}}
\newcommand{\warn}[0]{\cellcolor{orange!75}?}
\newcommand{\na}[0]{\cellcolor{gray!25}}
\newcommand{\s}[1]{\cellcolor{cyan!25}#1}
\newcommand{\scross}[0]{\ding{53}~}
\newcommand{\svalid}[0]{\ding{51}~}
\newcommand{\swarn}[0]{?~}
\newcommand{\pytracer}[0]{PyTracer\xspace}

\definecolor{lightred}{rgb}{1.0,0.6,0.6}
\definecolor{lightgreen}{rgb}{0.6,1.0,0.6}

\DeclareRobustCommand{\add}[1]{{\sethlcolor{lightgreen}\hl{#1}}}
\DeclareRobustCommand{\remove}[1]{{\sethlcolor{lightred}\hl{#1}}}

\DeclareRobustCommand{\addCite}[1]{{\sethlcolor{lightgreen}\hl{\mbox{#1}}}}
\DeclareRobustCommand{\removeCite}[1]{{\sethlcolor{lightred}\hl{\mbox{#1}}}}

\setlength{\arrayrulewidth}{0.1pt}

\include{href}

\makeatletter
\newcommand{\thickhline}{%
    \noalign {\ifnum 0=`}\fi \hrule height 1pt
    \futurelet \reserved@a \@xhline
}
\newcolumntype{"}{@{\hskip\tabcolsep\vrule width 1pt\hskip\tabcolsep}}
\makeatother

\lstdefinestyle{customPython}{
  belowcaptionskip=1\baselineskip,
  breaklines=true,
  xleftmargin=\parindent,
  language=Python,
  showstringspaces=false,
  basicstyle=\scriptsize\ttfamily,
  keywordstyle=\bfseries\color[rgb]{0.580, 0.000, 0.827},
  %{purple!40!lightgray},
  commentstyle=\itshape\color{green!40!black},
  identifierstyle=\bfseries\color{cyan!75!black},
  stringstyle=\color{orange},
  deletekeywords={double,float},
  classoffset=1, % starting new class
  otherkeywords={double,float},
  morekeywords={double,float},
  keywordstyle=\bfseries\color{green!55!black},
  classoffset=0
}

\lstdefinestyle{customC}{
  belowcaptionskip=1\baselineskip,
  breaklines=true,
  xleftmargin=\parindent,
  language=C,
  showstringspaces=false,
  basicstyle=\scriptsize\ttfamily,
  keywordstyle=\bfseries\color[rgb]{0.580, 0.000, 0.827},
  %{purple!40!lightgray},
  commentstyle=\itshape\color{green!40!black},
  identifierstyle=\bfseries\color{cyan!75!black},
  stringstyle=\color{orange},
  deletekeywords={double,float},
  classoffset=1, % starting new class
  otherkeywords={double,float},
  morekeywords={double,float},
  keywordstyle=\bfseries\color{green!55!black},
  classoffset=0
}

\begin{document}

\makeatletter
\let\orig@lstnumber=\thelstnumber
\newcommand\lstsetnumber[1]{\gdef\thelstnumber{#1}}
\newcommand\lstresetnumber{\global\let\thelstnumber=\orig@lstnumber}
\makeatother

%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{PyTracer: Automatically Profiling Numerical Instabilities in Python}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%
%
%\IEEEcompsocitemizethanks is a special \thanks that produces the bulleted
% lists the Computer Society journals use for "first footnote" author
% affiliations. Use \IEEEcompsocthanksitem which works much like \item
% for each affiliation group. When not in compsoc mode,
% \IEEEcompsocitemizethanks becomes like \thanks and
% \IEEEcompsocthanksitem becomes a line break with idention. This
% facilitates dual compilation, although admittedly the differences in the
% desired content of \author between the different types of papers makes a
% one-size-fits-all approach a daunting prospect. For instance, compsoc 
% journal papers have the author affiliations above the "Manuscript
% received ..."  text while in non-compsoc journals this is reversed. Sigh.

\author{Yohan Chatelain$^1$, Nigel Yong Sao Young$^1$,  Gregory Kiar$^2$,  Tristan Glatard$^1$\\
    $^1$Department of Computer Science and Software Engineering, Concordia University, Montreal, Canada\\
    $^2$Center for the Developing Brain, Child Mind Institute, New York, NY, USA}
%% \author{Michael~Shell,~\IEEEmembership{Member,~IEEE,}
%%         John~Doe,~\IEEEmembership{Fellow,~OSA,}
%%         and~Jane~Doe,~\IEEEmembership{Life~Fellow,~IEEE}% <-this % stops a space
%% \IEEEcompsocitemizethanks{\IEEEcompsocthanksitem M. Shell was with the Department
%% of Electrical and Computer Engineering, Georgia Institute of Technology, Atlanta,
%% GA, 30332.\protect\\
%% % note need leading \protect in front of \\ to get a newline within \thanks as
%% % \\ is fragile and will error, could use \hfil\break instead.
%% E-mail: see http://www.michaelshell.org/contact.html
%% \IEEEcompsocthanksitem J. Doe and J. Doe are with Anonymous University.}% <-this % stops an unwanted space
%% \thanks{Manuscript received April 19, 2005; revised August 26, 2015.}}

% note the % following the last \IEEEmembership and also \thanks - 
% these prevent an unwanted space from occurring between the last author name
% and the end of the author line. i.e., if you had this:
% 
% \author{....lastname \thanks{...} \thanks{...} }
%                     ^------------^------------^----Do not want these spaces!
%
% a space would be appended to the last name and could cause every name on that
% line to be shifted left slightly. This is one of those "LaTeX things". For
% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
% "AB" then you have to do: "\textbf{A}\textbf{B}"
% \thanks is no different in this regard, so shield the last } of each \thanks
% that ends a line with a % and do not let a space in before the next \thanks.
% Spaces after \IEEEmembership other than the last one are OK (and needed) as
% you are supposed to have spaces between the names. For what it is worth,
% this is a minor point as most people would not even notice if the said evil
% space somehow managed to creep in.

% The paper headers
\markboth{Journal of \LaTeX\ Class Files,~Vol.~14, No.~8, August~2015}%
{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Computer Society Journals}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

% The publisher's ID mark at the bottom of the page is less important with
% Computer Society journal papers as those publications place the marks
% outside of the main text columns and, therefore, unlike regular IEEE
% journals, the available text space is not reduced by their presence.
% If you want to put a publisher's ID mark on the page you can do it like
% this:
%\IEEEpubid{0000--0000/00\$00.00~\copyright~2015 IEEE}
% or like this to get the Computer Society new two part style.
%\IEEEpubid{\makebox[\columnwidth]{\hfill 0000--0000/00/\$00.00~\copyright~2015 IEEE}%
%\hspace{\columnsep}\makebox[\columnwidth]{Published by the IEEE Computer Society\hfill}}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark (Computer Society jorunal
% papers don't need this extra clearance.)

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\IEEEtitleabstractindextext{%
    \begin{abstract}
        Numerical stability is a crucial requirement of reliable scientific
        computing. However, despite the pervasiveness of Python in data science,
        analyzing large Python programs remains challenging due to the lack of
        scalable numerical analysis tools available for this language. To fill
        this gap, we developed \pytracer, a profiler to quantify numerical
        instability in Python applications. \pytracer transparently instruments
        Python code to produce numerical traces and visualize them interactively
        in a Plotly dashboard. We designed \pytracer to be agnostic to numerical
        noise model , allowing for tool evaluation through Monte-Carlo
        Arithmetic, random rounding, random data perturbation, or structured
        noise for a particular application. We illustrate \pytracer's
        capabilities by testing the numerical stability of key functions in both
        SciPy and Scikit-learn, two dominant Python libraries for mathematical
        modeling. Through these evaluations, we demonstrate \pytracer as a
        scalable, automatic, and generic framework for numerical profiling in
        Python.
    \end{abstract}
}
% Note that keywords are not normally used for peerreview papers.
%% \begin{IEEEkeywords}
%% Computer Society, IEEE, IEEEtran, journal, \LaTeX, paper, template.
%% \end{IEEEkeywords}}

% make the title area
\maketitle

% To allow for easy dual compilation without having to reenter the
% abstract/keywords data, the \IEEEtitleabstractindextext text will
% not be used in maketitle, but will appear (i.e., to be "transported")
% here as \IEEEdisplaynontitleabstractindextext when the compsoc 
% or transmag modes are not selected <OR> if conference mode is selected 
% - because all conference papers position the abstract like regular
% papers do.
\IEEEdisplaynontitleabstractindextext
% \IEEEdisplaynontitleabstractindextext has no effect when using
% compsoc or transmag under a non-conference mode.

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\IEEEraisesectionheading{\section{Introduction}\label{sec:introduction}}
% Computer Society journal (but not conference!) papers do something unusual
% with the very first section heading (almost always called "Introduction").
% They place it ABOVE the main text! IEEEtran.cls does not automatically do
% this for you, but you can achieve this effect with the provided
% \IEEEraisesectionheading{} command. Note the need to keep any \label that
% is to refer to the section immediately after \section in the above as
% \IEEEraisesectionheading puts \section within a raised box.

\IEEEPARstart{T}{he} scientific Python ecosystem has become a central component
of data analyses in recent years, owing to its rich offering of data
manipulation, array programming, numerical analysis, and visualization tools. In
particular, libraries such as NumPy~\cite{harris2020array},
SciPy~\cite{virtanen2020scipy}, scikit-learn~\cite{pedregosa2011scikit} or
PyTorch~\cite{paszke2019pytorch} are used in hundreds of publications each year,
and serve as a reference set of high-quality open-source core scientific
libraries. Researchers have built numerous domain-specific software tools from
this ecosystem, leveraging Python's simplicity and flexibility.

Numerical stability is a crucial requirement of reliable scientific data
processing. In unstable analyses, small numerical perturbations introduced by
data noise, software and hardware updates, or parallelization have potential to
introduce substantial deviations in final results, leading to potentially
different scientific conclusions, ultimately threatening the reliability of
computational analyses. To prevent this, numerical stability analyses need to be
conducted systematically, however, no practical tool currently exists to conduct
such analyses for Python programs.

We present PyTracer, a numerical stability profiling and visualization tool for
Python programs. \pytracer adopts a dynamic approach that evaluates numerical
stability using program execution traces, and is applicable to large programs
without manual intervention. The transparent nature of \pytracer is particularly
valuable in contrast to theoretical approaches based on condition numbers or
backward error analysis~\addCite{\cite{higham2002accuracy}}, which require detailed modeling of the program and its
numerical implementation. Similarly, static code analysis techniques~\addCite{\cite{goubault2006static}} hardly
scale to large code bases, particularly in dynamically-typed languages such as
Python.

\pytracer estimates the variability of floating-point variables by combining
program traces obtained with different numerical perturbations, which requires
(\add{a}\remove{1}) generating numerically-perturbed executions, (\add{b}\remove{2}) tracing floating-point
computations along program executions, and (\add{c}\remove{3}) visualizing stability
evaluations. 
\add{PyTracer uses the ``fuzzy" Python environment instrumented with Monte-Carlo
arithmetic to address the requirement (a). \mbox{\pytracer} addresses the
remaining requirements with (b) a dynamic instrumentation of Python functions,
modules, and classes through meta-programming, (c) an interactive Plotly
dashboard to visualize stability measures throughout program executions.
\mbox{\pytracer} is available at
\mbox{\href{https://github.com/big-data-lab-team/pytracer.git}{github.com/big-data-lab-team/pytracer.git}}
under MIT License.}
\remove{\mbox{\pytracer} addresses these requirements with (1) a ``fuzzy" Python
    interpreter instrumented with Monte-Carlo arithmetic, (2) a dynamic
    instrumentation of Python functions, modules, and classes through
    meta-programming, (3) an interactive Plotly dashboard to visualize stability
    measures throughout program executions. }

This manuscript presents the design of \pytracer and demonstrates its potential
in various use cases. Through stability evaluations of the SciPy and
scikit-learn libraries, we demonstrate \pytracer as a practical solution to
review the numerical stability of extensive code bases.

\section{\pytracer design}

% \subsection{Pytracer workflow}
\pytracer adopts the three-step workflow shown in Figure~\ref{fig:workflow}.
First, \pytracer uses \textit{monkey patching} to replace the functions called
by the application with a wrapper that saves their arguments and returned values
in a trace file.
Second, it executes the application multiple times with noise injection, and
aggregates the resulting trace files to compute summary statistics about each
traced variable. Finally, a Plotly dashboard provides interactive visualizations
highlighting unstable code sections.

% \tristan{mention monkey patching}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figure/workflow.pdf}
    \caption{Pytracer workflow
        % \gkmod{}{Note: replace "Execution" with "Multiple Independent Executions", and maybe move it to be between the gears and the pkl files.}
    }
    \label{fig:workflow}
\end{figure}

\subsection{Python code instrumentation}

%\tristan{``Instrumentation" should appear in Figure 1}

\pytracer dynamically instruments Python modules without requiring source code
modifications in the application. To do so, it creates a new, instrumented
version of each module that preserves its original attributes so that the
application can transparently use it.
% \gkmod{}{Note: this form of monkey patching is basically global and
% transparent decorating, right?} \Yohan{Yes}
By default, \pytracer instruments all the attributes in a given module except
the special ones of the form \texttt{\_\_*\_\_}, which is common for reserved
Python methods. However, some attributes are not writable. In this case,
\pytracer \remove{preservers} \add{preserves} the original attributes and warns the user. The user can
also restrict the list of traced attributes through an inclusion/exclusion
mechanism. The main advantage of this instrumentation approach is the lack of
required modification in the application source code, in contrast with
\remove{decorator-based approaches} \add{a decorator-based approache~\mbox{\cite{hunt2019decorators}}}. 
In addition, this technique is scalable and
applicable in read-only environments such as Singularity containers or HPC
clusters.

\subsubsection{Intercepting module import}

Python loads modules through an import mechanism triggered by the
\texttt{import} keyword that involves two objects: \texttt{finders} and
\texttt{loaders}. \texttt{Finders} search for packages (or
namespaces\footnote{Python distinguishes between \texttt{regular} and
    \texttt{namespace} packages. A regular package is a directory that contains a
    \_\_init\_\_.py file and potentially subdirectories (sub-packages) that contain
    themselves a \texttt{\_\_init\_\_.py} file and so on recursively. The package
    hierarchy follows those of the directory. Namespace packages introduced in
    Python 3.3 (\href{https://www.python.org/dev/peps/pep-0420/}{PEP 420}) do not
    contain an \_\_init\_\_.py file and allow for flexible directory structures.
    Hence, parts of the package can be located in zip files, on the network, or in
    separated directories. There are no functional differences between both types of
    packages.}) on the local storage --- such as standard packages installed through
the pip package manager --- or on a distant server. \texttt{Finders} do not load
modules. Instead, they return a specification (\texttt{spec} object) that
includes information on where to find and how to load modules. \texttt{Loaders}
create module objects, initialize them, and fill import-related attributes such
as \texttt{\_\_spec\_\_}. \texttt{Loaders} then execute modules and populate
their namespace. Finally, modules are bound to their import name in the
\texttt{sys.modules} dictionary.

Python supports custom \texttt{finder} classes registered in the
\texttt{sys.meta\_path} list. When Python encounters an \texttt{import}
statement, it first looks for an existing binding in \texttt{sys.modules} and
then iterates over the \texttt{finder} classes in the \texttt{sys.meta\_path}
list until it finds the module. \pytracer adds a custom \texttt{finder} class at
the head of \texttt{sys.meta\_path} that intercepts the module import and
creates the module with a custom \texttt{loader} class.

\pytracer's \texttt{loader} class first loads the original module, then copies
it as a new instance of the \texttt{ModuleType} class. It then calls the
appropriate instrumentation function for each attribute depending on its type
(function, class, or instances) and replaces the original module in the
\texttt{sys.modules} map with the instrumented version. Finally, it updates all
existing references to the original module contained in the global symbol table.
As a result, the application will transparently call the instrumented module.

\subsubsection{Instrumenting module functions}

\remove{Listings~\mbox{\ref{fig:wrapper_creation}} and~\mbox{\ref{fig:generic_wrapper}} show
    \mbox{\pytracer}'s function instrumentation mechanism.}
For each function to instrument, \pytracer's \texttt{loader} class creates a
wrapper function, dynamically compiles it with the \texttt{compile} builtin
function, and substitutes it for the original module function.  Although simple,
this technique does not apply to callable class instances, i.e., class instances
that have a \texttt{\_\_call\_\_} attribute and are not of type
\texttt{FunctionType}. Indeed, applying the wrapper technique would 1) modify
the type of the class instance to \texttt{FunctionType}, which could cause
syntax and semantic errors, and 2) mask class attributes, leading to
\texttt{AttributeError} exceptions. \pytracer overrides the
\texttt{\_\_call\_\_} attribute with the wrapper function when possible to
overcome this issue. When the \texttt{\_\_call\_\_} attribute is not writable,
Pytracer does not instrument the class and returns a warning.

% \begin{listing}
%     \centering
%     \begin{lstlisting}[language=Python,style=customPython]
% def get_wrapper_function(module, qualname, name, function):
%     """Returns the instrumented function as a string.
%     This string will be compiled with the compile builtin function

%     Parameters:
%         module: Name of the function module
%         qualname: Qualified name of the function
%         name: Name of the function
%         function: Python object function

%     Returns:
%         wrapper_code: The code source of the wrapper
%     """
%     info = (id(function), function_module, qualname)
%     wrapper_code = f"""
%     def {function_wrapper_name}(*args, **kwargs):
%         return generic_wrapper({info},*args,**kwargs)"""
%     return wrapper_code
% \end{lstlisting}
%     \caption{Function to create the instrumented version of a function.
%         \pytracer uses the identifier of the function instead of the 
%         actual function to handle aliases and avoid duplicate instrumentation.}
%     \label{fig:wrapper_creation}
% \end{listing}

% \begin{listing}
%     \centering
%     \begin{lstlisting}[language=Python,style=customPython,]
% def generic_wrapper(self, info, *args, **kwargs):
%     """Generic wrapper saving inputs and outputs of the wrapped function.
%     The id_dict dict keeps a mapping between the original function identifier
%     and itself. Arguments are binded to avoid mispositioning.
%     Returns the actual output of the wrapped function.

%     Parameters:
%         info: Tuple with the function's id, function's module and function's name
%         *args: Positional calling arguments
%         **kwargs: Keyword calling arguments

%     Returns:
%         outputs: Outputs of the wrapped function
%     """
%     fid, module_name, function_name = info
%     function = original_function_cache.id_dict[fid]
%     bind = Binding(function, *args, **kwargs)
%     stack = self.backtrace()
%     time = elements()
%     self.inputs(time, module_name, function_name, function, bind.arguments, stack)
%     outputs = function(*bind.args, **bind.kwargs)
%     self.outputs(time, module_name, function_name, function, outputs, stack)
%     return outputs
% \end{lstlisting}
%     \caption{\pytracer's generic wrapper function that saves
%         the inputs and outputs of the wrapped function.}
%     \label{fig:generic_wrapper}
% \end{listing}

\subsubsection{Instrumenting classes}

\pytracer instruments classes by wrapping their callable attributes as described
previously. This instrumentation preserves class types, which avoids type
mismatches in the instrumented application. However, classes that describe
particular base types and NumPy's universal functions (class \texttt{ufunc})
contain read-only attributes that cannot be instrumented. By default, \pytracer
returns a warning when it encounters such classes.

\subsubsection{Instrumenting class instances}

\pytracer automatically instruments class instances since it instruments
classes. However, it cannot instrument or instantiate classes that have
read-only attributes --- in particular class \texttt{ufunc}, --- which is
problematic given their pervasiveness in scientific computing. In particular,
NumPy's universal functions include vectorized elementary mathematical and other
prevalent functions. Universal functions are implemented in C, which prevents
their modification.

To overcome this issue, \pytracer wraps the class instance into a transparent
wrap class (\texttt{twc}) that overloads 1) function
\texttt{\_\_getattribute\_\_}, called when an instance attribute is accessed ---
instrumented to return the attributes from the original class, and 2) function
\texttt{\_\_call\_\_}, used when an instance behaves like a function through the
() operator --- instrumented with function \texttt{generic\_wrapper}. The
\texttt{twc} returns the queried attribute of the instance instead of returning
its attributes, allowing transparent access to the instance.  Finally, \pytracer
overloads the builtin \texttt{type} function to preserve the type of the wrapped
instance.

\subsubsection{Instrumenting iterators}

In functional programming, iterators traverse containers lazily, meaning that
the next element in the sequence is only computed when the application uses it.
This technique allows for the manipulation of virtually infinite sequences with
finite memory. However, it implies that no complete mapping of the container
returned by an iterator exists in memory since the application computes each
element on the fly. Therefore, iterators are not serializable which implies that
\pytracer cannot save them to output traces. A workaround would be to convert
iterators into explicit containers, however, this would increase the memory
footprint significantly. \pytracer, therefore, does not instrument non-builtin
iterators.

\subsection{Detecting numerical instabilities}

\pytracer detects numerical instabilities by computing summary statistics across
multiple executions perturbed with numerical noise. While \pytracer's
instrumentation, trace aggregation, and visualization work with various types of
numerical noise, we experimented primarily with Monte-Carlo Arithmetic (MCA) as
it is an accurate model for floating-point errors~\addCite{\cite{verificarlo}}.

\subsubsection{Noise injection}
\label{sec:fuzzy}

Three  main  approaches  exist  to  detect  numerical  instabilities:
stochastic  arithmetic,  uncertainty  or  sensitivity analysis [23], and random
seed analysis [11].  \hl{PyTracer is agnostic to the noise model used and therefore
    works for all methods mentioned}.  For our use cases, we focus on stochastic
arithmetic since it does not make any assumption about the noise model, in
contrast to sensitivity analysis\add{.}

Stochastic arithmetic leverages randomness to estimate numerical instabilities
coming from the use of floating-point representations. The main idea is to treat
round-off errors as random variables and characterize them statistically. Monte
Carlo Arithmetic (MCA)~\cite{parker1997monte} uses this principle by introducing
two improvements: (i) a virtual precision parameter, allowing to simulate
reduced working precision, and (ii) different perturbation modes to introduce
perturbations on function inputs or output. MCA replaces each floating-point
number $x$ with a stochastic counterpart computed as:
\[
    inexact(x) =  x + \beta^{e_x - t}\xi
\]
where $\beta$ is the number base, $e_x$ is the magnitude of $x$, $t$ the virtual
precision and $\xi \in (-\frac{1}{2},\frac{1}{2})$ is a random uniform variable.
Virtual precision allows for the simulation of reduced working precision. MCA
can be applied in three modes: Random Rounding (RR), Precision Bounding (PB),
and full MCA, which apply function $inexact$ to the output, the inputs, or both,
respectively. While the RR mode is equivalent to stochastic rounding, the PB
mode can also identify catastrophic cancellations.

Stochastic arithmetic quantifies numerical error using the number of significant
digits $s$ estimated among the sampled values. A common formula to determine
this number from MCA samples is presented in~\cite{parker1997monte}:
\begin{equation}
    s = -\log_{\beta}{ \left| \dfrac{\sigma}{\mu} \right|} \label{eq:sig-digits}
\end{equation}
where $\mu$ and $\sigma$ are the sample mean and standard deviation of a
variable sampled with MCA.  
Sohier et al.~\cite{sohier2018confidence} recently provided a generalization of
this formula to include confidence intervals.

We enable MCA in Python programs via Verificarlo~\cite{verificarlo}, a
clang-based compiler~\cite{lattner2008llvm} that replaces floating-point
operations by a generic call to a configurable floating-point model. Several
floating-point models are
available~\cite{chatelain2019automatic,chatelain2019outils}, also called
backends. \pytracer leverages ``fuzzy"~\cite{kiar2020comparing}, a collection of
software tools compiled with Verificarlo. In particular, fuzzy provides
MCA-instrumented versions of the Python interpreter as well as the BLAS, LAPACK,
NumPy, SciPy, and scikit-learn libraries, which enables MCA for a wide range of
existing scientific Python programs. Fuzzy is available in Verificarlo's GitHub
organization at
\href{https://github.com/verificarlo/fuzzy}{\url{github.com/verificarlo/fuzzy}}.

\subsubsection{Trace format}

\pytracer stores traces in the pickle format, a binary format to serialize
Python objects. The main advantages of the pickle format compared to other
serialization approaches such as marshal or JSON are its ability to serialize
most Python objects and its native compression.  In addition, pickle writes
Python objects sequentially, which preserves the temporal ordering required by
subsequent trace analyses.

Traces store numerical values at the granularity of function calls. When the
application invokes a function, \pytracer saves a call input object containing
contextual information (function's name, module's function, stacktrace) and the
values of the function arguments. When the function returns, \pytracer saves a
similar call output object containing the returned value.

\subsubsection{Trace aggregation}

Once traces are available, \pytracer sequentially parses them and computes the
mean, standard deviation, and the number of significant digits (as in
Equation~\ref{eq:sig-digits}) of all saved values. \pytracer saves these summary
statistics in an HDF5 file, a hierarchical format that facilitates visual
browsing since only required information are loaded when requested by the user.
This operation critically relies on the ordering of call input and output
objects in the traces. Therefore, it assumes that the application is
single-threaded and that the control flow does not depend on a random state. 
\add{Therefore, \mbox{\pytracer} cannot merge traces resulting from a different control flow path
    and will stop the parsing when encountering one.}

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figure/pytracer_layout.pdf}
    \caption{Pytracer visualization layout.}
    \label{fig:visu-layout}
\end{figure}

To aggregate traces, \pytracer successively unpickles call objects in each
trace, checks that the call objects refer to the same function and stacktrace,
and terminates the parsing otherwise. PyTracer stores function arguments or
returned values in a NumPy array for each matching call object to compute
statistics. For compound objects such as tuples, dictionaries, or objects, it
recursively parses the fields until finding numeric types or NumPy arrays of
numeric types.

\subsection{Interactive visualization}
The third main component of \pytracer is its interactive trace visualizer built
with Plotly Dash~\cite{plotly}, a Python framework for web visualization. Plotly
Dash abstracts the low-level details of web applications in a highly
customizable development interface allowing to offload heavy computations to a
server if necessary. Figure~\ref{fig:visu-layout} presents the general layout of
the \pytracer visualizer, divided into five sub-components:
\begin{enumerate}
    \item Timeline view: This view displays the values of each input and output
          for a given function at a given invocation. By default, \pytracer
          shows the number of significant digits computed across the traces. The
          mean and standard deviation across perturbed samples are also
          available. The visualizer uses an upper triangle for inputs and a
          lower triangle for outputs. \pytracer internally represents all values
          as NumPy arrays. In the case of non-scalar values or types that cannot
          be trivially cast into NumPy arrays, \pytracer splits the value into
          several variables. Hence, \pytracer will transform a class containing
          $n$ floating-point parameters into $n$ variables. Each variable is
          prefixed with the original value and postfixed with the name of the
          parameter. Finally, non-scalar values (vectors or matrices) are
          represented by their average and can be inspected with the Zoom view
          (see below).
    \item Gantt chart view: This view displays the call graph as a Gantt chart
          representing the instrumented functions. Each function call is
          represented with a bar. The time overlaps reflect the caller/callee
          relation. For example, in the figure, \texttt{function1} calls
          \texttt{function2} that itself calls \texttt{function4} and
          \texttt{function5}. This view provides insights on the calling context
          for each function, which may influence numerical stability.
    \item  Source code view: This view displays a local snapshot of the source
          code file of the variable hovered in the timeline view.
    \item Zoom view: This view provide details about non-scalar values. For
          example, a matrix may have instabilities around its diagonal but not
          in other regions. The visualizer also provides additional metrics such
          as array shape, infinite norm, or condition number.
    \item  Functions list view: This view allows the user to select the
          functions plotted in the timeline view.
\end{enumerate}

\section{Numerical evaluations}

We evaluated \pytracer on two widely-used Python libraries:
SciPy~\cite{virtanen2020scipy}, the main library for scientific computing in
Python, and  scikit-learn~\cite{pedregosa2011scikit}, a reference
machine-learning framework that implements a wide range of state-of-the-art
algorithms. For each library, we executed the software examples provided in
their respective GitHub repositories. We traced each example five times, having
activated MCA in the Python interpreter \add{as well as} BLAS, LAPACK, NumPy, SciPy, and
scikit-learn. \add{Only instrumented the Python interpreter would lead to
    only perturbe pure Python floating-point operations and thus to miss
    operations in compiled codes (i.e. Cython and shared libraries).}
\hl{We repeated the experiment with two MCA modes, Random Rounding
    (RR) and Full MCA, resulting in 2 experiments for each example}. We used the
default virtual precision in Verificarlo, namely 24 bits for single-precision
floating-point values and 53 bits for double-precision values, which simulates
machine error. We used Python-3.8.5, NumPy-1.19.1, SciPy-1.5.4, and
scikit-learn-0.23.2.

\subsection{Results overview}

Table~\ref{tab:pytracer_results_summary} summarizes \pytracer's outcome on SciPy
and scikit-learn examples. Columns \textit{trace} and \textit{parse} report on
the tracing and aggregation steps with three possible outcomes: 1) green for
success, meaning that the execution ended well and no issues were raised by the
application, 2) red for numerical error in the application, meaning that the
execution raised a Python error before the end, and 3) orange for errors due to
the fuzzy noise injection that are not considered numerical errors, meaning that
the execution raised a Python error but due to a perturbation that should not
have been introduced (i.e. a floating-point value representing an integer).
Overall, the tracing and parsing of SciPy and scikit-learn examples in Random
Rounding mode was entirely successful for 32/40 examples and showed an average
precision loss of 8 bits. The remaining examples failed due to numerical errors
in the libraries, which we discuss hereafter. The instrumentation with Full MCA
is more invasive and reduced the number of successful executions to 12/40. Among
the 28/40 failed executions, one failed due to an actual numerical error
(\texttt{Bayesian Ridge Regression} example) and the other ones failed due to
perturbations that should not have been introduced.

Column \textit{sigbits} in Table~\ref{tab:pytracer_results_summary} summarizes
the precision for the main outputs of SciPy and scikit-learn examples. We
measured the precision in significant bits on the final primary example output
by taking the element-wise mean for non-scalar values and, if the example
produces many outputs, the average value among all of them. The numerical
quality across all examples varies from 0 to 52, with an average of 44
significant bits. For both modes, we note that only two examples failed due to a
numerical error (\texttt{face\_recognition} for RR and \texttt{Bayesian Ridge
    Regression} for Full MCA).

\begin{table*}[]
    \centering
    \small
    \begin{subfigure}[t]{\linewidth}
        \centering
        \begin{tabular}{|lll|c|c|c|c|c|c|}
            \hline
            \multicolumn{3}{|c}{ \multirow{2}{*}{Application} }                 & \multicolumn{3}{|c|}{Random Rounding (RR)}                                & \multicolumn{3}{c|}{Full MCA}                                                                                      \\
            \cline{4-9}
                                                                                &                                                                           &                                                          & trace  & parse    & sigbits & trace  & parse  & sigbits \\
            \hline
            \multicolumn{1}{|c|}{ \multirow{20}{2em}{ \rotatebox{90}{SciPy} } } & 
            \multicolumn{1}{c|}{ \multirow{3}{2em}{ \rotatebox{90}{FFT} }}      & \discreteCosineRF                                                         & \valid                                                   & \valid & \s{52}   & \warn   & \cross & \s{47}           \\ 
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \fftOneDRf                                               & \valid & \valid   & 41      & \warn  & \cross & -       \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \fftTwoDRf                                               & \valid & \valid   & \s{52}  & \valid & \valid & \s{52}  \\
            \cline{2-9}
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{  \multirow{5}{*}{ \rotatebox{90}{Interpolation} } }
                                                                                & \interOneDRf                                                              & \valid                                                   & \valid & 51       & \warn   & \cross & -                \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \multiRf                                                 & \valid & \valid   & \s{51}  & \warn  & \cross & \s{-}   \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \bsplineRf                                               & \valid & \valid   & 10      & \valid & \valid & 10      \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \splineOneDRf                                            & \valid & \cross   & \s{52}  & \warn  & \cross & \s{52}  \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \splineTwoDRf                                            & \valid & \valid   & 10      & \warn  & \cross & -       \\
            \cline{2-9}
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{ \multirow{12}{2em}{ \rotatebox{90}{Optimization} } }
                                                                                & \bfgsRf                                                                   & \valid                                                   & \valid & \s{48}   & \valid  & \valid & \s{46}           \\
            \cline{3-9}
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \globalRf                                                & \valid & \cross   & 17      & \warn  & \cross & -       \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & | SHGO\footnote{Simplicial Homology Global Optimization} & \na    & \na      & \s{25}  & \na    & \na    & \s{-}   \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & | Dual Annealing                                         & \na    & \na      & 4       & \na    & \na    & -       \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & | Differential Evolution                                 & \na    & \na      & \s{0}   & \na    & \na    & \s{-}   \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & | Basin Hopping                                          & \na    & \na      & 40      & \na    & \na    & -       \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & | SHGO Sobol                                             & \na    & \na      & \s{45}  & \na    & \na    & \s{-}   \\
            \cline{3-9}
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \slsqpRf                                                 & \valid & \valid   & 45      & \valid & \valid & 44      \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \lsmRf                                                   & \valid & \valid   & \s{48}  & \warn  & \cross & \s{-}   \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \nelderRf                                                & \valid & \valid   & 44      & \valid & \valid & 42      \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \ncgRf                                                   & \valid & \valid   & \s{48}  & \valid & \cross & \s{47}  \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \rootRf                                                  & \valid & \cross   & 52      & \warn  & \cross & 50      \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \rootLargRf                                              & \valid & \cross   & \s{29}  & \valid & \cross & \s{26}  \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \rootLargePredRf                                         & \valid & \cross   & 42      & \valid & \cross & 42      \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \trustExactRf                                            & \valid & \valid   & \s{48}  & \valid & \valid & \s{46}  \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \trustTruncRf                                            & \valid & \valid   & 48      & \valid & \valid & 46      \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \trustNCGRf                                              & \valid & \valid   & \s{52}  & \valid & \valid & \s{47}  \\
            \hline
            \multicolumn{2}{|c|}{ \multirow{20}{2em}{ \rotatebox{90}{Scikit-learn} } }
                                                                                & \AdaboostRf                                                               & \valid                                                   & \valid & 48       & \warn   & \cross & -                \\
            \cline{3-9}
            \multicolumn{2}{|c|}{}                                              & \brrRf                                                                    & \valid                                                   & \valid & \s{25}   & \cross  & \cross & \s{-}            \\
            \multicolumn{2}{|c|}{}                                              & | $1^{st}$ dataset                                                        & \na                                                      & \na    & 25       & \na     & \na    & -                \\
            \multicolumn{2}{|c|}{}                                              & | $2^{nd}$ dataset                                                        & \na                                                      & \na    & \s{nan}  & \na     & \na    & \s{-}            \\
            \cline{3-9}
            \multicolumn{2}{|c|}{}                                              & \onlineClassifierComparisonRf                                             & \valid                                                   & \valid & 20       & \warn   & \cross & -                \\
            \multicolumn{2}{|c|}{}                                              & \kmeansRf                                                                 & \valid                                                   & \valid & \s{52}   & \valid  & \cross & \s{-}            \\
            \multicolumn{2}{|c|}{}                                              & \covarianceRf                                                             & \valid                                                   & \cross & 48       & \warn   & \cross & -                \\
            \multicolumn{2}{|c|}{}                                              & \decisionRf                                                               & \valid                                                   & \valid & \s{50}   & \valid  & \valid & \s{50}           \\
            \multicolumn{2}{|c|}{}                                              & \digitsRf                                                                 & \valid                                                   & \valid & 52       & \warn   & \cross & 52               \\
            \multicolumn{2}{|c|}{}                                              & \faceRf                                                                   & \cross                                                   & \valid & \s{-}    & \warn   & \cross & \s{-}            \\
            \multicolumn{2}{|c|}{}                                              & \penaltyRf                                                                & \valid                                                   & \valid & -        & \valid  & \valid & -                \\
            \multicolumn{2}{|c|}{}                                              & \lassoRf                                                                  & \valid                                                   & \valid & \s{48}   & \warn   & \cross & \s{47}           \\
            \multicolumn{2}{|c|}{}                                              & \hyperplaneRf                                                             & \valid                                                   & \valid & 3        & \valid  & \cross & -                \\
            \multicolumn{2}{|c|}{}                                              & \mnistRf                                                                  & \valid                                                   & \valid & \s{43}   & \warn   & \valid & \s{-}            \\
            \multicolumn{2}{|c|}{}                                              & \multitaskRf                                                              & \valid                                                   & \valid & 50       & \warn   & \valid & 48               \\
            \multicolumn{2}{|c|}{}                                              & \ompRf                                                                    & \valid                                                   & \valid & \s{52}   & \valid  & \valid & \s{52}           \\
            \multicolumn{2}{|c|}{}                                              & \pcaRf                                                                    & \valid                                                   & \valid & 48       & \valid  & \valid & 48               \\
            \cline{3-9}
            \multicolumn{2}{|c|}{}                                              & \robustRf                                                                 & \valid                                                   & \valid & \s{50}   & \warn   & \cross & \s{48}           \\
            \multicolumn{2}{|c|}{}                                              & | Linear                                                                  & \na                                                      & \na    & 48       & \na     & \na    & 46               \\
            \multicolumn{2}{|c|}{}                                              & | RANSAC                                                                  & \na                                                      & \na    & \s{50}   & \na     & \na    & \s{48}           \\
            \cline{3-9}
            \multicolumn{2}{|c|}{}                                              & \toyRf                                                                    & \valid                                                   & \valid & 35       & \warn   & \cross & -                \\
            \cline{3-9}
            \multicolumn{2}{|c|}{}                                              & \svmRf                                                                    & \valid                                                   & \valid & \s{43.5} & \valid  & \cross & \s{-}            \\
            \multicolumn{2}{|c|}{}                                              & | Non-weighted                                                            & \na                                                      & \na    & 43       & \na     & \na    & -                \\
            \multicolumn{2}{|c|}{}                                              & | Weighted                                                                & \na                                                      & \na    & \s{44}   & \na     & \na    & \s{-}            \\
            \cline{3-9}
            \multicolumn{2}{|c|}{}                                              & \tomographyRf                                                             & \valid                                                   & \cross & -        & \warn   & \cross & -                \\
            \multicolumn{2}{|c|}{}                                              & \weightedRf                                                               & \valid                                                   & \valid & \s{44}   & \valid  & \cross & \s{-}            \\
            \hline
        \end{tabular}
    \end{subfigure}
    \caption{\pytracer execution summary on SciPy and Scikit-learn examples.
        % \tristan{Replace FFT with SciPy FFT, Interpolation with SciPy Interpolation and Optimization with SciPy Optimization, or add a SciPy header to these rows.} \tristan{Nitpick: you could add a color scheme to sigbits (for instance shades of blue), to make the column more easily readable.} \\
        \emph{trace}: outcome of \pytracer tracing, \emph{parse}: outcome of \pytracer trace parsing and aggregation, \emph{result}: precision for main outputs in significant bits. \valid: successful run, \warn: error from invalid noise injection, \cross: numerical error or divergent traces. \textit{nan} stands for (\texttt{Not-A-Number}). Example names are linked to their source code.}
    \label{tab:pytracer_results_summary}
\end{table*}

\subsection{Effect of MCA modes}
\label{sec:impact_mca_modes}
Random Rounding perturbs only the result of a floating-point operation while
Full MCA perturbs both the inputs and the result. Therefore, Full MCA may
trigger catastrophic cancellations while Random Rounding only introduces
round-off errors which generally have a lesser impact. Moreover, Random Rounding
preserves exact operations --- operations where the result can be exactly
represented at the virtual precision --- while Full MCA does not.

\pytracer can only aggregate traces obtained from the same control flow path. In
particular, program branches triggered by an unstable floating-point comparison
lead to different control flows and result in parsing errors. For example,
comparing a residual error or the distance between two points to a strict
threshold in an iterative scheme generally results in unstable branches. Full
MCA even perturbs exact values such as integers represented with floating-point
values, leading to many execution failures. For instance, the \texttt{fft1}
example from SciPy raised an \texttt{ValueError: operands could not be broadcast
    together with shapes (601,) (600,)} error due to an array shape mismatch. A
closer inspection showed that one of the arrays involved in the error was
created with the NumPy function \texttt{linspace} that itself calls function
\texttt{arange} (Listing~\ref{fig:pyarray_range_code}). Lines 3174-3178 show
that the function stores the array size in a floating-point value that will be
perturbed by Full MCA, as shown in Table~\ref{tab:mca_result_linspace},
resulting in array sizes that fluctuate between N and N+1. It is worth noting
that 78\% of execution issues encountered in Full MCA mode instrumentation are
due to this type of ``typing" errors.  
Those code regions should not be perturbed by MCA since the injected noise does
not make sense.

\begin{listing}
    \begin{lstlisting}[language=C,style=customC,numbers=left, firstnumber=3163, mathescape=true]
/* NUMPY_API Arange, */
NPY_NO_EXPORT PyObject *
PyArray_Arange(double start, double stop, double step, int type_num) {
    npy_intp length; 
    PyArrayObject *range; $\lstsetnumber{\ldots}$
    ...$\lstresetnumber\setcounter{lstnumber}{3173}$
    double delta, tmp_len; $\lstsetnumber{\ldots}$
    ...$\lstresetnumber\setcounter{lstnumber}{3176}$
    delta = stop - start; 
    tmp_len = delta/step; $\lstsetnumber{\ldots}$
    ...$\lstresetnumber\setcounter{lstnumber}{3189}$
        length = _arange_safe_ceil_to_intp(tmp_len); $\lstsetnumber{\ldots}$
    ...$\lstresetnumber\setcounter{lstnumber}{3201}$
    range = (PyArrayObject *)PyArray_New(&PyArray_Type, 1, &length, type_num, NULL, NULL, 0, 0, NULL);
    \end{lstlisting}
    \caption{Source code of the \texttt{PyArray\_Range} Cython function called by NumPy function \texttt{linspace} to create an array of equally-spaced elements. The temporary array size assigned in line 3178 is stored as a floating-point value and is therefore perturbed in Full MCA mode, leading to differences amplified by the use of ceil rounding at line 3190 and resulting in different array sizes across MCA-perturbed executions. See also Table~\ref{tab:mca_result_linspace}.}
    \label{fig:pyarray_range_code}

\end{listing}

\begin{table}[t]
    \centering
    % \begin{tabular}{ccccccc}
    %     \thickhline
    %     \textbf{Mode} & \textbf{start} & \textbf{stop} & \textbf{step} & \textbf{delta}                & \textbf{tmp\_len} & \textbf{ceil(tmp\_len)} \\
    %     \hline
    %     Exact         & 0              & 600           & 1             & 600                           & 600               & 600                     \\
    %     RR            & 0              & 600           & 1             & 600 $\pm$ 0.0                 & 600 $\pm$ 0.0     & 600 $\pm$ 0.0           \\
    %     MCA           & 0              & 600           & 1             & 599.9999999999999 $\pm$ 8e-14 & 600.0 $\pm$ 8e-14 & 600.01 $\pm$ 0.0995     \\
    %     \thickhline        
    % \end{tabular}
    \begin{tabular}{l|ccc}
        \hline                                                   
        \textbf{Mode}           & Exact & RR            & MCA                   \\
        \hline
        \texttt{start}          & 0     & 0             & 0                     \\
        \texttt{stop}           & 600   & 600           & 600                   \\
        \texttt{step}           & 1     & 1             & 1                     \\ 
        \texttt{delta}          & 600   & 600 $\pm$ 0.0 & 599.9...9 $\pm$ 8e-14 \\
        \texttt{tmp\_len}       & 600   & 600 $\pm$ 0.0 & 600.0 $\pm$ 8e-14     \\
        \texttt{ceil(tmp\_len)} & 600   & 600 $\pm$ 0.0 & 600.01 $\pm$ 0.0995   \\
        \thickhline        
    \end{tabular}
    \caption{Results obtained through lines 3177-3190 in Listing~\ref{fig:pyarray_range_code}. In contrast with RR, Full MCA does not preserve exact operations which leads to an array length oscillating between 600 and 601.}
    \label{tab:mca_result_linspace}
\end{table}

\subsection{SciPy}
\label{sec:scipy_tests}

We tested \pytracer on SciPy examples provided in the tutorial website section.
SciPy is organized in several libraries targeting specific computing domains. We
focused on \texttt{fftpack} (Fast Fourier Transform routines),
\texttt{interpolate} (interpolation and smoothing splines), and
\texttt{optimize} (Optimization and root-finding routines).

\subsubsection{FFT}

This package has three examples : \texttt{discrete\_cosine\_transform},
\texttt{fft\_1D}, and \texttt{fft\_2D}. All computations are done in double
precision. The results for \texttt{discrete\_cosine\_transform},\texttt{fft\_1D}
and  \texttt{fft\_2D} show precise numerical results with 48 significant bits on
average. We observe that the FFT computation is stable to noisy inputs.
Figures~\ref{fig:fft1D_inputs} (inputs) and~\ref{fig:fft1D_outputs} (outputs)
show the mean and standard deviation of the MCA samples. The three columns
correspond to the FFT computations of: 
\begin{enumerate}
    \item $z_1 = \sin(50 . 2\pi . x_i) + \dfrac{1}{2} \sin(80 . 2\pi . x_i),\; x_i = \frac{i}{600},\; i \in [0,600]$
    \item $\mathrm{blackman}(400) \times z_1$
    \item $ z_2= e^{50 . i 2\pi . x_i} + \dfrac{1}{2} e^{-80 . i2\pi .x_i },\; x_i = \frac{i}{800},\; i \in [0,400] $
\end{enumerate}

\add{The} top row in Figures~\ref{fig:fft1D_inputs} and~\ref{fig:fft1D_outputs} show the
input mean value while the bottom row shows the standard deviation. The points
with low magnitude in Figure~\ref{fig:fft1D_inputs} correspond to inputs near 0
when the input of $\sin$ or $\exp$ is close to $k\pi$, $k \in \mathbb{Z}$.
Indeed, since MCA introduces a slight noise, the result is not exactly 0. We can
see in Figure~\ref{fig:fft1D_inputs} that the maximal noise introduced by MCA on
the inputs is in the order of $10^{-14}$, two orders of magnitude higher than
the $ulp \simeq 10^-16$ for double precision. This slight noise introduced can
be interpreted as white noise on the input signal. We can see on the bottom row
of Figure~\ref{fig:fft1D_outputs} that the frequency noise is of the same order
of of magnitude as the input noise, which is expected. The two peaks at x=38 and
x=562 corresponds to the actual frequencies of the input signal.

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figure/FFT/fft_x.pdf}
    \caption{Absolute value of the mean and standard deviation values for
        fft 1D inputs within RR mode (log scale). The real part is shown in blue and the imaginary part in orange.}
    \label{fig:fft1D_inputs}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figure/FFT/fft_y.pdf}
    \caption{Absolute value of the mean and standard deviation values
        for fft 1D outputs within RR mode (log scale).}
    \label{fig:fft1D_outputs}
\end{figure}

\subsection{Interpolation}

This package has five examples: \texttt{interpolation\_1D},
\texttt{multivariate\_data}, \texttt{bspline}, \texttt{spline\_1D} and
\texttt{spline\_2D}.

\texttt{interpolation\_1D} interpolates function $f(x)=\cos(\frac{-x^2}{9})$
with 11 points $x\in[0,10]$. For the five interpolation methods tested (linear,
cubic, nearest, previous, next), the solution found is precise, up to 51 bits
out of the 53 available.

\texttt{multivariate\_data} interpolates the grid $f(x,y)=x(1-x)\cos(4\pi.x)
    \sin(4\pi.y^2)^2$ with $(x,y) \in [0,1] \times [0,1]$. It samples 1,000 points
for each coordinate and uses three interpolation methods: nearest, linear and
cubic. Our results show a precision of 51 bits on average for the three methods.

\texttt{bspline} compares two edge filters: \texttt{bisplev} evaluates a
bivariate B-spline and its derivative whereas \texttt{convolved2d} convolves two
2-dimensional arrays. Figure~\ref{fig:bspline_rr} shows the significant bits,
mean, and standard deviation maps for the two methods as well as the input image
(Fig.~\ref{fig:bspline_original_image}). The middle row shows \texttt{sepfir2d}
results and the bottom row shows \texttt{convol2d} results. Both methods exhibit
similar precision, with 11 significant bits on average. Standard deviations maps
on Figures~\ref{fig:bspline_bisplev_std} and~\ref{fig:bspline_convol2d_std} show
that regions with low spatial frequencies correspond to regions with low
standard deviation, similar to the FFT results. The SciPy tutorial mentions that
\texttt{bisplev} is faster than \texttt{convol2d}. The comparable numerical
precision observed in our experiments reinforce the use of \texttt{bisplev} over
\texttt{convol2d} in this example.

\texttt{spline\_1D} computes the 1D spline interpolation for $f(x)=\sin(x)$ on
points $x=\frac{2\pi k}{8}, k \in [0, 10]$. The spline interpolation uses
\texttt{splrep} to build the spline representation and \texttt{splrev} to
evaluate the spline at any point. It also computes the integral by using
\texttt{splint} and the root finder \texttt{sproot}. Although the traces
diverge, we can analyze the partial aggregation for the \texttt{splrep},
\texttt{splev} and \texttt{splint}. The three functions show results with an
average precision of 51 bits with RR mode.

\texttt{spline\_2D} computes the 2D spline interpolation for function
$z=(x+y)e^{-6(x^2+y^2)}$, with $(x,y) \in [-1,1]\times[-1,1]$ and a sampling of
21 points for each coordinate. \texttt{bisplrep} builds the spline
representation with the $21 \times 21$ points of $z$ over 71 grid sampling
points for $(x,y)$. Figure~\ref{fig:spline2d_rr} shows the results of the spline
evaluation with \texttt{bisplev}.  We observe a significant loss of precision in
places, which follows an interesting pattern that could be further investigated.

\begin{figure}
    \centering
    \begin{subfigure}{.3\linewidth}
        \includegraphics[width=\linewidth]{figure/bspline/original_image.pdf}
        \caption{Original image.}
        \label{fig:bspline_original_image}
    \end{subfigure}\\
    \begin{subfigure}{0.3\linewidth}
        \includegraphics[width=\linewidth]{figure/bspline/bspline_sig.pdf}
        \caption{\centering\texttt{sepfir2d} significant bits}
        \label{fig:bspline_bisplev_sig}
    \end{subfigure}
    \begin{subfigure}{0.3\linewidth}
        \includegraphics[width=\linewidth]{figure/bspline/bspline_mean_log.pdf}
        \caption{\centering\texttt{sepfir2d} mean (log)}
        \label{fig:bspline_bisplev_mean}
    \end{subfigure}
    \begin{subfigure}{0.3\linewidth}
        \includegraphics[width=\linewidth]{figure/bspline/bspline_std_log.pdf}
        \caption{\centering\texttt{sepfir2d} standard deviation (log)}
        \label{fig:bspline_bisplev_std}
    \end{subfigure}
    \begin{subfigure}{0.3\linewidth}
        \includegraphics[width=\linewidth]{figure/bspline/convol2d_sig.pdf}
        \caption{\centering\texttt{convol2d} significant bits}
        \label{fig:bspline_convol2d_sig}
    \end{subfigure}
    \begin{subfigure}{0.3\linewidth}
        \includegraphics[width=\linewidth]{figure/bspline/convol2d_mean_log.pdf}
        \caption{\centering\texttt{convol2d} mean (log)}
        \label{fig:bspline_convol2d_mean}
    \end{subfigure}
    \begin{subfigure}{0.3\linewidth}
        \includegraphics[width=\linewidth]{figure/bspline/convol2d_std_log.pdf}
        \caption{\centering\texttt{convol2d} standard deviation (log)}
        \label{fig:bspline_convol2d_std}
    \end{subfigure}
    \caption{\texttt{bspline} results within RR mode. \texttt{sepfir2d} and
        \texttt{convol2d} have a similar precision.
    }
    \label{fig:bspline_rr}
\end{figure}

\begin{figure}
    \centering
    \begin{subfigure}{.3\textwidth}
        \centering
        \includegraphics[width=\linewidth]{figure/spline_2d/bisplev_mean.pdf}
        \caption{Mean value}
        \label{fig:bisplev_mean}
    \end{subfigure}
    \begin{subfigure}{.3\textwidth}
        \centering
        \includegraphics[width=\linewidth]{figure/spline_2d/bisplev_mean_log.pdf}
        \caption{Absolute mean value (log)}
        \label{fig:bisplev_mean_log}
    \end{subfigure}
    \begin{subfigure}{.3\textwidth}
        \centering
        \includegraphics[width=\linewidth]{figure/spline_2d/bisplev_std_log.pdf}
        \caption{Standard deviation (log)}
        \label{fig:bisplev_std_log}
    \end{subfigure}
    \caption{2D Spline interpolation results. Figure~\ref{fig:bisplev_mean} shows the mean result
        of the 2D-spline interpolation. Figures~\ref{fig:bisplev_mean_log} and~\ref{fig:bisplev_std_log}
        show respectively in the logarithmic scale the absolute mean value and the standard deviation 
        of the 5 samples run within RR mode. 
        % \tristan{any reason why you are showing the ieee result?} \tristan{can you use the same color maps as in figure 8?}
    }
    \label{fig:spline2d_rr}
\end{figure}

\subsubsection{Optimization}

This package has eleven examples. Seven examples involve the same optimization
problem solved with and without constraints using different methods. One example
optimizes a chemical problem under constraints. The last three examples solve a
different problem each.

\paragraph{Unconstrained minimization of multivariate scalar functions}
Four Quasi-Newton-Raphson methods minimize function $f$ by using Newton's step:
$x_{k+1} = x_{k} - B^{-1}(x_k)\nabla f(x_k)$, where $\nabla$ and $B$ are the
gradient and the approximation of the Hessian matrix of $f$. The
\texttt{Broyden-Fletcher-Goldfarb-Shanno}~\cite{BFGS} (BFGS) and
\texttt{Newton-Conjugate-Gradient}~\cite{nocedal2006numerical} (NCG) are line
search methods that approximate $H$ by adding two symmetric rank-one matrices
and by using the Conjugate-Gradient method. The \texttt{Trust-Region Truncated
    Generalized Lanczos / Conjugate Gradient}~\cite{gould1999solving} (TR-Krylov),
\texttt{Trust-Region Newton-Conjugate-Gradient} (TR-NCG), and
\texttt{Trust-Region Nearly Exact}~\cite{nocedal2006numerical} (TR-Exact) are
trust-region methods approximating $H$ by solving the trust-region subproblem
restricted to a truncated Krylov subspace and by solving nonlinear equations for
each quadratic subproblem. In addition, the
\texttt{Nelder-Mead}~\cite{singer2009nelder} simplex-based method iteratively
transforms a simplex until its vertices are getting closer to a local minima of
the function. All methods are applied to the Rosenbrock function.

\paragraph{Constrained minimization of multivariate scalar functions}
Two optimizers solve an optimization problem by taking into account constraints.
The \texttt{Sequential Least SQuares Programming} uses the
SLSQP~\cite{kraft1988software} method to optimize the Rosenbrock function. The
\texttt{Least-squares minimization} uses the Trust Region Reflective
algorithm~\cite{li1993centering} to solve a fitting problem from an enzymatic
reaction.

The SLSQP and unconstrained minimization methods minimize the
\textbf{Rosenbrock} function of $N$ variables that reaches its minimum (0) when
$x_i=1,\; \forall i \leq N-1$:
$f(x) = \sum_{i=1}^{N-1} 100(x_{i+1}-x^2_i)^2 + (1-x_i)^2$.
Figure~\ref{fig:unconstrained_optimization} shows the number of significant bits
of the solution for the different methods used. The $(p)$ notation for the NCG,
Trust-Krylov, and Trust-NCG methods in the legend refers to the variant where
the Hessian matrix is replaced by a function computing the matrix-vector product
of the Hessian and an arbitrary vector, to save memory and time. All the methods
have a good precision ranging from 43 to 53 significant bits. The Nelder-Mead
method is the least precise one with $\simeq$ 43 significant bits on average,
while the Trust-Region-Newton-CG is the most precise one with $\simeq$ 53
significant bits, the highest precision achievable in double precision. The
remaining methods have a similar precision.

The SLSQP also minimizes the Rosenbrock function with the following constraint:
$x_0$ + $2x_1$ $\leq$ 1 
$x_0^2$ + $x_1$ $\leq$ 1 
$x_0^2$ - $x_1$ $\leq$ 1 
$2x_0$ + $x_1$ = 1 
0 $\leq$ $x_0$ $\leq$ 1 
-0.5 $\leq$ $x_1$ $\leq$ 2. 
% \begin{eqnarray*}
%     x_0 + 2x_1 \leq 1 \\
%     x_0^2 + x_1 \leq 1 \\
%     x_0^2 - x_1 \leq 1 \\
%     2x_0 + x_1 = 1 \\
%     0 \leq x_0 \leq 1 \\
%     -0.5 \leq x_1 \leq 2 
% \end{eqnarray*}
The problem has a unique solution $[x_0, x_1] = [0.4149, 0.1701]$. Results
obtained with RR show a precision of 47 and 44 significant bits for the first
and the second solution. %\tristan{which method is tested?}.

The \texttt{least-square minimization} example solves a fitting problem from an enzymatic reaction~\cite{kowalik1968analysis} with 11 residuals defined as:
\begin{eqnarray*}
    f_i(x) &=& \frac{x_0(u_i^2 + u_ix_1)}{u_i^2 + u_ix_2+x_3}-y_i,\; i=0,...,10 \\
    &0& \leq x_j \leq 100,\; j=0,..,3
\end{eqnarray*}
where where $y_i$ are measurement values, $u_i$ are values of the independent
variable, and $x_i$ the unknown. Results within RR show a precision of 51, 46,
47, and 48 significant bits for each component of the solution
respectively.% \tristan{which solver is used?}.

\paragraph{Root finding}
% We tested three root finder algorithms with one for small problems using the hybrid Powell and Levenberg-Marquardt methods  (\texttt{root\_finding}) and two for large problems using a Krylov approximation for the inverse Jacobian with and without the help of a preconditioner (\texttt{root\_finding\_large\_preconditionned} and \texttt{root\_finding\_large}).

The \texttt{root\_*} examples use three algorithms to find the root of
non-linear equations. \texttt{root\_finding} uses the hybrid Powell method to
solve a single-variable transcendental equation and the Levenberg-Marquardt
method for a set of non-linear equations. \texttt{root\_finding\_large} and
\texttt{root\_finding\_large\_preconditioned} examples use the Krylov method to
approximate the inverse Jac\add{o}bian with and without the help of a preconditioner.
They solve an integrodifferential equation on a square $(x,y) \in [0,1] \times
    [0,1]$, $P(x,1)=1$, and $P=0$ elsewhere on the boundary. The function $P$ is
discretized with a Cartesian grid $P_{n,n}=P(nh,nh)$, with $n=75$, and its
derivatives are approximated by $\partial_x^2P(x,y) \simeq (P(x+h,y) - 2P(x,y) +
    P(x-h,y))/h^2$.

% the two following equations: $x + 2\cos(x) = 0$ and 
% the set $x_0\cos(x_1)=4$; $x_0x_1-x_1 = 5$ while the 
% \texttt{root\_finding\_large} and \texttt{root\_finding\_large\_preconditioned} tests 
% solve the equation:
% \[
%     (\partial_x^2 + \partial_y^2)P + 5 \left(\int_0^1 \int_0^1 cosh(P)\, dx dy \right)^2 = 0
% \]

All the examples diverge even within RR, which is not surprising since
root-finding methods can easily take a few extra steps to find the root
depending on the starting point. Nevertheless, for the
\texttt{root\_finding\_large} example, 3 traces among the 5 samples can be
merged for both RR and MCA modes. Figure~\ref{fig:root_finding_large} shows the
mean and standard deviation of the result for these 3 samples. We can see that a
similar solution was found for RR and Full MCA. Full MCA leads to a lower
precision than RR although both standard deviation maps remain in the same order
of magnitude. Finally, it is interesting to note the impact of preconditioning
on the result numerical quality: as Table~\ref{tab:pytracer_results_summary}
shows, the precision doubles between \texttt{root\_finding\_large} and
\texttt{root\_finding\_large\_preconditioned}.

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figure/unconstrained_optimization_comparison.pdf}
    \caption{Comparison of results precision for different optimization solvers within RR mode.}
    \label{fig:unconstrained_optimization}
\end{figure}

\begin{figure}
    \centering
    \begin{subfigure}{0.45\linewidth}
        \includegraphics[width=\linewidth]{figure/root_finding/solution_mean_RR.pdf}
        \caption{Mean solution (RR)}
        \label{fig:mean_solution_rr}
    \end{subfigure}
    \begin{subfigure}{0.45\linewidth}
        \includegraphics[width=\linewidth]{figure/root_finding/solution_std_RR.pdf}
        \caption{Standard deviation solution (RR)}
        \label{fig:stdev_rr}
    \end{subfigure}

    \begin{subfigure}{0.45\linewidth}
        \includegraphics[width=\linewidth]{figure/root_finding/solution_mean_MCA.pdf}
        \caption{Mean solution (Full MCA)}
        \label{fig:mean_solution_mca}
    \end{subfigure}
    \begin{subfigure}{0.45\linewidth}
        \includegraphics[width=\linewidth]{figure/root_finding/solution_std_MCA.pdf}
        \caption{Standard deviation (Full MCA)}
        \label{fig:stdev_mca}
    \end{subfigure}
    \caption{Solution of \texttt{root\_finding\_large} example.
        The standard deviation has two maximum for both RR and Full MCA modes. The standard deviation remains lower than
        the stopping criterion threshold fixed at $6.10^{-6}$. }
    \label{fig:root_finding_large}
\end{figure}

% \begin{itemize}
% \item broyden: RR 47 bits solution / MCA divergence 
% \item global\_optimization: RR divergence / MCA issues
% \item sqlsp: RR 46 bits solution / MCA 46 bits solution
% \item least\_square\_minimization : RR 48 bits solution / MCA index issue
% \item nelder\_mean\_simplex : RR 43 bits solution / MCA stable
% \item newton\_conjugate\_gradient: RR 48 bits solution / MCA divergence
% \item root\_finding\_large\_preconditionned : RR divergence / MCA divergence 3/5 converged
% \item root\_finding\_large : RR divergence / MCA divergence (fgmres very low!) 3/5 converged
% \item trust\_region\_exact : RR 47 bit solution / RR 45 bit solution
% \item trust\_region\_truncated\_generalized\_lanczos : RR 48 bits solution / MCA 46 bits solution
% \end{itemize}

% \begin{figure}
%     \centering
%     \includegraphics[width=\linewidth]{figure/nelder_mead/timeline_sig.png}
%     \caption{Nelder-Mead: average of number of significant bits over time. 
%     Solver ends when accuracy is reaching the tolerance.
%     }
%     \label{fig:my_label}
% \end{figure}

\subsection{Scikit learn}
\label{sec:sklearn_tests}

Scikit-learn offers a well-supplied and documented list of examples that
facilitates its use with Pytracer. Among the several available examples, we
choose the following representative set listed in
Table~\ref{tab:pytracer_results_summary}. The following paragraphs focus on the
most interesting stability results observed across all experiments.
% \tristan{Shall we just refer to Table 4 and remove this lengthy list? You could add the links in the table and mention in the table caption that the tests are linked to their source code}:

% \subsubsection{Classifier comparisions}

% This test compares the performance of 6 online solvers on the hand-written digits dataset.
% These 6 solvers are:
% \begin{itemize}
%     \item Stochastic Gradient Descent 
%     \item Averaged Stochastic Gradient Decent
%     \item Perceptron
%     \item Passive-Aggressive I \& II
%     \item SAG: Minimizing Finite Sums with the Stochastic Average Gradient
% \end{itemize}

% Each solver is trained and predicts 20 times on 6 different training size ratio.
% Figure~\ref{fig:classifier_comparisons_general} shows the all functions
% traced by Pytracer over time. We can see the estimated number of significant digits varies a lot between 
% values from 1 to 80 bits. Figure~\ref{fig:classifier_comparisons_mean_predictions} shows 
% the average prediction score for each classifiers. The yellow points 
% represent the average correct prediction score for one round and one classifier (line. 48) while blue points
% tshows average predictions over the 20 rounds (line. 49).
% We can see that precision is pretty low with an average solution below 10 significant bits, with 
% an exception for values between the 24k and 32k invocation that correspond to the Perceptron
% classifier. Indeed, this classifier is pretty stable with an average number of significant digits around 52 bits.

% \begin{figure}
%     \begin{lstlisting}[language=Python,style=customPython,numbers=left, firstnumber=19]
% heldout = [0.95, 0.90, 0.75, 0.50, 0.01]
% rounds = 20
% X, y = datasets.load_digits(return_X_y=True)

% classifiers = [
%     ("SGD", SGDClassifier(max_iter=100)),
%     ("ASGD", SGDClassifier(average=True)),
%     ("Perceptron", Perceptron()),
%     ("Passive-Aggressive I", PassiveAggressiveClassifier(loss='hinge',C=1.0, tol=1e-4)),
%     ("Passive-Aggressive II", PassiveAggressiveClassifier(loss='squared_hinge',C=1.0, tol=1e-4)),
%     ("SAG", LogisticRegression(solver='sag', tol=1e-1, C=1.e4 / X.shape[0]))
% ]

% xx = 1. - np.array(heldout)

% for name, clf in classifiers:
%     print("training %s" % name)
%     rng = np.random.RandomState(42)
%     yy = []
%     for i in heldout:
%         yy_ = []
%         for r in range(rounds):
%             X_train, X_test, y_train, y_test = train_test_split(X, y, test_size=i, random_state=rng)
%             clf.fit(X_train, y_train)
%             y_pred = clf.predict(X_test)
%             yy_.append(1 - np.mean(y_pred == y_test))
%         yy.append(np.mean(yy_))
%     plt.plot(xx, yy, label=name)
%     \end{lstlisting}
%     \label{fig:wrapper_code}
% \caption{}
% \end{figure}

% \begin{figure}
%     \centering
%     \caption{Caption}
%     \includegraphics[width=\linewidth]{figure/classifier_comparisons/SAG_predicition_s.png}
%     \caption{Number of significant digits (General view).}
%     \label{fig:classifier_comparisons_general}
% \end{figure}

% \begin{figure}
%     \centering
%     \caption{Caption}
%     \includegraphics[width=\linewidth]{figure/classifier_comparisons/zoom_SAG_fit_coef_s.png}
%     \caption{Number of significant digits (General view).}
%     \label{fig:classifier_comparisons_general}
% \end{figure}

% \subsubsection{Adaboost}

% \tristan{For all the paragraphs below, add a sentence to describe the results as you did for scipy. If you have nothing more to say than what is reported in Table 4, then remove the paragraphs describing the algorithms and just refer to the sklearn doc linked from the table. If you have interesting things to discuss only for some examples, mention in the section intro that the text focuses on the most interesting stability patterns observed across all experiments.}

% \paragraph{Adaboost}
% This test compares a decision tree boosted with 300 estimators using the AdaBoost.R2~\cite{drucker1997improving} algorithm with
% a single decision tree regressor on a 1D sinusoidal dataset with a small amount of Gaussian noise and a depth fixed at 4. 
% The example shows that as the number of boosts is increased the regressor can fit more detail.

% RR Good 48 \\
% MCA: arange issue /  bootstrap\_idx = random\_state.choice(a=np.arange(\_num\_samples(X)), size=\_num\_samples(X), replace=True, p=sample\_weight) / len(a) != len(p)

% \paragraph{Online classifier comparison}
% This example compares how the following online solvers perform on the hand-written digits dataset:
% Stochastic Gradient Descent (SGD) with maximum iterations fixed to 100, 
% Averaged Stochastic Gradient Descent (ASGD),
% Perceptron, Passive-Aggressive I with the hinge loss function, tolerance 
% and regularization fixed at $10^{-4}$ and $1$,
% Passive-Aggressive II with the squared hinge loss function, tolerance 
% and regularization fixed at $10^{-4}$ and $1$, and Stochastic Average Gradient (SAG)
% with the tolerance and regularization fixed at $10^{-4}$ and $5.10^{-8}$.

% RR: stable 48 bits. \\
% MCA: Index issue
\begin{figure}
    \centering
    \includegraphics[width=0.5\linewidth]{figure/BRR/call_path.pdf}
    \caption{Call paths of the Bayesian Ridge Regression (BRR) and
        Ordinary Least Square (OLS) function. Color represents the calling library:
        orange for scikit-learn, blue for SciPy and yellow for LAPACK.
        Dashed color represents the original path in red and the alternative one in green.
        Figure~\ref{fig:brr_svd_sig} shows the original method  leads to numerical instabilities while the alternative one converges.
    }
    \label{fig:call_path_brr}
\end{figure}

\paragraph{Bayesian Ridge Regression}

This example compares a Bayesian Ridge Regression (BRR) to the Ordinary Least
Squares (OLS) estimator on a synthetic dataset and for one-dimensional
regression using polynomial feature expansion. Although the example does not
raise a runtime error, the regression coefficients computed from the fitting are
non-significant. \pytracer traces reveal that the SVD solver used is the root
cause for this error. Figure~\ref{fig:call_path_brr} shows the two call paths
for BRR and OLS. The LAPACK library has two main methods for computing the SVD:
\texttt{gesdd} and \texttt{gesvd}. The former uses a Divide \& Conquer approach,
while the latter uses a QR decomposition. While both methods are expected to
have the same accuracy~\cite{nakatsukasa2013stable},
Figure~\ref{fig:brr_svd_sig} shows that \texttt{gesdd} is totally imprecise,
with an average number of significant digits of 0 for BRR and even
\texttt{Not-a-Number} (\texttt{NaN}) values for OLS. On the same figure, we can
see that using the \texttt{gesvd} (\remove{dgelsd} \add{dgelss}) method instead of the \texttt{gesdd}
(\remove{dgelss} \add{dgelsd}) one, the SVD converges for RR and Full MCA, with a number of
significant bits of 44 on average. This observation supports the on-going
discussion among LAPACK developers on the instability of \texttt{gesdd} (see
\href{https://github.com/Reference-LAPACK/lapack/issues/316}{github.com/Reference-LAPACK/lapack/issues/316}).

% \tristan{t'es sur que c'est pas l'inverse? Sur la fig 9 on voit que dgelsd n'est pas precise du tout, ca suggere qu'elle utiliserait gesdd}. 
% \tristan{ce paragraph est difficile a lire (mais neanmoins super interessant !). tu devrais le restructurer en expliquant (1) quelle methode utilise quel solver de quelle librairie, (2) la performance observee sur la fig 9 (tu pourais d'ailleurs ajouter le nom du solver (gesdd ou gesvd) entre parenthese dans la legende pour rendre l'explication plus claire, (3) le fait que ca confirme la discussion sur github et autres commentaires.}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figure/BRR/BRR_coefs_sig.pdf}
    \caption{Precision of Bayesian Ridge Regression coefficients in RR mode.
        \texttt{BRR\_*} show the \texttt{BayesianRidge} results
        using the \texttt{gesdd} and \texttt{gesvd} methods. 
        \texttt{OLS\_*} show the \texttt{LinearRegression} results
        with the \texttt{dgelsd} and \texttt{dgelss} methods.
        The Divide \& Conquer method 
        (\texttt{gesdd} and  \texttt{dgelsd})
        has low precision
        for \textit{BRR} and results in \texttt{NaN} values (star points) for \textit{OLS}. By switching to the \texttt{gesvd} method, 
        results have a precision of 42 bits on average.
    }
    \label{fig:brr_svd_sig}
\end{figure}

% the coefficient weights are slightly shifted toward zeros, which stabilises them.
% As the prior on the weights is a Gaussian prior, the histogram of the estimated weights is Gaussian.
% The estimation of the model is done by iteratively maximizing the marginal log-likelihood of the observations.
% We also plot predictions and uncertainties for Bayesian Ridge Regression for one dimensional regression using polynomial feature expansion. Note the uncertainty starts going up on the right side of the plot. This is because these test samples are outside of the range of the training samples.

% RR: solution mean 15-24 / std 20-30 \\
% MCA: SVD did not converge

% \paragraph{K-Means Clustering}
% This test uses the K-means to cluster the IRIS dataset.
% The plots display firstly what a K-means algorithm would yield using three clusters. It is then shown what the effect of a bad initialization is on the classification process: By setting n\_{init} to only 1 (default is 10), the amount of times that the algorithm will be run with different centroid seeds is reduced. The next plot displays what using eight clusters would deliver and finally the ground truth.

% RR Very stable (labels with highest precision) \\
% MCA: Divergence

% \paragraph{Covariance estimation}
% This example shows covariance estimation with Mahalanobis distances on Gaussian distributed data.

% For Gaussian distributed data, the distance of an observation xi to the mode of the distribution can be computed using its Mahalanobis distance:

% d(μ,Σ)(xi)2=(xi−μ)TΣ−1(xi−μ)
% where μ and Σ are the location and the covariance of the underlying Gaussian distributions.

% In practice, μ and Σ are replaced by some estimates. The standard covariance maximum likelihood estimate (MLE) is very sensitive to the presence of outliers in the data set and therefore, the downstream Mahalanobis distances also are. It would be better to use a robust estimator of covariance to guarantee that the estimation is resistant to “erroneous” observations in the dataset and that the calculated Mahalanobis distances accurately reflect the true organization of the observations.

% The Minimum Covariance Determinant estimator (MCD) is a robust, high-breakdown point (i.e. it can be used to estimate the covariance matrix of highly contaminated datasets, up to

% nsamples−nfeatures−12 outliers) estimator of covariance. The idea behind the MCD is to find

% nsamples+nfeatures+12 observations whose empirical covariance has the smallest determinant, yielding a “pure” subset of observations from which to compute standards estimates of location and covariance. The MCD was introduced by P.J.Rousseuw in 1.

% This example illustrates how the Mahalanobis distances are affected by outlying data. Observations drawn from a contaminating distribution are not distinguishable from the observations coming from the real, Gaussian distribution when using standard covariance MLE based Mahalanobis distances. Using MCD-based Mahalanobis distances, the two populations become distinguishable. Associated applications include outlier detection, observation ranking and clustering.

% RR: Traces diverge but precision seems stable (~48bits) \\
% Main computation is in a loop that might be take extra steps to ends. \\
% MCA: arange issue

% \paragraph{Decision Tree Regression}
% This test uses a decision tree to fit a sine curve with added noise.

% A 1D regression with decision tree.
% The decision trees is used to fit a sine curve with addition noisy observation. As a result, it learns local linear regressions approximating the sine curve.
% We can see that if the maximum depth of the tree (controlled by the max\_depth parameter) is set too high, the decision trees learn too fine details of the training data and learn from the noise, i.e. they overfit.

% RR Fit ok (50bits) , missing predict \\
% MCA Fit ok (50bits) , missing predict

% \paragraph{Digits Classification}
% This test uses a Support Vector Classifier to recognize images of hand-written
% digits. The digits dataset consists of 8x8 pixel images of digits.

% RR Very stable (highest precision) \\
% MCA: Very stable (highest precision)

\paragraph{Face Recognition}

The face recognition example uses eigenfaces (eigenvectors) and Support Vector
Machines (SVMs) to classify faces from the Labeled Faces in the Wild dataset.
This example uses a Principal Component Analysis (PCA) to extract the eigenfaces
and then trains an SVM classifier to predict the name associated with a face.
This example is the only one to raise an error with RR due to the
non-convergence of the Singular Value Decomposition (SVD) performed in the PCA.

In this example, the PCA uses the Randomized SVD~\cite{halko2011finding} method
that first computes a low-rank approximation of the input matrix using a
stochastic algorithm and then uses this approximation as input for the SVD. The
example raises an error during the computation of the SVD. The
\texttt{randomized\_svd} scikit-learn function uses the Divide \& Conquer SVD
method \texttt{gesdd} mentioned previously. Replacing \texttt{gesdd} for
\texttt{gesvd} solves the SVD convergence issue.
Figure~\ref{fig:randomized_svd_S} shows the precision of the eigenvalues
computed with gesvd \add{in single precision}, which varies between 15 and 20 significant bits out of the
24 available in single precision. The classifier metrics (precision, recall,
F1-score, and support) are identical to the IEEE execution for all MCA samples.

In conclusion, \pytracer and the MCA model were helpful to detect the
instability of the \texttt{gesdd} method mentioned in several GitHub issues,
including the LAPACK repository. Furthermore, it is worth mentioning that a Pull
Request has recently been submitted to let the user select the SVD method and
avoid this kind of issue. Thus, \pytracer can detect instabilities that affect
fundamental scientific computing components such as the LAPACK library.
% The Pytracer findings can therefore be extended to many application domains.
% % \tristan{conclusion needs to be discussed. Note to self: generaliser un peu la conclusion: pytracer permet de se rendre compte de problemes de stabilite dont plusieurs librairies sont dependantes.}

\begin{figure}
    \centering
    \begin{subfigure}{0.3\linewidth}
        \includegraphics[width=\linewidth]{figure/face_recognition/randomized_svd_ret_U_sig.pdf}
        \caption{Left matrix $U$ of the SVD decomposition}
        \label{fig:randomized_svd_U}
    \end{subfigure}
    \begin{subfigure}{0.3\linewidth}
        \includegraphics[width=\linewidth]{figure/face_recognition/svd_ret_S_sig.pdf}
        \caption{Singular values $\Sigma$ of the SVD decomposition}
        \label{fig:randomized_svd_S}
    \end{subfigure}
    \begin{subfigure}{0.3\linewidth}
        \includegraphics[width=\linewidth]{figure/face_recognition/randomized_svd_ret_V_sig.pdf}
        \caption{Right matrix $V$ of the SVD decomposition}
        \label{fig:randomized_svd_V}
    \end{subfigure}
    \caption{Precision of the \texttt{randomized\_svd} function in the
        \texttt{face\_recognition} example using the \texttt{gesvd} SVD method and RR.
        Figures~\ref{fig:randomized_svd_U},~\ref{fig:randomized_svd_S}, and~\ref{fig:randomized_svd_V}
        show respectively the values $U$, $\Sigma$ and $V$.
        Numerical precision is lower for smaller singular values and this loss of precision translates to singular vectors. It would be interesting to investigate the use of numerical precision for dimensionality reduction.
    }
    \label{fig:face_recognition_svd}
\end{figure}

% RR: Crash due to FP32 use. Does FP64 converge? No. Does sgsvd converge? Yes

% MCA: arange issue

% File "/usr/local/lib/python3.8/site-packages/sklearn/datasets/\_lfw.py", line 169, in \_load\_imgs
%     faces[i, ...] = face
% ValueError: could not broadcast input array from shape (49,37) into shape (50,37)

% \paragraph{L1 Penalty}
% This test compares the percentage of zero coefficients for logistic regression
% by varying the penalty (L1, L2 and Elastic-Net) and regularization (1, 0.1, 0.01) parameters. 
% A large regularization value gives the model more freedom than a small value, which puts more constraints on the model.

% of L1, L2 and Elastic-Net penalty.

% Comparison of the sparsity (percentage of zero coefficients) of solutions when L1, L2 and Elastic-Net penalty are used for different values of C. We can see that large values of C give more freedom to the model. Conversely, smaller values of C constrain the model more. In the L1 penalty case, this leads to sparser solutions. As expected, the Elastic-Net penalty sparsity is between that of L1 and L2.
% We classify 8x8 images of digits into two classes: 0-4 against 5-9. The visualization shows coefficients of the models for varying C.

% RR: ok \\
% MCA: stable for L1 and L2, not for Len but coef have less zeroes.

% \paragraph{Lasso and Elastic-Net}
% This test compares the Lasso and Elastic-Net regression models
% on a corrupted signal.

% Estimates Lasso and Elastic-Net regression models on a manually generated sparse signal corrupted with an additive noise. Estimated coefficients are compared with the ground-truth.

% RR: Good (48 bits on coefs) \\
% MCA: Good like RR if we exclude one trace that fails due to arange issue

\paragraph{Separating hyperplane}

This example uses a linear Support Vector Machine (SVM) classifier trained using
the Stochastic Gradient Descent (SGD) method to find the maximum separating
hyperplane between sets of 2D points. The example trains the SVM with 50 random
points separated in 2 clusters with hyperparameter $\alpha=0.01$ and using the
hinge loss function. Once the model is trained, the example generates a
$[-1,5]\times[-1,5]$ Cartesian grid discretized with 100 points and predicts the
class label for each point with integer coordinates.
Figure~\ref{fig:hyperplan_sig_zoom} shows the number of significant bits
obtained with RR for each point of the grid. 
We can see that yellow regions 
close to the separating hyperplane are highly unstable with a number of
significant bits below 0, meaning that the classifier is unsure about the
predicted class. 
%\tristan{Describe and comment on the other 2 tests, why are they more stable? Is it because the training set has more points?}

% \begin{figure}
%     \centering
%     \includegraphics[width=\linewidth]{figure/separating_hyperplan/hyperplan.png}
%     \caption{Ratio $mean/std$ of Separating hyperplane test result within RR mode.
%     Circle and square points represent respectively the train and test dataset.
%     The color represents the predicted class. We can see that 
%     instability is high around the hyperplane which is coherent with 
%     the fact the classifier has difficulty to decide in which side of the hyperplane the 
%     point is.}
%     \label{fig:separating_hyperplan}
% \end{figure}

\begin{figure}
    \centering
    % \begin{subfigure}{.3\linewidth}
    % \includegraphics[width=\linewidth]{figure/separating_hyperplan/hyperplan.png}
    % \caption{}
    % \label{fig:hyperplan_sig}
    % \end{subfigure}
    \begin{subfigure}{.45\linewidth}
        \includegraphics[width=\linewidth]{figure/separating_hyperplan/hyperplane_zoom.pdf}
        \caption{Separating hyperplane}
        \label{fig:hyperplan_sig_zoom}
    \end{subfigure} \\
    \begin{subfigure}{.45\linewidth}
        \includegraphics[width=\linewidth]{figure/SVM/non_weighted.pdf}
        \caption{SVM (non weighted)}
        \label{fig:SVM_nw_sig}
    \end{subfigure}
    \begin{subfigure}{.45\linewidth}
        \includegraphics[width=\linewidth]{figure/SVM/weighted.pdf}
        \caption{SVM (weighted)}
        \label{fig:SVM_w_sig}
    \end{subfigure} \\
    \begin{subfigure}{.45\linewidth}
        \includegraphics[width=\linewidth]{figure/Weighted/non_weighted.pdf}
        \caption{SGD weighted (non-weighted)}
        \label{fig:weighted_nw_sig}
    \end{subfigure}
    \begin{subfigure}{.45\linewidth}
        \includegraphics[width=\linewidth]{figure/Weighted/weighted.pdf}
        \caption{SGD weighted (weighted)}
        \label{fig:weighted_w_sig}
    \end{subfigure}
    \caption{
        Three examples of separating hyperplane prediction. All the figures show the number of significant bits of the prediction in RR mode. The separating hyperplane corresponds to a lower-precsion region in each figure. Using a large discretization step amplifies
        this instability.
        % respectively the results of examples    \textit{SVM} and~\textit{Weighted samples}. Figures~\ref{fig:SVM_nw_sig} and~\ref{fig:weighted_nw_sig} show the unweighted classification while Figures~\ref{fig:SVM_w_sig} and~\ref{fig:weighted_nw_sig} the classification taking into account the weights. We can see that taking into account the weight classes allows a better separation. 
        % \tristan{add labels to subfigures} \tristan{don't use this yellow for class 1, take a color that isn't in the color map. Add a legend to show the class colors.}
    }
    \label{fig:separating_hyperplan}
\end{figure}

% \paragraph{MNIST classification}
% This test fits a logistic regression with the SAGA solver with the L1 penalty on the MNIST digits classification task subset.

% Here we fit a multinomial logistic regression with L1 penalty on a subset of the MNIST digits classification task. We use the SAGA algorithm for this purpose: this a solver that is fast when the number of samples is significantly larger than the number of features and is able to finely optimize non-smooth objective functions which is the case with the l1-penalty. Test accuracy reaches > 0.8, while weight vectors remains sparse and therefore more easily interpretable.
% Note that this accuracy of this l1-penalized linear model is significantly below what can be reached by an l2-penalized linear model or a non-linear multi-layer perceptron model on this dataset.

% RR: Very stable \\
% MCA: Parse by excluding a trace that failed due to an arange issue

% \paragraph{Multitask Lasso}
% This test compares the Lasso and MultiTaskLasso regression models
% on a corrupted signal.

% The multi-task lasso allows to fit multiple regression problems jointly enforcing the selected features to be the same across tasks. This example simulates sequential measurements, each task is a time instant, and the relevant features vary in amplitude over time while being the same. The multi-task lasso imposes that features that are selected at one time point are select for all time point. This makes feature selection by the Lasso more stable.

% RR: Stable (44~53 bits for coef) \\
% MCA: Stable like RR if we exclude a trace that failed due index issue

% \paragraph{Orthogonal matching}
% This test uses the Orthogonal Matching Pursuit algorithm
% for recovering a sparse signal from a noisy measurement encoded with a dictionary.

% RR: Very stable \\
% MCA: Very stable

% \paragraph{PCA decomposition}
% This test makes a PCA decomposition on a synthetic dataset.
% These figures aid in illustrating how a point cloud can be very flat in one direction–which is where PCA comes in to choose a direction that is not flat.

% rr: stable (48 bits on components) \\
% mca: stable (48 bits on components)

% \paragraph{Robust Linear Regression}
% This test compares the linear and  RANSAC regression methods on noisy data.
% In this example we see how to robustly fit a linear model to faulty data using the RANSAC algorithm.

% RR: Very stable \\
% MCA: Very stable if we exclude two traces that failed due to arange/index issues

% \paragraph{Segmentation Toy}
% This test uses spectral clustering to separate connected circles in an image.

% In this example, an image with connected circles is generated and spectral clustering is used to separate the circles.

% In these settings, the Spectral clustering approach solves the problem know as ‘normalized graph cuts’: the image is seen as a graph of connected voxels, and the spectral clustering algorithm amounts to choosing graph cuts defining regions while minimizing the ratio of the gradient along the cut, and the volume of the region.

% As the algorithm tries to balance the volume (ie balance the region sizes), if we take circles with different sizes, the segmentation fails.

% In addition, as there is no useful information in the intensity of the image, or its gradient, we choose to perform the spectral clustering on a graph that is only weakly informed by the gradient. This is close to performing a Voronoi partition of the graph.

% In addition, we use the mask of the objects to restrict the graph to the outline of the objects. In this example, we are interested in separating the objects one from the other, and not from the background.

% RR Stable \\
% MCA: Stable if we exclude three traces that failed due to index issue

\paragraph{SVM}
This example uses a plain Support Vector
Classification~\cite{Platt99probabilisticoutputs} (SVC) to find the best
hyperplane separating an unbalanced dataset with two classes of 1000 and 100
points. This example tests two SVC configurations using a linear kernel with and
without a class weight parameter. The SVC predicts the class label for each
coordinate over a $[-5,5] \times [-5,5]$ grid discretized with 900 points.
Figures~\ref{fig:SVM_nw_sig} and~\ref{fig:SVM_w_sig} show the result within RR
of the classification with and without taking into account the unbalancing. We
can see that the number of significant bits is lower on the separating line and
that taking into account class imbalance allows for a better separation.
% Find the optimal separating hyperplane using an SVC for classes that are unbalanced.

% We first find the separating plane with a plain SVC and then plot (dashed) the separating hyperplane with automatically correction for unbalanced classes.

% RR: stable (48 bits) \\
% MCA: stable (48 bits)

% \paragraph{Tomography}
% This test compares the LASSO and Ridge regression models to 
% reconstruct an image from  projections for a dataset
% acquired from computed tomography.

% This test shows the reconstruction of an image from a set of parallel projections, acquired along different angles. Such a dataset is acquired in computed tomography (CT).

% Without any prior information on the sample, the number of projections required to reconstruct the image is of the order of the linear size l of the image (in pixels). For simplicity we consider here a sparse image, where only pixels on the boundary of objects have a non-zero value. Such data could correspond for example to a cellular material. Note however that most images are sparse in a different basis, such as the Haar wavelets. Only l/7 projections are acquired, therefore it is necessary to use prior information available on the sample (its sparsity): this is an example of compressive sensing.

% The tomography projection operation is a linear transformation. In addition to the data-fidelity term corresponding to a linear regression, we penalize the L1 norm of the image to account for its sparsity. The resulting optimization problem is called the Lasso. We use the class Lasso, that uses the coordinate descent algorithm. Importantly, this implementation is more computationally efficient on a sparse matrix, than the projection operator used here.

% The reconstruction with L1 penalization gives a result with zero error (all pixels are successfully labeled with 0 or 1), even if noise was added to the projections. In comparison, an L2 penalization (Ridge) produces a large number of labeling errors for the pixels. Important artifacts are observed on the reconstructed image, contrary to the L1 penalization. Note in particular the circular artifact separating the pixels in the corners, that have contributed to fewer projections than the central disk.

% RR Divergence \\
% MCA Index issue

\paragraph{SGD weighted samples}
This example uses an SGD to separate weighted points. It divides the training
set into two classes of 10 points with a bigger weight for the second class. The
SGD is trained with hyperparameters $\alpha=0.01$ and a maximum number of
iterations to 100. The prediction is made on a $[-4,5]\times[-4,5]$ grid
discretized with 250\add{,}000 points. Figures~\ref{fig:weighted_nw_sig}
and~\ref{fig:weighted_w_sig} show the number of significant bits of prediction
within RR along with dataset points in blue and \remove{violet} \add{purple}. We can see that the
separating plane corresponds to a region of lower precision, similar to the
previous examples. 
% Finally, taking into account the points' weight allows for a better separation \tristan{is this related to numerical stability? If not, you can ommit the comment.}.

% RR Stable (30~48 bits on coef) \\
% MCA Unstable (0 bits on coef)

% \subsection{PyAFQ}

% PyAFQ is a Python package for Automated Fiber Quantification.
% It aims to find delineation of the major fiber tracts in individual human brains and quantify the tissue properties within the tracts. PyAFQ \pytracer tracing produces traces of 1TB with an average slowdown $\times 6$ compared to the actual execution. Moreover, \pytracer was able to visualize the aggregated traces, demonstrating its scalability fluently.

\subsection{Overhead evaluation}

\pytracer's tracing overhead consists of (1) module instrumentation time
(initialization overhead) and (2) data write time during the execution (runtime
overhead). The initialization overhead is proportional to the number of
functions and classes traced. For example, we measured an average initialization
time of 8 seconds during our experiments with  NumPy, SciPy, and Scikit-learn.
This overhead can be reduced by excluding intensively used functions such as
\texttt{numpy.array} from the instrumentation.
Figure~\ref{fig:performance_tracing} summarizes \pytracer's tracing overhead for
SciPy and scikit-learn examples. \textit{Slowdown} represents the slowdown of
\pytracer without fuzzy measured as the ratio between the actual execution time
(without \pytracer) and the execution time using \pytracer. The \textit{Slowdown
    RR} and \textit{Slowdown MCA} show the slowdown within the fuzzy environment.
Without fuzzy, \pytracer introduces an average slowdown of $17\times$ high for
the examples running less than 8 seconds because \pytracer spends most of its
time in the initialization step. Since this step is a constant factor of the
number of functions instrumented, we also measured the instrumentation cost
itself (\textit{Slowdown (Instrumentation)}) calculated by subtracting the
initialization time to \textit{Slowdown}. We then obtain an average slowdown of
$3.5\times$. Finally, the \remove{fuzzy instrumentation} \add{using MCA operations (throuhg fuzzy) instead of the native ones} increases the average slowdown
to $37\times$ for RR and $43\times$ for Full MCA.

% \pytracer's tracing overhead consists of (1) module instrumentation time (initialization overhead), and (2) data write time during the execution (runtime overhead). The initialization overhead is proportional to the number of functions and classes traced. However, the initialization overhead is amortized \tristan{what do you mean by amortized? I don't get this sentence} when traced applications are above 8 seconds,  proportional to 
% the number of functions and classes traced. Figure~\ref{fig:performance_tracing} summaries \pytracer's tracing overhead for SciPy and scikit-learn examples. We can see that \pytracer introduces a slowdown of reduced toif we subtract the initialization step 
% \tristan{not sure where to get these numbers from the table, to be clarified. Are these average values? Why are there two values?}. 
% Fuzzy instrumentation increases the slowdown to for RR and   for Full MCA.
% \tristan{Figure 13 is not referred in the text. Choose between the figure and the table, no need to put both.}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figure/performance.pdf}
    \caption{
        \pytracer tracing overheads. \textit{Slowdown} represents the slowdown of \pytracer without fuzzy measured as the ratio between the actual execution time (without \pytracer) and the execution time using \pytracer. 
        \textit{Slowdown (Instrumentation)} show the actual instrumentation cost computed by subtracting the initialization step.
        The \textit{Slowdown (RR)} and \textit{Slowdown (MCA)} are the slowdown of \pytracer 
        using the fuzzy environment with RR and Full MCA modes.
        The first \remove{first} \add{last} four examples measure the initialization steps depending on the modules instrumented. 
        % \tristan{there's too much text in this caption, move some of it to the text, the caption should just explain the difference between the 4 slowdowns. From the text I'm not sure what is the difference between slowdown and slowdown(RR)/slowdown(MCA).}
    }
    \label{fig:performance_tracing}
\end{figure}

% \begin{figure}
%     \centering
%     \includegraphics[width=\linewidth]{figure/parsing_time.pdf}
%     \caption{\pytracer time to aggregate traces.}
%     \label{fig:performance_parsing_time}
% \end{figure}

% \begin{figure}
%     \centering
%     \includegraphics[width=\linewidth]{figure/parsing.pdf}
%     \caption{\pytracer trace and aggregation files size.
%     The figure presents the size measured for the trace and the aggregation size files.}
%     \label{fig:performance_parsing_size}
% \end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figure/performance_parsing.pdf}
    \caption{\pytracer aggregation performance.
        % \tristan{change xlabel to "Trace size (bytes) and ylabel to "Aggregation time (seconds)}
    }
    \label{fig:performance_parsing}
\end{figure}

% \begin{table}
% \centering
% \begin{subfigure}[t]{.75\linewidth}
%     \centering
%     \begin{tabular}{|l|c|c|c|c|}
%     \hline 
%     Application & Original & Pytracer only & RR & MCA \\
%     \hline 
%     Loading & 0.456 & 8.152 & - & - \\
%     \hline
%     discrete\_cosine\_transform & 0.280 & 8.176 & 10.203 & - \\ 
%     fft\_1D & 0.650 & 9.075 & 11.345 & - \\
%     fft\_2D & 0.311 & 9.500 & 11.929  & 16.433   \\
%     \hline
%     interpolation\_1D & 0.377 & 8.803  & 11.954  & -   \\
%     multivariate\_data &0.626 & 8.648 &  12.711  & - \\
%     bspline & 0.109  & 9.050 & 13.396  & 16.830   \\
%     spline\_1D & 0.111 & 8.618 & 10.960   & - \\
%     spline\_2D & 0.655 & 8.649 & 19.424  & - \\
%     \hline
%     broyden\_fletcher\_goldfarb\_shanno &0.475 & 8.292 & 11.033 & 13.685 \\
%     global\_optimization &1.257 & 55.916  & 93.322  & - \\
%     slsqp & 0.469 & 8.192 & 10.795 & 12.439  \\
%     least\_square\_minimization & 0.505 & 10.151  & 13.888 & -  \\
%     nelder\_mead\_simplex & 0.471 & 8.418  & 11.382  & 12.792  \\
%     newton\_conjugate\_gradient & 0.487 & 8.548  & 11.956  & 14.281 \\
%     root\_findings & 0.462 & 8.198  & 10.787   & - \\
%     root\_finding\_large & 0.801 & 11.525 & 41.493   &  82.701  \\
%     root\_finding\_large\_preconditioned & 0.702 & 10.580 & 33.199   & 57.579  \\
%     trust\_region\_exact & 0.482 & 8.170 & 10.400  & 12.524 \\
%     trust\_region\_truncated\_generalized\_lanczos & 0.459 & 8.457 & 11.216 & 13.245  \\
%     trust\_region\_newton\_conjugate\_gradient & 0.479 & 8.538 & 11.179 & 13.488 \\
%     \hline
%     Adaboost &1.388 &14.553 &  20.999 & - \\
%     Bayesian Ridge Regression & 1.086 & 10.344 & 90.299 & -  \\
%     Online classifier comparison & 15.964 & 72.691 & 2104.114 & -  \\
%     K-Means Clustering & 1.621 & 17.614  & 34.005 & 52.902   \\
%     Covariance estimation & 1.357 & 11.755 & 23.964  & -   \\
%     Decision Tree Regression & 1.185 & 11.194 & 15.760 & 15.918  \\
%     Digits Classification & 1.672 & 12.725 & 40.634 & - \\
%     Face Recognition & 32.159  & 74.356 & - & - \\
%     L1 Penalty & 1.568 & 12.591 & 65.547 & 128.220 \\
%     Lasso and elastic net & 1.057 & 9.903 & 17.078 & -  \\
%     Separating hyperplan & 1.376 & 11.532  & 22.333 & 25.036 \\
%     MNIST classification & 19.650 & 37.838 & 402.389  & - \\
%     Multitask Lasso & 1.090 &10.196 & 15.025  & - \\
%     Orthogonal matching & 1.161 & 12.129 & 21.779   & 31.135 \\
%     PCA decomposition & 1.239 & 10.642 & 15.520 & 17.952    \\
%     Robust Linear Regression & 1.370 & 11.144  & 14.280  & - \\
%     Segmentation toy &1.679 & 22.385 & 87.811  & - \\
%     SVM & 1.403 & 11.737 & 23.203 & 24.126   \\
%     Tomography & 11.108 & 40.177  & 1376.134 & -  \\
%     Weighted samples & 1.358 & 11.768 & 21.255 & 20.541   \\
%     \hline
%     \end{tabular}
% \end{subfigure}
%     \caption{\pytracer tracing overhead in seconds. \tristan{what is the unit in this table?}}
%     \label{tab:pytracer_overhead}
% \end{table}

% \begin{table}[]
%     \centering
%     \begin{tabular}{c|c|c|c|c}
%          Pytracer & Verificarlo Instrumentation & Backend & Time (hours:minutes:seconds) & Overhead \\
%          \hline 
%          No & No & - & 00:30:51 & 1 \\
%          No & Yes & IEEE & 01:32:56 & 3 \\
%          No & Yes & RR & 16:52:08 & 32 \\
%          Yes & No & - & 03:11:07 & 6  \\
%          Yes & Yes & IEEE & 11:14:33 & 22 \\
%          Yes & Yes & RR & 25:05:54 & 48 \\
%     \end{tabular}
%     \caption{Overheads of \pytracer on PyAFQ experiment with and without fuzzy.}
%     \label{tab:pytracer_overhead}
% \end{table}

% \begin{table}[]
%     \centering
%     \begin{tabular}{c|c|c|c|c}
%          Pytracer & Verificarlo Instrumentation & Backend & Time (hours:minutes:seconds) & Overhead \\
%          \hline 
%          No & No & - & 15 & 1 \\
%          No & Yes & IEEE & 49 & 3 \\
%          No & Yes & RR &  & 441 & 29 \\
%          Yes & No & - & 03:11:07 & 6  \\
%          Yes & Yes & IEEE & 11:14:33 & 22 \\
%          Yes & Yes & RR & 25:05:54 & 48 \\
%     \end{tabular}
%     \caption{Caption}
%     \label{tab:pytracer_overhead}
% \end{table}

Figure~\ref{fig:performance_parsing} shows the postprocessing time and trace
sizes produced by \pytracer. The linear correlation between the log of the size
and the log of the time shows a significant correlation (p-value $\leq 10^{-8}$
for RR and p-value $\leq 10^{-3}$ for MCA).
% \tristan{Mention linear correlation with p-value here.}

In conclusion, while \pytracer's overhead is substantial, it remains tractable
for real-life examples.

\section{Discussion}

% Numpy   2509 (2403 not tests) (1848) functions, 1196 (1154) (817) classes (pure Python)
% Scipy   4031 (2187) (506) functions, 1532 (472) (109) classes (pure Python)
% Sklearn 4807 (1010) (81) functions, 546 (417) (41) classes (pure Python)

% \pytracer is a tool for visualizing the numerical quality of function inputs and outputs along the execution of a Python program. It generates numerical traces and post-processes them to measure numerical discrepancies between executions. \pytracer does not require any particular noise model to estimate uncertainties. We tested it with MCA through the fuzzy ecosystem on two broadly used Python libraries.

Through our analysis of SciPy and scikit-learn, we demonstrated \pytracer's
scalability, automaticity, and applicability to a wide range of Python program.
The relatively low instrumentation overhead of 3.5$\times$ on average enables
the analysis of real-scale applications. \pytracer's instrumentation is fully
automatic and does not require manual intervention from the user. Furthermore,
\pytracer works on any Python libraries, with any data format supported by
pickle and any numerical noise model.

\pytracer is the first tool to enable a systematic numerical analysis of Python
codes. Existing tools focus on traditional languages of the HPC community, such
as C, C++, and FORTRAN. However, Python's high-level abstraction allows for
quick application implementations while maintaining high-performance thanks to
specialized libraries such as NumPy. Therefore, \pytracer meets the numerical
analysis needs of an ever-increasing number of applications.

\pytracer allows users to pinpoint the root causes of numerical instabilities by
working at the function level. Numerical stability evaluations are complex since
calculations are strongly interdependent. Function-level instrumentation limits
the size of the traces since only inputs and outputs of a function are dumped.
Moreover, it decreases instrumentation overhead and helps the user identify
instabilities quickly. Working at the function level is thus a suitable
compromise between analysis conducted at the floating-point operation level and
user debug prints. Indeed,  fine-grained analyses are costly and hard to
interpret as they overwhelm users with information. In contrast, looking at
intermediate results through print statements is lightweight but not reliable
since developers may not systematically instrument unstable functions. Finally,
the SciPy environment wraps many computation kernels written in C, such as BLAS
or LAPACK. By working at the function level, \pytracer allows the user to
evaluate the numerical precision of such kernels and their impact on the code
without getting deep into technical implementation aspects.

Integrating \pytracer at the early stages of the development cycle would improve
the overall numerical quality of the developed code. \pytracer's automaticity
facilitates its usage and enables its integration in unit testing. For example,
\pytracer could be used in regression tests to ensure that updates do not
degrade the numerical quality of a calculation.

\pytracer can also be useful for selecting hyperparameters, as is the case in
machine learning. Indeed, it is common for data scientists to explore a range of
hyperparameters and select the ones giving the best scores. With \pytracer, the
user has a practical way to quantify numerical quality and understand why
hyperparameters produce inconsistent results instead of tweaking hyperparameters
to make solvers converge. For instance, the Face Recognition example
(Figure~\ref{fig:face_recognition_svd}) shows how numerical precision could
inform the selection of the number of components in dimensionality reduction.

The visualization provides a global view of the code and interactions between
computations. In addition, the Plotly dashboard helps navigate through the code
easily, without the need for ad-hoc scripts to visualize data. Finally, the
dashboard can be extended to support other data formats to visualize. At this
stage, \pytracer currently supports n-dimensional NumPy arrays as well as native
Python data.

As we showed in Section~\ref{sec:impact_mca_modes}, using Full MCA mode leads to
runtime errors that have a considerable impact on the proper functioning of the
execution, which is unfortunate since these errors do not reflect actual
numerical errors but are related to perturbations that should not occur.
Conversely, RR mode is far more conservative since it preserves exact operations
and is easier to use even though it does not simulate all perturbations.
Therefore, from our experiments, we recommend the use of RR over Full MCA. More
research is required to address the issues encountered with Full MCA.

Finally, the wide range of results precision obtained from our experiments
demonstrate that looking at the numerical variability is essential and that such
analyses would benefit from being conducted systematically. We showed that
well-known scientific libraries are prone to numerical error even though they
have been well tested for decades. The omnipresence of these libraries in
scientific codes should thus impel users to integrate numerical analysis in
their development workflow in addition to testing their implementation.

\section{Related work}

Several tools have been developed to assess numerical quality, and can be
divided into static and dynamic approaches. Static approaches typically analyze
source code, while dynamic approaches trace floating-point computations
throughout executions. A detailed review of these approaches is presented
in~\cite{cherubin2020tools}. 
% \pytracer adopts the dynamic approach due to its scalability to large code
% bases. Dynamic approaches typically trace floating-point computations, detect
% instabilities in traces, and visualize summary statistics about instabilities.
% This section reviews existing approaches for each of these steps.
% \label{sec:soa} \paragraph{Numerical tracing techniques}
Numerous tools exists to trace floating-point computations for C, C++, or
FORTRAN programs due to the prevalence of these languages in High-Performance
Computing (HPC). The main tracing techniques are \textit{source-to-source},
\textit{compiler-based transformations}, and \textit{dynamic binary
    instrumentation}.

% \gkmod{}{Note: this paragraph just reads like a list with no real
% interpretation/commentary. I think it would be better to state the strengths
% of this technique, cite examples, and then state limitations/why it isn't the
% only technique used.}
The source-to-source technique requires a rewriting of the application to modify
floating-point types and provides a fine-grained control on the analyzed code.
CADNA~\cite{jezequel2008cadna} is a library for C, C++, and Fortran implementing
CESTAC~\cite{vignes1993stochastic} stochastic arithmetic.
Shaman~\cite{demeure_phd} is a C++ library that uses a first-order error model
to propagate numerical error. MCAlib~\cite{frechtling2015mcalib} is a C library
that implements the Monte Carlo Arithmetic (MCA)~\cite{parker1997monte} using
the MPFR~\cite{fousse2007mpfr} library. \add{ADAPT \mbox{~\cite{menon2018adapt}} uses
    Automatic Differentiation to compute a sensitivity profile for each variable of
    given program}. Finally, the work in~\cite{tang2016software} proposed a
source-to-source framework for executing targeted code in infinite and fixed
precision with and without stochastic arithmetic. The main drawback of the
source-to-source approach is its lacks of scalability since rewriting large
codes can be a tedious task.

% \gkmod{}{Note: similar to the above, this is just a list rather than a
% highlight of how the compiler-based approach can be more scalable than
% source-to-source because it requires less manual intervention into a codebase
% (e.g.), then have your list, and finally say why it is still limited/not a
% global fix to this problem.}

The compiler-based approach instruments floating-point expressions at compile
time and so allows the user to automatically instrument large codebases.
Verificarlo~\cite{verificarlo} is a compiler that supports different types of
arithmetic instrumentations, including MCA. The work in~\cite{bao2013fly}
modified the GNU Compiler Collection (GCC) to track local floating-point errors
across executions. pLiner~\cite{guo2020pliner} is a root-cause analyzer based on
the Clang compiler that detects floating-point operations responsible for result
variability using a source-to-source transformation at the Abstract Syntax Tree
(AST) level to rewrite parts of code with higher precision.
PFPSanitizer~\cite{chowdhary2020debugging,chowdhary2021parallel} \add{and
    NSan~\mbox{\cite{courbet2021nsan}} are compiler based approaches}
that use shadow execution to detect numerical issues using higher
precision.
FLiT~\cite{sawaya2017flit} is a framework to detect variability
induced by compilers and their optimizations. The work
in~\cite{wang2012development} proposes a numerical debugger based on
GDB~\cite{stallman1988debugging} for Discrete Stochastic Arithmetic (DSA) on
FPGA as an Eclipse plugging. Similarly, Cadtrace~\cite{jezequel2008cadna} and
Shaman propose a GDB-based tool to use the CADNA and Shaman libraries with GDB,
respectively. The main limitation of the compiler-based approach is that one
must have access to the source code.

Dynamic Binary Instrumentation (DBI) operates directly on executables, without
the need for recompilation or manual changes. Therefore, it is applicable to any
programming language. CraftHPC~\cite{lam2013dynamic} uses DBI to detect
catastrophic cancellations at runtime. Verrou~\cite{fevotte2016verrou} is a
Valgrind~\cite{nethercote2007valgrind} based tool that dynamically replaces
floating-point operations with their stochastic arithmetic counterparts.
FPDebug~\cite{benz2012dynamic} uses DBI to detect numerical inaccuracies by
using shadow computations with higher precision.
Herbgrin\remove{g}\add{d}~\cite{sanchez2017finding} is a Valgrind-based tool to detect numerical
instabilities. It uses a shadow memory to detect precision losses by comparison
with results obtained \remove{at} \add{in} higher precision. It is combined with symbolic
computation to backtrack the root of the error. We can note that all methods
based on a high-precision oracle suppose that computing with a large number of
bits is sufficient to obtain an accurate result, which is not always true (see
for example, the famous Muller's sequence~\cite{bajard1996introduction}).
Although versatile, the DBI looses high-level information useful to understand
the source-code logic.

% \gkmod{}{Note: I like this comparison here, but do think that the comments
% above about providing some more info when each is mentioned would still help
% this section.}
Being agnostic to the programming language is a serious advantage of the DBI
method compared to source-to-source and compiler-bases methods. However, working
at the binary level makes it difficult to access high-level information needed
to debug and understand the source-code logic. Conversely, the source-to-source
approach provides a fine-grained control on the analyzed code, but it lacks
scalability, as rewriting large codes can be a tedious task. Finally, the
compiler-based approach takes advantage of the best of both by being automatic
and having access to high-level information. However, like source-to-source, the
compiler-based method is not suitable for analyzing closed-source libraries.

To the best of our knowledge, \pytracer is the first tool dedicated to the
numerical analysis of Python code. Existing tools for tracing Python code focus
on performance profiling for time (cProfile) or memory consumption (memprofile).
Anteater~\cite{faust2019anteater} is a Python tracer tool to debug Python
applications. It performs source transformations at the AST level but only deals
with native numeric Python types. Moreover, the user needs to manually tag each
variable to trace. Finally, according to the authors, Anteater does not scale to
large traces. cProfile, memprofile, and Anteater use Python decorators as the
primary instrumentation technique, a Python mechanism to instrument a function
by adding a line over a function declaration. While this method is appropriate
when targeting specific functions, it is not feasible for large codebases where
potentially unstable code sections are unknown.

% \label{sec:detecting-instabilities}
% \paragraph{Dynamically detecting numerical instabilities}
% \label{sec:mca}
% :
%  \Yohan{Either detail this section with UA/SA or remove it since MCA is introduced in 'noise injection' subsection. } \tristan{I agree, let's just remove it, the core of the contribution is on the tracing, not in the detection of instabilities.}

% % We review here the existing theoretical framework that might be used to detect numerical instabilities.
% % Our experiments rely on the Monte Carlo Arithmetic (MCA), a stochastic arithmetic allowing 
% % to detect numerical errors by adding random perturbations in the floating-point computations.
% % However, Pytracer do not rely on any assumption about the perturbation model used and thus leaves the user free to use
% % an alternative model. As such, we describe two others options that we thought might be beneficial and complementary to MCA: 
% % input data uncertainty analysis and random seed sensitivity.

% % \subsubsection{Stochastic Arithmetic}

% % The Monte Carlo Arithmetic is a variant of so-called stochastic arithmetic, whose purpose is to model round-off errors by a random variable.

% Three main approaches exist to detect numerical instabilities dynamically: stochastic arithmetic, uncertainty or sensitivity analysis~\cite{helton2006survey}, and random seed analysis~\cite{click2011quality}.
% Stochastic arithmetic leverages randomness to estimate numerical instabilities coming from the use of floating-point representations. The main idea is to treat round-off errors as random variables and characterize them statistically.
% The first published complete stochastic arithmetic framework was CESTAC~\cite{vignes1993stochastic}. In CESTAC, each floating-point computation is performed multiple times with different rounding modes: round-up or round-down. From these stochastic samples, one can then derive the number of significant digits in the computed result. Discrete Stochastic Arithmetic~\cite{vignes2004discrete} extends CESTAC with comparison operators implemented in the CADNA library.
% Monte Carlo Arithmetic (MCA)~\cite{parker1997monte} uses the same principle but introduces two differences:
% (i) a virtual precision parameter, allowing to simulate reduced working precision, and (ii) different perturbation modes to introduce perturbations
% on function inputs or output.
% Stochastic arithmetic quantifies numerical error using the number of significant digits $s$ estimated among the sampled values. A common formula to determine this number from MCA samples is presented in~\cite{parker1997monte}.
% Sohier et al.~\cite{sohier2018confidence} recently provided a generalization of this formula to include confidence intervals.

% % \subsubsection{Uncertainty, sensitivity, and random seed analysis}
% % \tristan{I'm wondering if we could remove this sub-section, as it doesn't seem to be critical to the understanding of the paper, wdyt?}
% % Uncertainty analysis (UA) is a method to estimate the output distribution given an input distribution.
% % Sensitivity analysis (SA) is a variation of UA that measures the local response to a small perturbation~\cite{loucks2017water}. 
% % Several approaches for UA and SA exist, among them differential analysis, response surface methodology, Monte Carlo analysis, and variance decomposition (see~\cite{helton2006survey} for a review). In particular, Monte Carlo analysis is a sampling-based technique that samples inputs space to estimate output response. Monte Carlo analysis shares similarities with Monte Carlo Arithmetic. For instance, the PRECISE framework~\cite{chaitin1996lectures} shares with MCA the idea of varying virtual precision to estimate output sensitivity.

% % Furthermore, the quality of a Random Number Generator is a crucial characteristic when dealing with stochastic model and differ from cryptography requirements. A good RNG provides a tradeoff between performance, period, correlation, or uniformity~\cite{james2020review} for numerical simulation.  
% On the other hand, poor-quality RNGs may have a substantial impact on results~\cite{click2011quality}. First, measuring variation induced by seed is a good indicator of program stability that we can compare against variability induced by floating-point arithmetic.

% \paragraph{Visualization of numerical data}
% \tristan{now that it's here, I'm not sure if this section really adds to the paper. It could also go.}

% Floating-point values approximate real number in the computer arithmetic but
% data manipulated by scientific applications include a large variety of mathematical objects
% beyond scalar values. 
% Indeed, scientific visualization 
% \gkmod{}{Note: You kind of jump right in to this section. Visualization of what in particular? Why do we want (need?) to visualize this sort of data?} 
% covers a wide range of numerical simulation domains, from general-purpose numerical visualization with tools such as VTK~\cite{schroeder2000visualizing} to topological data analysis~\cite{tierny2018topological}, including matrix~\cite{wu2008matrix} or tensor~\cite{kindlmann2006diffusion} visualization.
% To be able to visualize those objects is crucial for the practitioner 
% as information about the structure of the object may be as important as the basic elements that constitute it (i.e. vectors or matrices). 
% Despite the large number of existing tools to identify numerically unstable sections of source code, few allow the user to have a global view of the numerical precision of their computations or to compare several executions of the same program under various perturbations.

% Veritracer~\cite{chatelain2018veritracer} was the first tool to propose an embedded visualizer of numerical stability over the execution of C, C++, and Fortran code. Veritracer extends Verificarlo~\cite{verificarlo} with contextual information about traced variables, which allows the user to identify instabilities quickly.
% However, Veritracer neither offers a dynamic interface nor includes information about high-level structures such as \texttt{struct} or \texttt{array}.
% FPAvizual~\cite{gu2014fpavisual} visualizes the effects of floating-point arithmetic but only for educational purposes and so does not target computational intensive applications\gkmod{}{Q: What does this mean?}.
% Anteater~\cite{faust2019anteater} proposes a visualizer for Python code, but as mentioned previously, it only deals with native Python types, and the user needs to select variables to trace manually. Anteater provides an insightful reflection 
% around visual debugging beyond floating-point analysis, which has inspired the design of \pytracer's visualization interface.

% Through our analysis of SciPy and Scikit-learn, we demonstrated \pytracer's scalability, automaticity, and applicability to any Python program. The relatively low instrumentation overhead of $3.5\times$ on average enables the analysis of real applications. \pytracer's instrumentation is fully automatic and does not require manual intervention from the user. \pytracer works on any Python libraries, with any data format supported by Pickle and any numerical noise model.

% \pytracer is the first tool to enable a systematic numerical analysis of Python codes. Existing tools focus on traditional languages of the HPC community such as C, C++, and FORTRAN. However, Python's high-level abstraction allows for quickly developing applications while maintaining high performance thanks to specialized libraries such as NumPy. Therefore, \pytracer meets the numerical analysis needs of an ever-increasing number of applications.

% Furthermore, \pytracer allows users to pinpoint numerical instabilities root causes precisely by working at the function level. Indeed, evaluating the numerical stability is complex since calculations 
% are strongly correlated between them. 
% The instrumentation at the function level limits the size of the traces since only inputs and outputs of a function are dumped. It moreover decreases the overhead of the instrumentation and helps the user to identify instabilities quickly. 
% Working at function level is thus a suitable compromise between (1) analyzing at the floating-point operation level and (2) user debug prints.
% (1) Indeed, the fine-grained analysis is costly and hard to analyze due to the large amount of computations evolved in HPC codes. It overwhelms the user with too much information, making analysis more difficult but can be interesting at the final step of the understanding. 
% (2) Looking at intermediate results through print statements focuses on valuable information
% (like residuals or norms) but may miss numerically unstable functions unsuspected by the developer.
% Finally, the SciPy environment wraps a lot of computation kernels written in C, like BLAS or LAPACK.
% By working at function level, \pytracer allows the user to evaluate the numerical precision of such kernels
% and their impact on the code without getting deep into technical implementation aspects.

% Python is a popular language to prototype proof-of-concept. Integrating \pytracer at early stages of the development cycle allows identifying numerical instabilities to write code of better quality.
% \pytracer's automaticity facilitates the usage and enables its integration in unit testing.
% It can be for example used in a non-regression test to ensure that the numerical quality of a calculation
% is not degraded by updates. Moreover, the Plotly dashboard offers a quick visualization to 
% analyze more finely the potential issues.

% \pytracer can also be useful to select the hyperparameters of a code, as it is the case in machine learning.
% Indeed, we find many cases in the literature where users explore a range of hyperparameters
% to select the ones giving the best scores. With \pytracer, 
% the user has a practical way to quantify the numerical quality and to understand why hyperparameters
% lead to unstable outputs, instead of tweaking hyperparameters to make solvers converge,

% The visualization permits a global view of the code and to understand interactions between 
% computations. The Plotly dashboard help to easily navigate through the code.
% It moreover removes the need for ad-hoc scripts to visualize data. 
% Finally, the dashboard can be extended to support other data format to visualize.
% \pytracer currently supports n-dimensional NumPy arrays as well as native Python data.

% As we showed in Section~\ref{sec:impact_mca_modes}, using Full MCA mode leads to runtime errors that have a considerable impact on the proper functioning of the execution. This is unfortunate since these errors do not reflect actual numerical errors but errors related to perturbations that should not occur.
% Conversely, RR mode is far more conservative since it preserves exact operations and is easier to use even though it does not simulate all perturbations.
% From our experiments, we do not recommend the Full MCA mode for now and more researches are needed 
% to help fixing issues we encountered.

% Fixing this issue would enable the study of catastrophic cancellations, which are common sources of numerical errors. 
% In addition to the runtime errors it raises, mismatching shapes is also a problem when aggregating traces as \pytracer will not compute statistics for a set of arrays with different shapes \tristan{So do you conclude that Full MCA is not usable while RR is? If so, you should recommend to use Pytracer with RR only.} \tristan{are you still talking about Full MCA here? if not, make another paragraph. If so, make it clearer.}. Two solutions exist to overcome this issue: the first is not to instrument the \texttt{PyArray\_Arange} function with MCA. The advantage is that it is easy to set up since it requires one extra line in the exclusion file. However, the same kind of problem may appear somewhere else because, as we have seen, the evolved operations are elementary. Therefore, the chances that developers reproduce the same pattern are very probable. Therefore, the second solution is to ensure that the Full MCA mode preserves the exact operations. This solution has the advantage of not worrying about which functions should be instrumented or not with MCA, but it is hard to achieve it in practice. Indeed, we rely on the MCA implementation provided by Verificarlo that uses an internal precision to add the random perturbation that is higher than the working precision for applying the random noise. Verificarlo then rounds the result to the working precision before returning it, so any residual noise lower than the working precision is lost. Keeping track of that noise would require recording each intermediate computation in a shadow memory or keeping a tag about their exactness. This option would require more extra works than the first one but would ensure more consistent results. \tristan{I find this paragraph too verbose, you should contract it to a few key sentences.}

% \tristan{The following paragraphs are future work, not discussion. I'd remove them and flesh out the discussion from the ideas mentioned in the current 3rd paragraph}

% Also, we would like to extend the post-process analysis to aggregate traces where the execution flow path diverges. Therefore, a posteriori reconstruction is challenging as one must precisely identify the flow path taken and find synchronization points to pursue the analysis. We retrieve, for example, this divergence behavior in two general classes of numerical schemes: iterative methods and optimization processes. One solution is to integrate methods used in code coverage analysis to keep track of flow paths taken and the backtraces. However, this would increase the complexity of the postprocess analysis as well as an extra slowdown factor. However, by coupling the Gantt chart with static analysis, \pytracer would catch some divergence patterns occurring within a function. Indeed, by instrumenting at the function level, \pytracer views the function's callees (i.e., the functions called within a function) and thus should partially detect flow divergences.

% By design principle, \pytracer is flexible on the noise model used, although we only tested it with the MCA model in our experiments. Nevertheless, we would like to compare different models against MCA in future works to see the resulting differences. Furthermore, we would like to test \pytracer on parallel codes.

% Finally, we would extend \pytracer to instrument elementary arithmetic operations which can be helpful to more profound analyses. We nevertheless believe that instrumentation at the function level presents two advantages: it first limits the size of the traces since only inputs and outputs of a function are saved instead of a full report which, on the one hand, decreases the overhead of the instrumentation and, on the other hand, help the user to identify instabilities quickly. Indeed, full reporting can overwhelm the user with too much information, making analysis more difficult at first. Second, the analysis at fine granularity is interesting at the final step of the understanding. Such an analysis would require a different instrumentation approach by, for example, working at the Abstract Syntax Tree (AST) level since it is trivial to get it in Python.
% Moreover, working at the AST level allows doing dependency analysis. Another extension would be to extract and replay a function, in the same vein as the CERE tool does but for Python, and to use an AST instrumentation to analyze the numerical quality of elementary operations step by step within the fuzzy environment.

% \pytracer is a tool for visualizing the numerical quality of function inputs/outputs over time.
% It generates numerical traces that are aggregated in postprocessing to measure numerical discrepancies between executions.
% \pytracer does not assume the noise model used to estimate uncertainties but has been tested with the MCA 
% through the fuzzy ecosystem on two broadly used scientific libraries for Python with partial success.

% We demonstrated the Pytracer's capabilities to analyze intensively used Python libraries such as NumPy or Scikit-learn, such as 
% 1) Scalability. The low instrumentation overhead of 3.5x enables analysis of real applications and permits having a global representation of the numerical quality of an application. 
% 2) Automaticity. Pytracer instrumentation is fully automatic and does not require manual intervention from the user, allowing the instrumentation of 
% the NumPy, SciPy and Scikit-learn libraries.
% // that cumulate 5600 and 2043 pure Python functions and classes. 
% 3) Agnosticity. Pytracer works on any python libraries, with any data format supported by pickle, and any numerical noise model flavor.

% Pytracer is the first tool to bring numerical analysis of Python codes available to everyone. Indeed, existing tools developed in the last years focused on traditional HPC languages community such as C,C++, and FORTRAN. However, the Python high-level abstraction allows you to quickly develop applications while maintaining high performance with the help of specialized libraries like NumPy. Therefore, it is quite natural to see more fields systematically using Python, as is the case in neuroimaging. Thus, Pytracer meets the numerical analysis needs of an ever increasing number of applications writing in Python.

% By working at function level, Pytracer allows users to precisely pinpoint numerical instabilities root causes. Furthermore, the automaticity facilitates the usage and its integration in unit testing. Hence, instead of tweaking hyperparameters to make solvers converging, the user has a practical way with Pytracer to quantify the numerical quality of its computations.

% As we showed in section~\ref{sec:impact_mca_modes}, using the Full MCA mode leads to runtime errors that have a considerable impact on the proper functioning of the execution, which is unfortunate since these
% errors do not reflect actual numerical errors but errors related to perturbations that should not occur.
% Fixing this misbehavior would greatly help use Full MCA mode since this mode allows simulating catastrophic cancellations, which are familiar sources of numerical errors. By preserving exact operations, RR mode is far more conservative and so easy to use even though it does not simulate all perturbations that might be encountered.

% Also, using a post-process analysis to aggregate traces, \pytracer falls in the same pitfalls as veritracer and is limited to aggregate traces where the execution flow path extremely diverges. Therefore, a posteriori reconstruction
% is challenging as one must precisely identify the flow path taken and find synchronization points to pursue the analysis. We retrieve, for example, this divergence behavior in two general classes of numerical schemes: iterative methods and optimization processes. One solution to overcome this issue is to integrate methods used in code coverage analysis to keep track of flow paths taken and the backtraces.
% However, this would increase the complexity of the postprocess analysis as well as an extra slowdown factor.
% However, by coupling the Gantt chart with static analysis, \pytracer would catch some divergence patterns occurring within a function. Indeed, by instrumenting at the function level, \pytracer views the function's callees (i.e., the functions called within a function) and thus should partially detect flow divergences.

% By design principle, \pytracer is flexible on the noise model used, although we only tested it with the MCA model in our experiments.
% Nevertheless, we would like to compare different models against MCA in future works to see the resulting differences. 
% In particular, although \pytracer only works on a sequential application for the moment, it can be used to analyze parallel sections internal to a function.
% The only restriction is that the section does not call other functions traced by \pytracer; we would fall into the previous issue of divergence if it would not be the case.
% Since \pytracer only cares about inputs and outputs of a function, the order of internal operations is transparent for it and thus should be able to handle those cases.

% Finally, \pytracer does not instrument elementary arithmetic operations, and it can be a barrier to more profound analysis.
%Such an analysis would require a different instrumentation approach by, for example, working at the Abstract Syntax Tree (AST) level since it is trivial to get it in Python.
% An ideal way to do that would be to extract and replay a function, in the same vein as the CERE~\cite{castro2015cere} tool does but for Python, and use an AST instrumentation to analyze the numerical quality of elementary operations step by step within the fuzzy environment.
% Lastly, AST analysis can serve as a basis for finding relationships between variables, valuable information for the user.

% In addition to the runtime errors it raises, mismatching shapes is also a problem when aggregating traces as PyTracer will not compute statistics for a set of arrays with different shapes.
% Two solutions exist to overcome this issue: the first is not to instrument the \texttt{PyArray\_Arange} function with MCA. The advantage is that it is easy to set up since it requires one extra line in the exclusion file.
% However, the same kind of problem may appear somewhere else
% because, as we have seen, the operations being evolved are elementary. Therefore, the chances that developers reproduce the same pattern are very probable. The second solution is to ensure that the Full MCA mode preserves the exact operations.
% This solution has the advantage of not worrying about which functions should be instrumented or not with MCA, but it is hard to achieve it in practice. Indeed, we rely on the MCA implementation provided by Verificarlo that uses an internal precision to add the random perturbation that is higher than the working precision for applying the random noise. Verificarlo then rounds the result to the working precision before returning it, so any residual noise lower than the working precision is lost. 
% Keeping track of that noise would require recording each intermediate computation in a shadow memory or keeping a tag about their exactness.
% \Yohan{This idea emerged by discussing with Verificarlo's team. 
% I must find a way to express that and not give the impression that we discovered or suggest it.}
% This option would require more extra works than the first one but would ensure more consistent results.

\section{Conclusion}

% \tristan{I moved this from the opening of Section 4, as this reads more like a summary of \pytracer's features than an experiment description. It should be refactored in 1-2 paragraphs summarizing the main properties of \\pytracer.}

Evaluating the numerical stability of scientific Python programs is a tedious task that requires substantial expertise. Automatic tools to help users understand the behavior of their code are therefore crucial.
In this perspective, \pytracer is the first tool to offer a complete framework for analyzing and visualizing the numerical stability of Python codes. It works by automatically instrumenting Python libraries to produce numerical traces visualized through an interactive Plotly Dash server. \pytracer does not require manual code modification, and we tested it with major scientific libraries such as NumPy, SciPy, and Scikit-learn. Our results show that \pytracer is scalable, with an average instrumentation slowdown of $3.5\times$. Moreover, \pytracer's visualization interface remains responsive even with traces of up to 1TB. 
In our experiments, \pytracer enabled the systematic evaluation of 40 examples from the SciPy and scikit-learn libraries.

In future work, we plan to extend the post-processing analysis to aggregate traces when the execution flow path diverges by leveraging techniques used in code coverage analysis. Moreover, we would like to extend \pytracer to instrument elementary arithmetic operations which can be helpful to more profound analyses. Finally, although \pytracer supports multiple noise models, we only tested it with Monte Carlo Arithmetic. Therefore, we would like to compare different models against MCA. 
\pytracer is available at \href{https://github.com/big-data-lab-team/pytracer}{github://big-data-lab-team/pytracer} under MIT License.

% % use section* for acknowledgment
% \ifCLASSOPTIONcompsoc
%     % The Computer Society usually uses the plural form
%     \section*{Acknowledgments}
% \else
%     % regular IEEE prefers the singular form
%     \section*{Acknowledgment}
% \fi

% The authors would like to thank...

% Can use something like this to put references on a page
% by themselves when using endfloat and the captionsoff option.
\ifCLASSOPTIONcaptionsoff
    \newpage
\fi

\bibliographystyle{IEEEtran}
\bibliography{paper-annotated}

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
%% \begin{thebibliography}{1}

%% \bibitem{IEEEhowto:kopka}
%% H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
%%   0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.

%% \end{thebibliography}

% biography section
% 
% If you have an EPS/PDF photo (graphicx package needed) extra braces are
% needed around the contents of the optional argument to biography to prevent
% the LaTeX parser from getting confused when it sees the complicated
% \includegraphics command within an optional argument. (You could create
% your own custom macro containing the \includegraphics command to make things
% simpler here.)
%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
% or if you just want to reserve a space for a photo:

\vskip -2\baselineskip plus -1fil
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{photos/yohan.jpg}}]{Yohan Chatelain}
    received his Ph.D. degree from the University of Paris-Saclay
    (UVSQ), Versailles, France in 2019.
    He is a postdoctoral researcher in the Big Data Infrastructures for
    Neuroinformatics lab at the University of Concordia, Montreal, Canada. His
    research topics include computer arithmetic, high-performance computing, and
    reproducibility. Yohan aims at democratizing the use of
    the analysis of stability for scientific computing codes through automatic tools to
    improve numerical quality. 
\end{IEEEbiography}

\vskip -2\baselineskip plus -1fil
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{photos/NigelPhoto.png}}]{Nigel Yong Sao Young}
    was born in Mauritius Island. He is a recent Computer
    Science graduate at Concordia University. He is currently working on his
    software startup, based in the city of Montreal.
\end{IEEEbiography}

%% % if you will not have a photo at all:
\vskip -2\baselineskip plus -1fil
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{photos/greg_kiar.png}}]{Gregory Kiar}
    is a research scientist at the Child Mind Institute. Throughout his degrees
    in Biomedical Engineering, Greg has developed techniques to study biosignal
    data, a turn-key tool that allows researchers to generate maps of brain
    connectivity from diffusion MRI data, and techniques to assess and improve
    the stability of research in neuroscience. 
    Greg’s research bridges the
    fields of numerical analysis and brain imaging to evaluate and improve the
    trustworthiness of techniques used to study the brain. 
\end{IEEEbiography}

\vskip -2\baselineskip plus -1fil
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{photos/glatard_tristan.jpg}}]{Tristan Glatard}
    is Associate Professor in the Department of Computer Science and Software
    Engineering at Concordia University in Montreal, and Canada Research Chair
    (Tier II) on Big Data Infrastructures for Neuroinformatics. Before that, he
    was research scientist at the French National Centre for Scientific Research
    and Visiting Scholar at McGill University. 
\end{IEEEbiography}

%% % insert where needed to balance the two columns on the last page with
%% % biographies
%% %\newpage

%% \begin{IEEEbiographynophoto}{Jane Doe}
%% Biography text here.
%% \end{IEEEbiographynophoto}

% You can push biographies down or up by placing
% a \vfill before or after them. The appropriate
% use of \vfill depends on what kind of text is
% on the last page and whether or not the columns
% are being equalized.

%\vfill

% Can be used to pull up biographies so that the bottom of the last one
% is flush with the other column.
%\enlargethispage{-5in}

% that's all folks
\end{document}