### 2022-05-10 IEEEtran/paper-annotated.tex
Update for major revision

%% bare_jrnl_compsoc.tex
%% V1.4b
%% 2015/08/26
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
%% Computer Society journal paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/pkg/ieeetran
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall the IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. The IEEE's font choices and paper sizes can   ***
% *** trigger bugs that do not appear when using other class files.       ***                          ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

\documentclass[10pt,journal,compsoc]{IEEEtran}

\usepackage{todonotes}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{color}
\usepackage{hyperref}
\usepackage{multirow}
\usepackage{pifont}
\usepackage{xcolor,colortbl}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=blue!50!black,citecolor=blue!75!black,filecolor=black,urlcolor=blue!75!black}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{xspace}
\usepackage{makecell}
\usepackage{ulem} % provides sout
\usepackage[frozencache=true,cachedir=minted-cache]{minted} 
\usepackage[htt]{hyphenat}      
\usepackage{subcaption}
\usepackage{graphicx,import}
\usepackage{booktabs}
\usepackage{soul}
%\usepackage{cancel}
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[10pt,journal,compsoc]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/pkg/ifpdf
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
\ifCLASSOPTIONcompsoc
  % IEEE Computer Society needs nocompress option
  % requires cite.sty v4.0 or later (November 2003)
  \usepackage[nocompress]{cite}
\else
  % normal IEEE
  \usepackage{cite}
\fi
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.
%
% Note that some packages require special options to format as the Computer
% Society requires. In particular, Computer Society  papers do not use
% compressed citation ranges as is done in typical IEEE papers
% (e.g., [1]-[4]). Instead, they list every citation separately in order
% (e.g., [1], [2], [3], [4]). To get the latter we need to load the cite
% package with the nocompress option which is supported by cite.sty v4.0
% and later. Note also the use of a CLASSOPTION conditional provided by
% IEEEtran.cls V1.7 and later.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at: 
% http://www.ctan.org/pkg/graphicx
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
%\usepackage{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics.
%
% Note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/amsmath

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/pkg/algorithms
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/pkg/algorithmicx

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/array

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=footnotesize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a sans serif font rather
% than the serif font used in traditional IEEE formatting and thus the need
% to invoke different subfig.sty package options depending on whether
% compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/pkg/subfig

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure.
% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
% corrections already built into the system in which case a warning will
% be issued if an attempt is made to load fixltx2e.sty as it is no longer
% needed.
% The latest version and documentation can be found at:
% http://www.ctan.org/pkg/fixltx2e

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/stfloats
% Do not use the stfloats baselinefloat ability as the IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that the IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/pkg/dblfloatfix

%\ifCLASSOPTIONcaptionsoff
%  \usepackage[nomarkers]{endfloat}
% \let\MYoriglatexcaption\caption
% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
%\fi
% endfloat.sty was written by James Darrell McCauley, Jeff Goldberg and 
% Axel Sommerfeldt. This package may be useful when used in conjunction with 
% IEEEtran.cls'  captionsoff option. Some IEEE journals/societies require that
% submissions have lists of figures/tables at the end of the paper and that
% figures/tables without any captions are placed on a page by themselves at
% the end of the document. If needed, the draftcls IEEEtran class option or
% \CLASSINPUTbaselinestretch interface can be used to increase the line
% spacing as well. Be sure and use the nomarkers option of endfloat to
% prevent endfloat from "marking" where the figures would have been placed
% in the text. The two hack lines of code above are a slight modification of
% that suggested by in the endfloat docs (section 8.4.1) to ensure that
% the full captions always appear in the list of figures/tables - even if
% the user used the short optional argument of \caption[]{}.
% IEEE papers do not typically make use of \caption[]'s optional argument,
% so this should not be an issue. A similar trick can be used to disable
% captions of packages such as subfig.sty that lack options to turn off
% the subcaptions:
% For subfig.sty:
% \let\MYorigsubfloat\subfloat
% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
% However, the above trick will not work if both optional arguments of
% the \subfloat command are used. Furthermore, there needs to be a
% description of each subfigure *somewhere* and endfloat does not add
% subfigure captions to its list of figures. Thus, the best approach is to
% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
% and instead reference/explain all the subfigures within the main caption.
% The latest version of endfloat.sty and its documentation can obtained at:
% http://www.ctan.org/pkg/endfloat
%
% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
% later in the document, say, to conditionally put the References on a 
% page by themselves.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\newcommand{\cross}[0]{\cellcolor{red!65}\ding{53}}
\newcommand{\valid}[0]{\cellcolor{green!75!black}\ding{51}}
\newcommand{\warn}[0]{\cellcolor{orange!75}?}
\newcommand{\na}[0]{\cellcolor{gray!25}}
\newcommand{\s}[1]{\cellcolor{cyan!25}#1}
\newcommand{\scross}[0]{\ding{53}~}
\newcommand{\svalid}[0]{\ding{51}~}
\newcommand{\swarn}[0]{?~}
\newcommand{\pytracer}[0]{PyTracer\xspace}

\definecolor{lightred}{rgb}{0.6,0.0,0.0}
\definecolor{lightgreen}{rgb}{0.8,1.0,0.8}

\DeclareRobustCommand{\add}[1]{\sethlcolor{lightgreen}\hl{#1}}
\DeclareRobustCommand{\remove}[1]{\textcolor{lightred}{\sout{#1}}}

\DeclareRobustCommand{\addCite}[1]{\textcolor{lightgreen}{\mbox{#1}}}

\setlength{\arrayrulewidth}{0.1pt}

\include{href}

\makeatletter
\newcommand{\thickhline}{%
    \noalign {\ifnum 0=`}\fi \hrule height 1pt
    \futurelet \reserved@a \@xhline
}
\newcolumntype{"}{@{\hskip\tabcolsep\vrule width 1pt\hskip\tabcolsep}}
\makeatother

\lstdefinestyle{customPython}{
  belowcaptionskip=1\baselineskip,
  breaklines=true,
  xleftmargin=\parindent,
  language=Python,
  showstringspaces=false,
  basicstyle=\scriptsize\ttfamily,
  keywordstyle=\bfseries\color[rgb]{0.580, 0.000, 0.827},
  %{purple!40!lightgray},
  commentstyle=\itshape\color{green!40!black},
  identifierstyle=\bfseries\color{cyan!75!black},
  stringstyle=\color{orange},
  deletekeywords={double,float},
  classoffset=1, % starting new class
  otherkeywords={double,float},
  morekeywords={double,float},
  keywordstyle=\bfseries\color{green!55!black},
  classoffset=0
}

\lstdefinestyle{customC}{
  belowcaptionskip=1\baselineskip,
  breaklines=true,
  xleftmargin=\parindent,
  language=C,
  showstringspaces=false,
  basicstyle=\scriptsize\ttfamily,
  keywordstyle=\bfseries\color[rgb]{0.580, 0.000, 0.827},
  %{purple!40!lightgray},
  commentstyle=\itshape\color{green!40!black},
  identifierstyle=\bfseries\color{cyan!75!black},
  stringstyle=\color{orange},
  deletekeywords={double,float},
  classoffset=1, % starting new class
  otherkeywords={double,float},
  morekeywords={double,float},
  keywordstyle=\bfseries\color{green!55!black},
  classoffset=0
}

\begin{document}

\makeatletter
\let\orig@lstnumber=\thelstnumber
\newcommand\lstsetnumber[1]{\gdef\thelstnumber{#1}}
\newcommand\lstresetnumber{\global\let\thelstnumber=\orig@lstnumber}
\makeatother

%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{PyTracer: Automatically Profiling Numerical Instabilities in Python}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%
%
%\IEEEcompsocitemizethanks is a special \thanks that produces the bulleted
% lists the Computer Society journals use for "first footnote" author
% affiliations. Use \IEEEcompsocthanksitem which works much like \item
% for each affiliation group. When not in compsoc mode,
% \IEEEcompsocitemizethanks becomes like \thanks and
% \IEEEcompsocthanksitem becomes a line break with idention. This
% facilitates dual compilation, although admittedly the differences in the
% desired content of \author between the different types of papers makes a
% one-size-fits-all approach a daunting prospect. For instance, compsoc 
% journal papers have the author affiliations above the "Manuscript
% received ..."  text while in non-compsoc journals this is reversed. Sigh.

\author{Yohan Chatelain$^1$, Nigel Yong Sao Young$^1$,  Gregory Kiar$^2$,  Tristan Glatard$^1$\\
    $^1$Department of Computer Science and Software Engineering, Concordia University, Montreal, Canada\\
    $^2$Center for the Developing Brain, Child Mind Institute, New York, NY, USA}
%% \author{Michael~Shell,~\IEEEmembership{Member,~IEEE,}
%%         John~Doe,~\IEEEmembership{Fellow,~OSA,}
%%         and~Jane~Doe,~\IEEEmembership{Life~Fellow,~IEEE}% <-this % stops a space
%% \IEEEcompsocitemizethanks{\IEEEcompsocthanksitem M. Shell was with the Department
%% of Electrical and Computer Engineering, Georgia Institute of Technology, Atlanta,
%% GA, 30332.\protect\\
%% % note need leading \protect in front of \\ to get a newline within \thanks as
%% % \\ is fragile and will error, could use \hfil\break instead.
%% E-mail: see http://www.michaelshell.org/contact.html
%% \IEEEcompsocthanksitem J. Doe and J. Doe are with Anonymous University.}% <-this % stops an unwanted space
%% \thanks{Manuscript received April 19, 2005; revised August 26, 2015.}}

% note the % following the last \IEEEmembership and also \thanks - 
% these prevent an unwanted space from occurring between the last author name
% and the end of the author line. i.e., if you had this:
% 
% \author{....lastname \thanks{...} \thanks{...} }
%                     ^------------^------------^----Do not want these spaces!
%
% a space would be appended to the last name and could cause every name on that
% line to be shifted left slightly. This is one of those "LaTeX things". For
% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
% "AB" then you have to do: "\textbf{A}\textbf{B}"
% \thanks is no different in this regard, so shield the last } of each \thanks
% that ends a line with a % and do not let a space in before the next \thanks.
% Spaces after \IEEEmembership other than the last one are OK (and needed) as
% you are supposed to have spaces between the names. For what it is worth,
% this is a minor point as most people would not even notice if the said evil
% space somehow managed to creep in.

% The paper headers
\markboth{Journal of \LaTeX\ Class Files,~Vol.~14, No.~8, August~2015}%
{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Computer Society Journals}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

% The publisher's ID mark at the bottom of the page is less important with
% Computer Society journal papers as those publications place the marks
% outside of the main text columns and, therefore, unlike regular IEEE
% journals, the available text space is not reduced by their presence.
% If you want to put a publisher's ID mark on the page you can do it like
% this:
%\IEEEpubid{0000--0000/00\$00.00~\copyright~2015 IEEE}
% or like this to get the Computer Society new two part style.
%\IEEEpubid{\makebox[\columnwidth]{\hfill 0000--0000/00/\$00.00~\copyright~2015 IEEE}%
%\hspace{\columnsep}\makebox[\columnwidth]{Published by the IEEE Computer Society\hfill}}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark (Computer Society jorunal
% papers don't need this extra clearance.)

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\IEEEtitleabstractindextext{%
    \begin{abstract}
        Numerical stability is a crucial requirement of reliable scientific
        computing. However, despite the pervasiveness of Python in data science,
        analyzing large Python programs remains challenging due to the lack of
        scalable numerical analysis tools available for this language. To fill
        this gap, we developed \pytracer, a profiler to quantify numerical
        instability in Python applications. \pytracer transparently instruments
        Python code to produce numerical traces and visualize them interactively
        in a Plotly dashboard. We designed \pytracer to be agnostic to numerical
        noise model , allowing for tool evaluation through Monte-Carlo
        Arithmetic, random rounding, random data perturbation, or structured
        noise for a particular application. We illustrate \pytracer's
        capabilities by testing the numerical stability of key functions in both
        SciPy and Scikit-learn, two dominant Python libraries for mathematical
        modeling. Through these evaluations, we demonstrate \pytracer as a
        scalable, automatic, and generic framework for numerical profiling in
        Python.
    \end{abstract}

    % Note that keywords are not normally used for peerreview papers.
    \begin{IEEEkeywords}
        Computer arithmetic, Quality analysis and evaluation, Visualization
        techniques and methodologies, Tracing,  Stochastic analysis
    \end{IEEEkeywords}
}

% make the title area
\maketitle

% To allow for easy dual compilation without having to reenter the
% abstract/keywords data, the \IEEEtitleabstractindextext text will
% not be used in maketitle, but will appear (i.e., to be "transported")
% here as \IEEEdisplaynontitleabstractindextext when the compsoc 
% or transmag modes are not selected <OR> if conference mode is selected 
% - because all conference papers position the abstract like regular
% papers do.
\IEEEdisplaynontitleabstractindextext
% \IEEEdisplaynontitleabstractindextext has no effect when using
% compsoc or transmag under a non-conference mode.

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\IEEEraisesectionheading{\section{Introduction}\label{sec:introduction}}
% Computer Society journal (but not conference!) papers do something unusual
% with the very first section heading (almost always called "Introduction").
% They place it ABOVE the main text! IEEEtran.cls does not automatically do
% this for you, but you can achieve this effect with the provided
% \IEEEraisesectionheading{} command. Note the need to keep any \label that
% is to refer to the section immediately after \section in the above as
% \IEEEraisesectionheading puts \section within a raised box.

\IEEEPARstart{T}{he} scientific Python ecosystem has become a central component
of data analyses in recent years, owing to its rich offering of data
manipulation, array programming, numerical analysis, and visualization tools. In
particular, libraries such as NumPy~\cite{harris2020array},
SciPy~\cite{virtanen2020scipy}, scikit-learn~\cite{pedregosa2011scikit} or
PyTorch~\cite{paszke2019pytorch} are used in hundreds of publications each year,
and serve as a reference set of high-quality open-source core scientific
libraries. Researchers have built numerous domain-specific software tools from
this ecosystem, leveraging Python's simplicity and flexibility.

Numerical stability is a crucial requirement of reliable scientific data
processing. In unstable analyses, small numerical perturbations introduced by
data noise, software and hardware updates, or parallelization have potential to
introduce substantial deviations in final results, leading to potentially
different scientific conclusions, ultimately threatening the reliability of
computational analyses. To prevent this, numerical stability analyses need to be
conducted systematically, however, no practical tool currently exists to conduct
such analyses for Python programs.

We present PyTracer, a numerical stability profiling and visualization tool for
Python programs. \pytracer adopts a dynamic approach that evaluates numerical
stability using program execution traces, and is applicable to large programs
without manual intervention. The transparent nature of \pytracer is particularly
valuable in contrast to theoretical approaches based on condition numbers or
backward error analysis~\addCite{\cite{higham2002accuracy}}, which require
detailed modeling of the program and its numerical implementation. Similarly,
static code analysis techniques~\addCite{\cite{goubault2006static}} hardly scale
to large code bases, particularly in dynamically-typed languages such as Python.

\pytracer estimates the variability of floating-point variables by combining
program traces obtained with different numerical perturbations, which requires
(\add{a}\remove{1}) generating numerically-perturbed executions,
(\add{b}\remove{2}) tracing floating-point computations along program
executions, and (\add{c}\remove{3}) visualizing stability evaluations.
\add{PyTracer uses the ``fuzzy" Python environment instrumented with Monte-Carlo
    arithmetic to address the requirement (a). \mbox{\pytracer} addresses the
    remaining requirements with (b) a dynamic instrumentation of Python functions,
    modules, and classes through meta-programming, (c) an interactive Plotly
    dashboard to visualize stability measures throughout program executions.
    \mbox{\pytracer} is available at
    \mbox{\href{https://github.com/big-data-lab-team/pytracer.git}{github.com/big-data-lab-team/pytracer.git}}
    under MIT License.} \remove{\mbox{\pytracer} addresses these requirements with
    (1) a ``fuzzy" Python interpreter instrumented with Monte-Carlo arithmetic, (2)
    a dynamic instrumentation of Python functions, modules, and classes through
    meta-programming, (3) an interactive Plotly dashboard to visualize stability
    measures throughout program executions. }

This manuscript presents the design of \pytracer and demonstrates its potential
in various use cases. Through stability evaluations of the SciPy and
scikit-learn libraries, we demonstrate \pytracer as a practical solution to
review the numerical stability of extensive code bases.

\section{Philosophy of \pytracer}

\add{PyTracer follows a three steps approach:}
\begin{enumerate}
    \item \add{Automatic instrumentation of Python programs to save the value of inputs and outputs of function in a file.}
    \item \add{Aggregation of the traces generated to produce a statistics summary.}
    \item \add{Interactive visualization to navigate through the code execution.}
\end{enumerate}

\add{
    The main goal of PyTracer is to help the developper visualizing the numerical
    stability of Python applications. PyTracer saves time for the user by
    instrumenting and aggregating values. In the same way that GDB makes the user
    save time when debugging an application by providing all information in one
    layout, PyTracer allows the user to visualize the overall numerical quality to
    quickly detect instabilities. The PyTracer visualization layout improves result
    analysis in any case since 1) the user does not have to manually instrument its
    code to dump raw data and creates specific post-processing scripts to visualize
    it 2) PyTracer allows the user to export the raw data traced to extend the
    analysis. 
}

\add{
    The main challenge addresses by PyTracer is the transparent instrumentation of Python 
    applications. Indeed, since Python is interpreted language, one has no complete 
    view of the function that will be executed by a given program. 
    Moreover, since we target Python application used in HPC environments, one has
    most of the time no access to source codes or with limited write access.
    Therefore, PyTracer tackles this issue by instrumenting the function dynamically 
    thanks to the Python introspection.
}

\add{
    The based-trace approach used by PyTracer permits to limit the 
    overhead at runtime since the statistics computation is reported to the postprocess analysis.
    However, it limits the aggregation to deterministic executions,
    although control-flow analysis can help to merge divergent execution paths.
}

\section{\pytracer design}

% \subsection{Pytracer workflow}
\pytracer adopts the three-step workflow shown in Figure~\ref{fig:workflow}.
First, \pytracer uses \textit{monkey patching} to replace the functions called
by the application with a wrapper that saves their arguments and returned values
in a trace file.
Second, it executes the application multiple times with noise injection, and
aggregates the resulting trace files to compute summary statistics about each
traced variable. Finally, a Plotly dashboard provides interactive visualizations
highlighting unstable code sections.

% \tristan{mention monkey patching}

\begin{figure*}
    \centering
    \includegraphics[width=0.75\linewidth]{figure/workflow.pdf}
    \caption{Pytracer workflow
        % \gkmod{}{Note: replace "Execution" with "Multiple Independent
        % Executions", and maybe move it to be between the gears and the pkl
        % files.}
    }
    \label{fig:workflow}
\end{figure*}

\subsection{Python code instrumentation}

%\tristan{``Instrumentation" should appear in Figure 1}

\pytracer dynamically instruments Python modules without requiring source code
modifications in the application. To do so, it creates a new, instrumented
version of each module that preserves its original attributes so that the
application can transparently use it.
% \gkmod{}{Note: this form of monkey patching is basically global and
% transparent decorating, right?} \Yohan{Yes}
By default, \pytracer instruments all the attributes in a given module except
the special ones of the form \texttt{\_\_*\_\_}, which is common for reserved
Python methods. However, some attributes are not writable. In this case,
\pytracer \remove{preservers} \add{preserves} the original attributes and warns
the user. The user can also restrict the list of traced attributes through an
inclusion/exclusion mechanism. The main advantage of this instrumentation
approach is the lack of required modification in the application source code, in
contrast with \remove{decorator-based approaches} \add{a decorator-based
    approache~\mbox{\cite{hunt2019decorators}}}. In addition, this technique is
scalable and applicable in read-only environments such as Singularity containers
or HPC clusters.

\subsubsection{Intercepting module import}

Python loads modules through an import mechanism triggered by the
\texttt{import} keyword that involves two objects: \texttt{finders} and
\texttt{loaders}. \texttt{Finders} search for packages (or
namespaces\footnote{Python distinguishes between \texttt{regular} and
    \texttt{namespace} packages. A regular package is a directory that contains a
    \_\_init\_\_.py file and potentially subdirectories (sub-packages) that contain
    themselves a \texttt{\_\_init\_\_.py} file and so on recursively. The package
    hierarchy follows those of the directory. Namespace packages introduced in
    Python 3.3 (\href{https://www.python.org/dev/peps/pep-0420/}{PEP 420}) do not
    contain an \_\_init\_\_.py file and allow for flexible directory structures.
    Hence, parts of the package can be located in zip files, on the network, or in
    separated directories. There are no functional differences between both types of
    packages.}) on the local storage --- such as standard packages installed through
the pip package manager --- or on a distant server. \texttt{Finders} do not load
modules. Instead, they return a specification (\texttt{spec} object) that
includes information on where to find and how to load modules. \texttt{Loaders}
create module objects, initialize them, and fill import-related attributes such
as \texttt{\_\_spec\_\_}. \texttt{Loaders} then execute modules and populate
their namespace. Finally, modules are bound to their import name in the
\texttt{sys.modules} dictionary.

Python supports custom \texttt{finder} classes registered in the
\texttt{sys.meta\_path} list. When Python encounters an \texttt{import}
statement, it first looks for an existing binding in \texttt{sys.modules} and
then iterates over the \texttt{finder} classes in the \texttt{sys.meta\_path}
list until it finds the module. \pytracer adds a custom \texttt{finder} class at
the head of \texttt{sys.meta\_path} that intercepts the module import and
creates the module with a custom \texttt{loader} class.

\pytracer's \texttt{loader} class first loads the original module, then copies
it as a new instance of the \texttt{ModuleType} class. It then calls the
appropriate instrumentation function for each attribute depending on its type
(function, class, or instances) and replaces the original module in the
\texttt{sys.modules} map with the instrumented version. Finally, it updates all
existing references to the original module contained in the global symbol table.
As a result, the application will transparently call the instrumented module.

\subsubsection{Instrumenting module functions}

\remove{Listings~\mbox{\ref{fig:wrapper_creation}} and~\mbox{\ref{fig:generic_wrapper}} show
    \mbox{\pytracer}'s function instrumentation mechanism.}
For each function to instrument, \pytracer's \texttt{loader} class creates a
wrapper function, dynamically compiles it with the \texttt{compile} builtin
function, and substitutes it for the original module function.  Although simple,
this technique does not apply to callable class instances, i.e., class instances
that have a \texttt{\_\_call\_\_} attribute and are not of type
\texttt{FunctionType}. Indeed, applying the wrapper technique would 1) modify
the type of the class instance to \texttt{FunctionType}, which could cause
syntax and semantic errors, and 2) mask class attributes, leading to
\texttt{AttributeError} exceptions. \pytracer overrides the
\texttt{\_\_call\_\_} attribute with the wrapper function when possible to
overcome this issue. When the \texttt{\_\_call\_\_} attribute is not writable,
Pytracer does not instrument the class and returns a warning.

% \begin{listing}
%     \centering
%     \begin{lstlisting}[language=Python,style=customPython]
% def get_wrapper_function(module, qualname, name, function):
%     """Returns the instrumented function as a string.
%     This string will be compiled with the compile builtin function

%     Parameters:
%         module: Name of the function module
%         qualname: Qualified name of the function
%         name: Name of the function
%         function: Python object function

%     Returns:
%         wrapper_code: The code source of the wrapper
%     """
%     info = (id(function), function_module, qualname)
%     wrapper_code = f"""
%     def {function_wrapper_name}(*args, **kwargs):
%         return generic_wrapper({info},*args,**kwargs)"""
%     return wrapper_code
% \end{lstlisting}
%     \caption{Function to create the instrumented version of a function.
%         \pytracer uses the identifier of the function instead of the 
%         actual function to handle aliases and avoid duplicate instrumentation.}
%     \label{fig:wrapper_creation}
% \end{listing}

% \begin{listing}
%     \centering
%     \begin{lstlisting}[language=Python,style=customPython,]
% def generic_wrapper(self, info, *args, **kwargs):
%     """Generic wrapper saving inputs and outputs of the wrapped function.
%     The id_dict dict keeps a mapping between the original function identifier
%     and itself. Arguments are binded to avoid mispositioning.
%     Returns the actual output of the wrapped function.

%     Parameters:
%         info: Tuple with the function's id, function's module and function's name
%         *args: Positional calling arguments
%         **kwargs: Keyword calling arguments

%     Returns:
%         outputs: Outputs of the wrapped function
%     """
%     fid, module_name, function_name = info
%     function = original_function_cache.id_dict[fid]
%     bind = Binding(function, *args, **kwargs)
%     stack = self.backtrace()
%     time = elements()
%     self.inputs(time, module_name, function_name, function, bind.arguments, stack)
%     outputs = function(*bind.args, **bind.kwargs)
%     self.outputs(time, module_name, function_name, function, outputs, stack)
%     return outputs
% \end{lstlisting}
%     \caption{\pytracer's generic wrapper function that saves
%         the inputs and outputs of the wrapped function.}
%     \label{fig:generic_wrapper}
% \end{listing}

\subsubsection{Instrumenting classes}

\pytracer instruments classes by wrapping their callable attributes as described
previously. This instrumentation preserves class types, which avoids type
mismatches in the instrumented application. However, classes that describe
particular base types and NumPy's universal functions (class \texttt{ufunc})
contain read-only attributes that cannot be instrumented. By default, \pytracer
returns a warning when it encounters such classes.

\subsubsection{Instrumenting class instances}

\pytracer automatically instruments class instances since it instruments
classes. However, it cannot instrument or instantiate classes that have
read-only attributes --- in particular class \texttt{ufunc}, --- which is
problematic given their pervasiveness in scientific computing. In particular,
NumPy's universal functions include vectorized elementary mathematical and other
prevalent functions. Universal functions are implemented in C, which prevents
their modification.

To overcome this issue, \pytracer wraps the class instance into a transparent
wrap class (\texttt{twc}) that overloads 1) function
\texttt{\_\_getattribute\_\_}, called when an instance attribute is accessed ---
instrumented to return the attributes from the original class, and 2) function
\texttt{\_\_call\_\_}, used when an instance behaves like a function through the
() operator --- instrumented with function \texttt{generic\_wrapper}. The
\texttt{twc} returns the queried attribute of the instance instead of returning
its attributes, allowing transparent access to the instance.  Finally, \pytracer
overloads the builtin \texttt{type} function to preserve the type of the wrapped
instance.

\subsubsection{Instrumenting iterators}

In functional programming, iterators traverse containers lazily, meaning that
the next element in the sequence is only computed when the application uses it.
This technique allows for the manipulation of virtually infinite sequences with
finite memory. However, it implies that no complete mapping of the container
returned by an iterator exists in memory since the application computes each
element on the fly. Therefore, iterators are not serializable which implies that
\pytracer cannot save them to output traces. A workaround would be to convert
iterators into explicit containers, however, this would increase the memory
footprint significantly. \pytracer, therefore, does not instrument non-builtin
iterators.

\subsection{Detecting numerical instabilities}

\pytracer detects numerical instabilities by computing summary statistics across
multiple executions perturbed with numerical noise. While \pytracer's
instrumentation, trace aggregation, and visualization work with various types of
numerical noise, we experimented primarily with Monte-Carlo Arithmetic (MCA) as
it is an accurate model for floating-point errors~\addCite{\cite{verificarlo}}.

\subsubsection{Noise injection}
\label{sec:fuzzy}

Three  main  approaches  exist  to  detect  numerical  instabilities:
stochastic  arithmetic,  uncertainty  or  sensitivity analysis [23], and random
seed analysis [11].  \add{PyTracer is agnostic to the noise model used and therefore
    works for all methods mentioned}.  For our use cases, we focus on stochastic
arithmetic since it does not make any assumption about the noise model, in
contrast to sensitivity analysis\add{.}

Stochastic arithmetic leverages randomness to estimate numerical instabilities
coming from the use of floating-point representations. The main idea is to treat
round-off errors as random variables and characterize them statistically. Monte 
Carlo Arithmetic (MCA)~\cite{parker1997monte} uses this principle by introducing
two improvements: (i) a virtual precision parameter, allowing to simulate
reduced working precision, and (ii) different perturbation modes to introduce
perturbations on function inputs or output. MCA replaces each floating-point
number $x$ with a stochastic counterpart computed as:
\[
    inexact(x) =  x + \beta^{e_x - t}\xi
\]
where $\beta$ is the number base, $e_x$ is the magnitude of $x$, $t$ the virtual
precision and $\xi \in (-\frac{1}{2},\frac{1}{2})$ is a random uniform variable.
Virtual precision allows for the simulation of reduced working precision. MCA
can be applied in three modes \add{but we only used the} Random Rounding (RR) \add{in our
    experiments} \remove{, Precision Bounding (PB), and full MCA,} which apply
function $inexact$ to the output. \remove{, the inputs, or both, respectively.
    While the RR mode is equivalent to stochastic rounding, the PB mode can also
    identify catastrophic cancellations.} \add{Random Rounding simulates roundoff errors.}

Stochastic arithmetic quantifies numerical error using the number of significant
digits $s$ estimated among the sampled values. A common formula to determine
this number from MCA samples is presented in~\cite{parker1997monte}:
\begin{equation}
    s = -\log_{\beta}{ \left| \dfrac{\sigma}{\mu} \right|} \label{eq:sig-digits}
\end{equation}
where $\mu$ and $\sigma$ are the sample mean and standard deviation of a
variable sampled with MCA. \add{This can be seen as the ratio signal over noise.}
Sohier et al.~\cite{sohier2021confidence} recently provided a generalization of
this formula to include confidence intervals.

We enable MCA in Python programs via Verificarlo~\cite{verificarlo}, a
clang-based compiler~\cite{lattner2008llvm} that replaces floating-point
operations by a generic call to a configurable floating-point model. Several
floating-point models are
available~\cite{chatelain2019automatic,chatelain2019outils}, also called
backends. \pytracer leverages ``fuzzy"~\cite{kiar2020comparing}, a collection of
software tools compiled with Verificarlo. In particular, fuzzy provides
MCA-instrumented versions of the Python interpreter as well as the BLAS, LAPACK,
NumPy, SciPy, and scikit-learn libraries, which enables MCA for a wide range of
existing scientific Python programs. Fuzzy is available in Verificarlo's GitHub
organization at
\href{https://github.com/verificarlo/fuzzy}{\url{github.com/verificarlo/fuzzy}}.

\subsubsection{Trace format}

\pytracer stores traces in the pickle format, a binary format to serialize
Python objects. The main advantages of the pickle format compared to other
serialization approaches such as marshal or JSON are its ability to serialize
most Python objects and its native compression.  In addition, pickle writes
Python objects sequentially, which preserves the temporal ordering required by
subsequent trace analyses.

Traces store numerical values at the granularity of function calls. When the
application invokes a function, \pytracer saves a call input object containing
contextual information (function's name, module's function, stacktrace) and the
values of the function arguments. When the function returns, \pytracer saves a
similar call output object containing the returned value.

\subsubsection{Trace aggregation}

Once traces are available, \pytracer sequentially parses them and computes the
mean, standard deviation, and the number of significant digits (as in
Equation~\ref{eq:sig-digits}) of all saved values. \pytracer saves these summary
statistics in an HDF5 file, a hierarchical format that facilitates visual
browsing since only required information are loaded when requested by the user.
This operation critically relies on the ordering of call input and output
objects in the traces. Therefore, it assumes that the application is
single-threaded and that the control flow does not depend on a random state.
\add{Therefore, \mbox{\pytracer} cannot merge traces resulting from a different
    control flow path and will stop the parsing when encountering one.}

\begin{figure*}
    \centering
    \includegraphics[width=0.75\linewidth]{figure/pytracer_layout.pdf}
    \caption{Pytracer visualization layout.}
    \label{fig:visu-layout}
\end{figure*}

To aggregate traces, \pytracer successively unpickles call objects in each
trace, checks that the call objects refer to the same function and stacktrace,
and terminates the parsing otherwise. PyTracer stores function arguments or
returned values in a NumPy array for each matching call object to compute
statistics. For compound objects such as tuples, dictionaries, or objects, it
recursively parses the fields until finding numeric types or NumPy arrays of
numeric types.

\subsection{Interactive visualization}
The third main component of \pytracer is its interactive trace visualizer built
with Plotly Dash~\cite{plotly}, a Python framework for web visualization. Plotly
Dash abstracts the low-level details of web applications in a highly
customizable development interface allowing to offload heavy computations to a
server if necessary. Figure~\ref{fig:visu-layout} presents the general layout of
the \pytracer visualizer, divided into five sub-components:
\begin{enumerate}
    \item Timeline view: This view displays the values of each input and output
          for a given function at a given invocation. By default, \pytracer
          shows the number of significant digits computed across the traces. The
          mean and standard deviation across perturbed samples are also
          available. The visualizer uses an upper triangle for inputs and a
          lower triangle for outputs. \pytracer internally represents all values
          as NumPy arrays. In the case of non-scalar values or types that cannot
          be trivially cast into NumPy arrays, \pytracer splits the value into
          several variables. Hence, \pytracer will transform a class containing
          $n$ floating-point parameters into $n$ variables. Each variable is
          prefixed with the original value and postfixed with the name of the
          parameter. Finally, non-scalar values (vectors or matrices) are
          represented by their average and can be inspected with the Zoom view
          (see below).
    \item Gantt chart view: This view displays the call graph as a Gantt chart
          representing the instrumented functions. Each function call is
          represented with a bar. The time overlaps reflect the caller/callee
          relation. For example, in the figure, \texttt{function1} calls
          \texttt{function2} that itself calls \texttt{function4} and
          \texttt{function5}. This view provides insights on the calling context
          for each function, which may influence numerical stability.
    \item  Source code view: This view displays a local snapshot of the source
          code file of the variable hovered in the timeline view.
    \item Zoom view: This view provide details about non-scalar values. For
          example, a matrix may have instabilities around its diagonal but not
          in other regions. The visualizer also provides additional metrics such
          as array shape, infinite norm, or condition number.
    \item  Functions list view: This view allows the user to select the
          functions plotted in the timeline view.
\end{enumerate}

\section{Results}

We evaluated \pytracer on two widely-used Python libraries:
SciPy~\cite{virtanen2020scipy}, the main library for scientific computing in
Python, and  scikit-learn~\cite{pedregosa2011scikit}, a reference
machine-learning framework that implements a wide range of state-of-the-art
algorithms. For each library, we executed the software examples provided in
their respective GitHub repositories. We traced each example five times, having
activated MCA in the Python interpreter \add{as well as} BLAS, LAPACK, NumPy,
SciPy, and scikit-learn. \add{Only instrumented the Python interpreter would
    lead to only perturbe pure Python floating-point operations and thus to miss
    operations in compiled codes (i.e. Cython and shared libraries).} \add{We
    repeated the experiment with two MCA modes, Random Rounding (RR) and Full MCA,
    resulting in 2 experiments for each example}. We used the default virtual
precision in Verificarlo, namely 24 bits for single-precision floating-point
values and 53 bits for double-precision values, which simulates machine error.
We used Python-3.8.5, NumPy-1.19.1, SciPy-1.5.4, and scikit-learn-0.23.2.

\add{Monte Carlo Arithmetic samples the floating-point variability of a given
    program, and hence the more MCA repetitions we have, the more accurate will be
    the sampling distribution. In theory, an infinite number of MCA repetitions
    should give the exact distribution. While in practice low sample size (3$\simeq$30)
    can give a good estimate, the issue is to capture "rare events" that could make the
    computation strongly diverge from the mean. For now, determining the number of
    MCA samples to use is still an open research question. 
    Recent theoretical works}~\cite{sohier2021confidence}
\add{on the significant digits computation allow computing an associated
    confidence interval depending on the sample size and the underlying distribution of the outputs.}

% \subsection{Results overview}

% Table~\ref{tab:pytracer_results_summary} summarizes \pytracer's outcome on SciPy
% and scikit-learn examples. Columns \textit{trace} and \textit{parse} report on
% the tracing and aggregation steps with three possible outcomes: 1) green for
% success, meaning that the execution ended well and no issues were raised by the
% application, 2) red for numerical error in the application, meaning that the
% execution raised a Python error before the end, and 3) orange for errors due to
% the fuzzy noise injection that are not considered numerical errors, meaning that
% the execution raised a Python error but due to a perturbation that should not
% have been introduced (i.e. a floating-point value representing an integer).
% Overall, the tracing and parsing of SciPy and scikit-learn examples in Random
% Rounding mode was entirely successful for 32/40 examples and showed an average
% precision loss of 8 bits. The remaining examples failed due to numerical errors
% in the libraries, which we discuss hereafter. The instrumentation with Full MCA
% is more invasive and reduced the number of successful executions to 12/40. Among
% the 28/40 failed executions, one failed due to an actual numerical error
% (\texttt{Bayesian Ridge Regression} example) and the other ones failed due to
% perturbations that should not have been introduced.

% Column \textit{sigbits} in Table~\ref{tab:pytracer_results_summary} summarizes
% the precision for the main outputs of SciPy and scikit-learn examples. We
% measured the precision in significant bits on the final primary example output
% by taking the element-wise mean for non-scalar values and, if the example
% produces many outputs, the average value among all of them. The numerical
% quality across all examples varies from 0 to 52, with an average of 44
% significant bits. For both modes, we note that only two examples failed due to a
% numerical error (\texttt{face\_recognition} for RR and \texttt{Bayesian Ridge
%     Regression} for Full MCA).

\begin{table*}[]
    \centering
    \normalsize
    \begin{subfigure}[t]{\linewidth}
        \centering
        \begin{tabular}{|lll|c|c|c|}
            \hline
            \multicolumn{3}{|c}{ \multirow{2}{*}{Application} }                 & \multicolumn{3}{|c|}{Random Rounding (RR)}                                                                                                                         \\
            \cline{4-6}
                                                                                &                                                                           &                                                          & trace  & parse    & sigbits \\
            \hline
            \multicolumn{1}{|c|}{ \multirow{20}{2em}{ \rotatebox{90}{SciPy} } } & 
            \multicolumn{1}{c|}{ \multirow{3}{2em}{ \rotatebox{90}{FFT} }}      & \discreteCosineRF                                                         & \valid                                                   & \valid & \s{52}             \\ 
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \fftOneDRf                                               & \valid & \valid   & 41      \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \fftTwoDRf                                               & \valid & \valid   & \s{52}  \\
            \cline{2-6}
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{  \multirow{5}{*}{ \rotatebox{90}{Interpolation} } }
                                                                                & \interOneDRf                                                              & \valid                                                   & \valid & 51                 \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \multiRf                                                 & \valid & \valid   & \s{51}  \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \bsplineRf                                               & \valid & \valid   & 10      \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \splineOneDRf                                            & \valid & \cross   & \s{52}  \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \splineTwoDRf                                            & \valid & \valid   & 10      \\
            \cline{2-6}
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{ \multirow{12}{2em}{ \rotatebox{90}{Optimization} } }
                                                                                & \bfgsRf                                                                   & \valid                                                   & \valid & \s{48}             \\
            \cline{3-6}
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \globalRf                                                & \valid & \cross   & 17      \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & | SHGO\footnote{Simplicial Homology Global Optimization} & \na    & \na      & \s{25}  \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & | Dual Annealing                                         & \na    & \na      & 4       \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & | Differential Evolution                                 & \na    & \na      & \s{0}   \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & | Basin Hopping                                          & \na    & \na      & 40      \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & | SHGO Sobol                                             & \na    & \na      & \s{45}  \\
            \cline{3-6}
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \slsqpRf                                                 & \valid & \valid   & 45      \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \lsmRf                                                   & \valid & \valid   & \s{48}  \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \nelderRf                                                & \valid & \valid   & 44      \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \ncgRf                                                   & \valid & \valid   & \s{48}  \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \rootRf                                                  & \valid & \cross   & 52      \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \rootLargRf                                              & \valid & \cross   & \s{29}  \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \rootLargePredRf                                         & \valid & \cross   & 42      \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \trustExactRf                                            & \valid & \valid   & \s{48}  \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \trustTruncRf                                            & \valid & \valid   & 48      \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \trustNCGRf                                              & \valid & \valid   & \s{52}  \\
            \hline
            \multicolumn{2}{|c|}{ \multirow{20}{2em}{ \rotatebox{90}{Scikit-learn} } }
                                                                                & \AdaboostRf                                                               & \valid                                                   & \valid & 48                 \\
            \cline{3-6}
            \multicolumn{2}{|c|}{}                                              & \brrRf                                                                    & \valid                                                   & \valid & \s{25}             \\
            \multicolumn{2}{|c|}{}                                              & | $1^{st}$ dataset                                                        & \na                                                      & \na    & 25                 \\
            \multicolumn{2}{|c|}{}                                              & | $2^{nd}$ dataset                                                        & \na                                                      & \na    & \s{nan}            \\
            \cline{3-6}
            \multicolumn{2}{|c|}{}                                              & \onlineClassifierComparisonRf                                             & \valid                                                   & \valid & 20                 \\
            \multicolumn{2}{|c|}{}                                              & \kmeansRf                                                                 & \valid                                                   & \valid & \s{52}             \\
            \multicolumn{2}{|c|}{}                                              & \covarianceRf                                                             & \valid                                                   & \cross & 48                 \\
            \multicolumn{2}{|c|}{}                                              & \decisionRf                                                               & \valid                                                   & \valid & \s{50}             \\
            \multicolumn{2}{|c|}{}                                              & \digitsRf                                                                 & \valid                                                   & \valid & 52                 \\
            \multicolumn{2}{|c|}{}                                              & \faceRf                                                                   & \cross                                                   & \valid & \s{-}              \\
            \multicolumn{2}{|c|}{}                                              & \penaltyRf                                                                & \valid                                                   & \valid & -                  \\
            \multicolumn{2}{|c|}{}                                              & \lassoRf                                                                  & \valid                                                   & \valid & \s{48}             \\
            \multicolumn{2}{|c|}{}                                              & \hyperplaneRf                                                             & \valid                                                   & \valid & 3                  \\
            \multicolumn{2}{|c|}{}                                              & \mnistRf                                                                  & \valid                                                   & \valid & \s{43}             \\
            \multicolumn{2}{|c|}{}                                              & \multitaskRf                                                              & \valid                                                   & \valid & 50                 \\
            \multicolumn{2}{|c|}{}                                              & \ompRf                                                                    & \valid                                                   & \valid & \s{52}             \\
            \multicolumn{2}{|c|}{}                                              & \pcaRf                                                                    & \valid                                                   & \valid & 48                 \\
            \cline{3-6}
            \multicolumn{2}{|c|}{}                                              & \robustRf                                                                 & \valid                                                   & \valid & \s{50}             \\
            \multicolumn{2}{|c|}{}                                              & | Linear                                                                  & \na                                                      & \na    & 48                 \\
            \multicolumn{2}{|c|}{}                                              & | RANSAC                                                                  & \na                                                      & \na    & \s{50}             \\
            \cline{3-6}
            \multicolumn{2}{|c|}{}                                              & \toyRf                                                                    & \valid                                                   & \valid & 35                 \\
            \cline{3-6}
            \multicolumn{2}{|c|}{}                                              & \svmRf                                                                    & \valid                                                   & \valid & \s{43.5}           \\
            \multicolumn{2}{|c|}{}                                              & | Non-weighted                                                            & \na                                                      & \na    & 43                 \\
            \multicolumn{2}{|c|}{}                                              & | Weighted                                                                & \na                                                      & \na    & \s{44}             \\
            \cline{3-6}
            \multicolumn{2}{|c|}{}                                              & \tomographyRf                                                             & \valid                                                   & \cross & -                  \\
            \multicolumn{2}{|c|}{}                                              & \weightedRf                                                               & \valid                                                   & \valid & \s{44}             \\
            \hline
        \end{tabular}
    \end{subfigure}
    \caption{\pytracer execution summary on SciPy and Scikit-learn examples.
        % \tristan{Replace FFT with SciPy FFT, Interpolation with SciPy
        % Interpolation and Optimization with SciPy Optimization, or add a SciPy
        % header to these rows.} \tristan{Nitpick: you could add a color scheme
        % to sigbits (for instance shades of blue), to make the column more
        % easily readable.} \\
        \emph{trace}: outcome of \pytracer tracing, \emph{parse}: outcome of
        \pytracer trace parsing and aggregation, \emph{result}: precision for
        main outputs in significant bits. \valid: successful run, \remove{\warn: error
            from invalid noise injection,} \cross: numerical error or divergent
        traces. \textit{nan} stands for (\texttt{Not-A-Number}). Example names
        are linked to their source code.}
    \label{tab:pytracer_results_summary}
\end{table*}

\subsection{Pytracer is applicable to scientific computation libraries of reference}

Table~\ref{tab:pytracer_results_summary} summarizes \pytracer's outcome on SciPy
and scikit-learn examples. Columns \textit{trace} and \textit{parse} report on
the tracing and aggregation steps with two possible outcomes: 1) green for
success, meaning that the execution ended well and no issues were raised by the
application, and 2) red for numerical error in the application, meaning that the
execution raised a Python error before the end. \remove{, and 3)
    orange for errors due to the fuzzy noise injection that are
    not considered numerical errors, meaning that the execu-
    tion raised a Python error but due to a perturbation that
    should not have been introduced (i.e. a floating-point value
    representing an integer).}
Overall, the tracing and parsing of SciPy and scikit-learn examples in Random
Rounding mode was entirely successful for 32/40 examples and showed an average
precision loss of 8 bits. The remaining examples failed due to numerical errors
in the libraries, which we discuss hereafter. 
\remove{The instrumentation with Full MCA is
    more invasive and reduced the number of successful exe-
    cutions to 12/40. Among the 28/40 failed executions, one
    failed due to an actual numerical error (Bayesian Ridge
    Regression example) and the other ones failed due to
    perturbations that should not have been introduced.}

Column \textit{sigbits} in Table~\ref{tab:pytracer_results_summary} summarizes
the precision for the main outputs of SciPy and scikit-learn examples. We
measured the precision in significant bits on the final primary example output
by taking the element-wise mean for non-scalar values and, if the example
produces many outputs, the average value among all of them. The numerical
quality across all examples varies from 0 to 52, with an average of 44
significant bits. We note that only one example failed due to a
numerical error (\texttt{face\_recognition}
\remove{for RR and \texttt{Bayesian Ridge Regression} for full MCA}).

\noindent \remove{\textbf{3.2 Effect of MCA modes}}

\remove{
    Random Rounding perturbs only the result of a floating-point operation while
    Full MCA perturbs both the inputs and the result. Therefore, Full MCA may
    trigger catastrophic cancellations while Random Rounding only introduces
    round-off errors which generally have a lesser impact. Moreover, Random Rounding
    preserves exact operations --- operations where the result can be exactly
    represented at the virtual precision --- while Full MCA does not.}

\remove{\pytracer can only aggregate traces obtained from the same control flow path. In
    particular, program branches triggered by an unstable floating-point comparison
    lead to different control flows and result in parsing errors. For example,
    comparing a residual error or the distance between two points to a strict
    threshold in an iterative scheme generally results in unstable branches. Full
    MCA even perturbs exact values such as integers represented with floating-point
    values, leading to many execution failures. For instance, the \texttt{fft1}
    example from SciPy raised an \texttt{ValueError: operands could not be broadcast
        together with shapes (601,) (600,)} error due to an array shape mismatch. A
    closer inspection showed that one of the arrays involved in the error was
    created with the NumPy function \texttt{linspace} that itself calls function
    \texttt{arange} (Listing~\ref{fig:pyarray_range_code}). Lines 3174-3178 show
    that the function stores the array size in a floating-point value that will be
    perturbed by Full MCA, as shown in Table~\ref{tab:mca_result_linspace},
    resulting in array sizes that fluctuate between N and N+1. It is worth noting
    that 78\% of execution issues encountered in Full MCA mode instrumentation are
    due to this type of ``typing" errors.  
    Those code regions should not be perturbed by MCA since the injected noise does
    not make sense.}

\subsection{\pytracer supports the visualization of compound data type}

\add{PyTracer supports native floating point formats as well as compound data type,
    such as complex floating point values, N-dimensional arrays and arbitrary Python
    classes. For the Python classes, PyTracer recursively parses the class attributes to
    extract scalar and N-dimensional arrays floating-points values.}

Figures~\ref{fig:fft1D_inputs} and~\ref{fig:fft1D_outputs} \add{show the
    visualization of the \texttt{fft\_1D} inputs and outputs values. Results are
    complex value encoded on double precision. PyTracer allows visualizing the
    numerical precision of real (blue points) and imaginary (orange points) parts.}
\add{The} top row in Figures~\ref{fig:fft1D_inputs} and~\ref{fig:fft1D_outputs}
show the input mean value while the bottom row shows the standard deviation. The
points with low magnitude in Figure~\ref{fig:fft1D_inputs} correspond to inputs
near 0 when the input of $\sin$ or $\exp$ is close to $k\pi$, $k \in
    \mathbb{Z}$. Indeed, since MCA introduces a slight noise, the result is not
exactly 0. We can see in Figure~\ref{fig:fft1D_inputs} that the maximal noise
introduced by MCA on the inputs is in the order of $10^{-14}$, two orders of
magnitude higher than the $ulp \simeq 10^-16$ for double precision. This slight
noise introduced can be interpreted as white noise on the input signal. We can
see on the bottom row of Figure~\ref{fig:fft1D_outputs} that the frequency noise
is of the same order of of magnitude as the input noise, which is expected. The
two peaks at x=38 and x=562 corresponds to the actual frequencies of the input
signal.

Figure~\ref{fig:spline2d_rr} \add{shows the results of the spline evaluation with
    \texttt{bisplev} from the \texttt{spline\_2D} example. It computes the 2D spline
    interpolation for function $z=(x+y)e^{-6(x^2+y^2)}$, with $(x,y) \in
        [-1,1]\times[-1,1]$. The result is a 2D-array that is natively supported by
    PyTracer. We observe a significant loss of precision in places, which follows
    an interesting pattern that could be further investigated.}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figure/FFT/fft_x.pdf}
    \caption{Absolute value of the mean and standard deviation values for fft 1D
        inputs within RR mode (log scale). The real part is shown in blue and
        the imaginary part in orange.}
    \label{fig:fft1D_inputs}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figure/FFT/fft_y.pdf}
    \caption{Absolute value of the mean and standard deviation values
        for fft 1D outputs within RR mode (log scale).}
    \label{fig:fft1D_outputs}
\end{figure}

\begin{figure*}
    \centering
    \begin{subfigure}{.3\linewidth}
        \centering
        \includegraphics[width=\linewidth]{figure/spline_2d/bisplev_mean.pdf}
        \caption{Mean value}
        \label{fig:bisplev_mean}
    \end{subfigure}
    \begin{subfigure}{.3\linewidth}
        \centering
        \includegraphics[width=\linewidth]{figure/spline_2d/bisplev_mean_log.pdf}
        \caption{Absolute mean value (log)}
        \label{fig:bisplev_mean_log}
    \end{subfigure}
    \begin{subfigure}{.3\linewidth}
        \centering
        \includegraphics[width=\linewidth]{figure/spline_2d/bisplev_std_log.pdf}
        \caption{Standard deviation (log)}
        \label{fig:bisplev_std_log}
    \end{subfigure}
    \caption{2D Spline interpolation results. Figure~\ref{fig:bisplev_mean}
        shows the mean result of the 2D-spline interpolation.
        Figures~\ref{fig:bisplev_mean_log} and~\ref{fig:bisplev_std_log} show
        respectively in the logarithmic scale the absolute mean value and the
        standard deviation of the 5 samples run within RR mode. 
    }
    \label{fig:spline2d_rr}
\end{figure*}

\subsection{\pytracer allows visualizing numerical stability at a fine grained level}

\add{We illustrate on two examples how PyTracer is effective for analyzing
    the numerical stability at a fine grained level: Face recognition and Bayesian Ridge Regression.}
The face recognition example uses eigenfaces (eigenvectors) and Support Vector
Machines (SVMs) to classify faces from the Labeled Faces in the Wild dataset.
This example uses a Principal Component Analysis (PCA) to extract the eigenfaces
and then trains an SVM classifier to predict the name associated with a face.
\add{When executing within fuzzy environment, this example crashed with the following
    Python error raised: \texttt{"SVD did not converge"}.}
This example is the only one to raise an error due to the
non-convergence of the Singular Value Decomposition (SVD) performed in the PCA.
The PCA uses the Randomized SVD~\cite{halko2011finding} method
that first computes a low-rank approximation of the input matrix using a
stochastic algorithm and then uses this approximation as input for the SVD. 
\add{When walking through the code with PyTracer, we see that}
the \texttt{randomized\_svd} scikit-learn function uses the Divide \& Conquer SVD
method \texttt{gesdd}. The LAPACK library has two main methods for computing the SVD:
\texttt{gesdd} and \texttt{gesvd}. The former uses a Divide \& Conquer approach,
while the latter uses a QR decomposition. While both methods are expected to
have the same accuracy~\cite{nakatsukasa2013stable}, the Divide \& Conquer
makes the SVD to not converge with MCA.

Replacing \texttt{gesdd} for \texttt{gesvd} solves the SVD convergence issue.
\pytracer allows to quickly visualize the new precision of the result.
Figure~\ref{fig:randomized_svd_S} shows the precision of the eigenvalues
computed displayed in decreasing order with \texttt{gesvd} \add{in single precision}, 
which varies between 15 and 20
significant bits out of the 24 available in single precision. The classifier
metrics (precision, recall, F1-score, and support) are identical to the IEEE
execution for all MCA samples. \add{We can note that even with using the \texttt{gesvd},
    the precision of the singular values (vectors) is not uniform. 
    Indeed, we can see that the precision of the lowest singular values (vectors) decreases 
    until reaching 0 significant bits.}

The second example, compares a Bayesian Ridge Regression (BRR) to the Ordinary
Least Squares (OLS) estimator on a synthetic dataset and for one-dimensional
regression using polynomial feature expansion. Although the example does not
raise a runtime error, the regression coefficients computed from the fitting are
non-significant. \pytracer traces reveal that the SVD solver used is the root
cause for this error. \add{Indeed, the BayesianRidge model first computes an SVD
    before iteratively updating the coefficients until reaching the convergence
    criterion. However, PyTracer reveals an issue in the singular values computed.}
\add{We can see on Figure}~\ref{fig:BRR_svd} \add{that the last singular values is
    \texttt{NaN} due to numerical error during the SVD although the remaining
    singular values exhibit high numerical precision (between 44 and 50 bits).
    Surprinsingly, this \texttt{NaN} value does not raise issue and is silently
    propagated. Indeed, the stopping criterion compares that the sum of the absolute
    differences between coefficients of two successive iterations but coefficients
    contain \texttt{NaN} since they are computed from singular values. Hence, the
    following condition is never met \texttt{if iter\_ != 0 and
        np.sum(np.abs(coef\_old\_ - coef\_)) < self.tol:} since the comparison with a
    \texttt{NaN} is always false.}

\add{The OLS uses the Least Square to fit the model. PyTracer shows that the
    coefficients fitted are \texttt{NaN} values. By looking at the callgraph, we can
    see that the scipy calls the \texttt{lstsq} method from the
    \texttt{scipy.linalg} package which uses the \texttt{gelsd} LAPACK function.
    Figure}~\ref{fig:OLS_svd} \add{shows the singular values returned by the
    \texttt{lstsq} method shows that one singular value has a \texttt{NaN} value
    that corrupted the calculation of the coefficients. By using the \texttt{gelss},
    PyTracer shows that the \texttt{NaN} value disapeared from the singular values
    (fig.}~\ref{fig:OLS_svd_fixed}\add{) and thus does not contaminated the computation of
    the coefficients that exhibit a precision of 42 bits in average. We can see that
    the full access to all values computed by the application with PyTracer allows
    the user to quickly identify the issue.}

In conclusion, \pytracer and the MCA model were helpful to detect the
instability of the \texttt{gesdd} method mentioned in several GitHub issues,
including the LAPACK repository. Furthermore, it is worth mentioning that a Pull
Request has recently been submitted to let the user select the SVD method and
avoid this kind of issue. Thus, \pytracer can detect instabilities that affect
fundamental scientific computing components such as the LAPACK library. \add{The
    unify view of all traces allows the user to quickly navigate through the
    different numerical precision of variables which saves time for the developer
    compare to the print statement debugging.}
% The Pytracer findings can therefore be extended to many application domains. %
% \tristan{conclusion needs to be discussed. Note to self: generaliser un peu la
% conclusion: pytracer permet de se rendre compte de problemes de stabilite dont
% plusieurs librairies sont dependantes.}

\begin{figure*}
    \centering
    \begin{subfigure}{0.3\linewidth}
        \includegraphics[width=\linewidth]{figure/face_recognition/randomized_svd_ret_U_sig.pdf}
        \caption{Left matrix $U$ of the SVD decomposition}
        \label{fig:randomized_svd_U}
    \end{subfigure}
    \begin{subfigure}{0.3\linewidth}
        \includegraphics[width=\linewidth]{figure/face_recognition/svd_ret_S_sig.pdf}
        \caption{Singular values $\Sigma$ of the SVD decomposition}
        \label{fig:randomized_svd_S}
    \end{subfigure}
    \begin{subfigure}{0.3\linewidth}
        \includegraphics[width=\linewidth]{figure/face_recognition/randomized_svd_ret_V_sig.pdf}
        \caption{Right matrix $V$ of the SVD decomposition}
        \label{fig:randomized_svd_V}
    \end{subfigure}
    \caption{Precision of the \texttt{randomized\_svd} function in the
        \texttt{face\_recognition} example using the \texttt{gesvd} SVD method
        and RR. Figures~\ref{fig:randomized_svd_U},~\ref{fig:randomized_svd_S},
        and~\ref{fig:randomized_svd_V} show respectively the values $U$,
        $\Sigma$ and $V$. Numerical precision is lower for smaller singular
        values and this loss of precision translates to singular vectors. It
        would be interesting to investigate the use of numerical precision for
        dimensionality reduction. }
    \label{fig:face_recognition_svd}
\end{figure*}

\begin{figure}
    \centering
    \begin{subfigure}{0.49\linewidth}
        \includegraphics[width=\linewidth]{figure/BRR_singular_values.png}
        \caption{BRR}
        \label{fig:BRR_svd}
    \end{subfigure}
    \begin{subfigure}{0.49\linewidth}
        \includegraphics[width=\linewidth]{figure/OLS_singular_values.png}
        \caption{OLS}
        \label{fig:OLS_svd}
    \end{subfigure}
    \begin{subfigure}{0.49\linewidth}
        \includegraphics[width=\linewidth]{figure/BRR_singular_values_fixed.png}
        \caption{BRR (fixed)}
        \label{fig:BRR_svd_fixed}
    \end{subfigure}
    \begin{subfigure}{0.49\linewidth}
        \includegraphics[width=\linewidth]{figure/OLS_singular_values_fixed.png}
        \caption{OLS (fixed)}
        \label{fig:OLS_svd_fixed}
    \end{subfigure}
    \caption{\add{Numerical precision of the singular values of the SVD from the
            Bayesian Ridge model fitting (left) and from the Ordinary Least Square model
            fitting (right) in the Bayesian Ridge Regression example. Red star
            represents \texttt{NaN} value. Both SVD give \texttt{NaN} coefficients.
            Switching from Divide \& Conquer to QR factorization improve numerical precision.
            PyTracer helps to quickly visualize improvements.}}
\end{figure}

% RR: Crash due to FP32 use. Does FP64 converge? No. Does sgsvd converge? Yes

% MCA: arange issue

% File "/usr/local/lib/python3.8/site-packages/sklearn/datasets/\_lfw.py", line
%     169, in \_load\_imgs faces[i, ...] = face ValueError: could not broadcast
%     input array from shape (49,37) into shape (50,37)

% \paragraph{L1 Penalty} This test compares the percentage of zero coefficients
% for logistic regression by varying the penalty (L1, L2 and Elastic-Net) and
% regularization (1, 0.1, 0.01) parameters. A large regularization value gives
% the model more freedom than a small value, which puts more constraints on the
% model.

% of L1, L2 and Elastic-Net penalty.

% Comparison of the sparsity (percentage of zero coefficients) of solutions when
% L1, L2 and Elastic-Net penalty are used for different values of C. We can see
% that large values of C give more freedom to the model. Conversely, smaller
% values of C constrain the model more. In the L1 penalty case, this leads to
% sparser solutions. As expected, the Elastic-Net penalty sparsity is between
% that of L1 and L2. We classify 8x8 images of digits into two classes: 0-4
% against 5-9. The visualization shows coefficients of the models for varying C.

% RR: ok \\
% MCA: stable for L1 and L2, not for Len but coef have less zeroes.

% \paragraph{Lasso and Elastic-Net} This test compares the Lasso and Elastic-Net
% regression models on a corrupted signal.

% Estimates Lasso and Elastic-Net regression models on a manually generated
% sparse signal corrupted with an additive noise. Estimated coefficients are
% compared with the ground-truth.

% RR: Good (48 bits on coefs) \\
% MCA: Good like RR if we exclude one trace that fails due to arange issue

% \paragraph{MNIST classification} This test fits a logistic regression with the
% SAGA solver with the L1 penalty on the MNIST digits classification task
% subset.

% Here we fit a multinomial logistic regression with L1 penalty on a subset of
% the MNIST digits classification task. We use the SAGA algorithm for this
% purpose: this a solver that is fast when the number of samples is
% significantly larger than the number of features and is able to finely
% optimize non-smooth objective functions which is the case with the l1-penalty.
% Test accuracy reaches > 0.8, while weight vectors remains sparse and therefore
% more easily interpretable. Note that this accuracy of this l1-penalized linear
% model is significantly below what can be reached by an l2-penalized linear
% model or a non-linear multi-layer perceptron model on this dataset.

% RR: Very stable \\
% MCA: Parse by excluding a trace that failed due to an arange issue

% \paragraph{Multitask Lasso} This test compares the Lasso and MultiTaskLasso
% regression models on a corrupted signal.

% The multi-task lasso allows to fit multiple regression problems jointly
% enforcing the selected features to be the same across tasks. This example
% simulates sequential measurements, each task is a time instant, and the
% relevant features vary in amplitude over time while being the same. The
% multi-task lasso imposes that features that are selected at one time point are
% select for all time point. This makes feature selection by the Lasso more
% stable.

% RR: Stable (44~53 bits for coef) \\
% MCA: Stable like RR if we exclude a trace that failed due index issue

% \paragraph{Orthogonal matching} This test uses the Orthogonal Matching Pursuit
% algorithm for recovering a sparse signal from a noisy measurement encoded with
% a dictionary.

% RR: Very stable \\
% MCA: Very stable

% \paragraph{PCA decomposition} This test makes a PCA decomposition on a
% synthetic dataset. These figures aid in illustrating how a point cloud can be
% very flat in one direction–which is where PCA comes in to choose a direction
% that is not flat.

% rr: stable (48 bits on components) \\
% mca: stable (48 bits on components)

% \paragraph{Robust Linear Regression} This test compares the linear and  RANSAC
% regression methods on noisy data. In this example we see how to robustly fit a
% linear model to faulty data using the RANSAC algorithm.

% RR: Very stable \\
% MCA: Very stable if we exclude two traces that failed due to arange/index
% issues

% \paragraph{Segmentation Toy} This test uses spectral clustering to separate
% connected circles in an image.

% In this example, an image with connected circles is generated and spectral
% clustering is used to separate the circles.

% In these settings, the Spectral clustering approach solves the problem know as
% ‘normalized graph cuts’: the image is seen as a graph of connected voxels, and
% the spectral clustering algorithm amounts to choosing graph cuts defining
% regions while minimizing the ratio of the gradient along the cut, and the
% volume of the region.

% As the algorithm tries to balance the volume (ie balance the region sizes), if
% we take circles with different sizes, the segmentation fails.

% In addition, as there is no useful information in the intensity of the image,
% or its gradient, we choose to perform the spectral clustering on a graph that
% is only weakly informed by the gradient. This is close to performing a Voronoi
% partition of the graph.

% In addition, we use the mask of the objects to restrict the graph to the
% outline of the objects. In this example, we are interested in separating the
% objects one from the other, and not from the background.

% RR Stable \\
% MCA: Stable if we exclude three traces that failed due to index issue

% Find the optimal separating hyperplane using an SVC for classes that are
% unbalanced.

% We first find the separating plane with a plain SVC and then plot (dashed) the
% separating hyperplane with automatically correction for unbalanced classes.

% RR: stable (48 bits) \\
% MCA: stable (48 bits)

% \paragraph{Tomography} This test compares the LASSO and Ridge regression
% models to reconstruct an image from  projections for a dataset acquired from
% computed tomography.

% This test shows the reconstruction of an image from a set of parallel
% projections, acquired along different angles. Such a dataset is acquired in
% computed tomography (CT).

% Without any prior information on the sample, the number of projections
% required to reconstruct the image is of the order of the linear size l of the
% image (in pixels). For simplicity we consider here a sparse image, where only
% pixels on the boundary of objects have a non-zero value. Such data could
% correspond for example to a cellular material. Note however that most images
% are sparse in a different basis, such as the Haar wavelets. Only l/7
% projections are acquired, therefore it is necessary to use prior information
% available on the sample (its sparsity): this is an example of compressive
% sensing.

% The tomography projection operation is a linear transformation. In addition to
% the data-fidelity term corresponding to a linear regression, we penalize the
% L1 norm of the image to account for its sparsity. The resulting optimization
% problem is called the Lasso. We use the class Lasso, that uses the coordinate
% descent algorithm. Importantly, this implementation is more computationally
% efficient on a sparse matrix, than the projection operator used here.

% The reconstruction with L1 penalization gives a result with zero error (all
% pixels are successfully labeled with 0 or 1), even if noise was added to the
% projections. In comparison, an L2 penalization (Ridge) produces a large number
% of labeling errors for the pixels. Important artifacts are observed on the
% reconstructed image, contrary to the L1 penalization. Note in particular the
% circular artifact separating the pixels in the corners, that have contributed
% to fewer projections than the central disk.

% RR Divergence \\
% MCA Index issue

% Finally, taking into account the points' weight allows for a better separation
% \tristan{is this related to numerical stability? If not, you can ommit the
% comment.}.

% RR Stable (30~48 bits on coef) \\
% MCA Unstable (0 bits on coef)

% \subsection{PyAFQ}

% PyAFQ is a Python package for Automated Fiber Quantification. It aims to find
% delineation of the major fiber tracts in individual human brains and quantify
% the tissue properties within the tracts. PyAFQ \pytracer tracing produces
% traces of 1TB with an average slowdown $\times 6$ compared to the actual
% execution. Moreover, \pytracer was able to visualize the aggregated traces,
% demonstrating its scalability fluently.

\subsection{Overhead remains tractable and evolves linearly with the trace size}

\pytracer's tracing overhead consists of (1) module instrumentation time
(initialization overhead) and (2) data write time during the execution (runtime
overhead). The initialization overhead is proportional to the number of
functions and classes traced. For example, we measured an average initialization
time of 8 seconds during our experiments with  NumPy, SciPy, and Scikit-learn.
This overhead can be reduced by excluding intensively used functions such as
\texttt{numpy.array} from the instrumentation.
Figure~\ref{fig:performance_tracing} summarizes \pytracer's tracing overhead for
SciPy and scikit-learn examples. \textit{Slowdown} represents the slowdown of
\pytracer without fuzzy measured as the ratio between the actual execution time
(without \pytracer) and the execution time using \pytracer. The \textit{Slowdown
    RR} \remove{and \textit{Slowdown MCA}} shows the slowdown within the fuzzy environment.
Without fuzzy, \pytracer introduces an average slowdown of $17\times$ high for
the examples running less than 8 seconds because \pytracer spends most of its
time in the initialization step. Since this step is a constant factor of the
number of functions instrumented, we also measured the instrumentation cost
itself (\textit{Slowdown (Instrumentation)}) calculated by subtracting the
initialization time to \textit{Slowdown}. We then obtain an average slowdown of
$3.5\times$. Finally, the \remove{fuzzy instrumentation} \add{using MCA
    operations (throuhg fuzzy) instead of the native ones} increases the average
slowdown to $37\times$ for RR\remove{and $43\times$ for Full MCA}.

% \pytracer's tracing overhead consists of (1) module instrumentation time
% (initialization overhead), and (2) data write time during the execution
% (runtime overhead). The initialization overhead is proportional to the number
% of functions and classes traced. However, the initialization overhead is
% amortized \tristan{what do you mean by amortized? I don't get this sentence}
% when traced applications are above 8 seconds,  proportional to the number of
% functions and classes traced. Figure~\ref{fig:performance_tracing} summaries
% \pytracer's tracing overhead for SciPy and scikit-learn examples. We can see
% that \pytracer introduces a slowdown of reduced toif we subtract the
% initialization step \tristan{not sure where to get these numbers from the
% table, to be clarified. Are these average values? Why are there two values?}.
% Fuzzy instrumentation increases the slowdown to for RR and   for Full MCA.
% \tristan{Figure 13 is not referred in the text. Choose between the figure and
% the table, no need to put both.}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figure/performance.pdf}
    \caption{
        \pytracer tracing overheads. \textit{Slowdown} represents the slowdown
        of \pytracer without fuzzy measured as the ratio between the actual
        execution time (without \pytracer) and the execution time using
        \pytracer. \textit{Slowdown (Instrumentation)} show the actual
        instrumentation cost computed by subtracting the initialization step.
        The \textit{Slowdown (RR)} \remove{and \textit{Slowdown (MCA)} are} \add{is} the slowdown
        of \pytracer using the fuzzy environment with \add{the} RR \remove{and Full MCA} mode\remove{s}. The
        first \remove{first} \add{last} four examples measure the initialization
        steps depending on the modules instrumented. 
        % \tristan{there's too much text in this caption, move some of it to the text, the caption should just explain the difference between the 4 slowdowns. From the text I'm not sure what is the difference between slowdown and slowdown(RR)/slowdown(MCA).}
    }
    \label{fig:performance_tracing}
\end{figure}

% \begin{figure}
%     \centering
%     \includegraphics[width=\linewidth]{figure/parsing_time.pdf}
%     \caption{\pytracer time to aggregate traces.}
%     \label{fig:performance_parsing_time}
% \end{figure}

% \begin{figure}
%     \centering
%     \includegraphics[width=\linewidth]{figure/parsing.pdf}
%     \caption{\pytracer trace and aggregation files size.
%     The figure presents the size measured for the trace and the aggregation size files.}
%     \label{fig:performance_parsing_size}
% \end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figure/performance_parsing.pdf}
    \caption{\pytracer aggregation performance.
        % \tristan{change xlabel to "Trace size (bytes) and ylabel to "Aggregation time (seconds)}
    }
    \label{fig:performance_parsing}
\end{figure}

% \begin{table}
% \centering
% \begin{subfigure}[t]{.75\linewidth}
%     \centering
%     \begin{tabular}{|l|c|c|c|c|}
%     \hline 
%     Application & Original & Pytracer only & RR & MCA \\
%     \hline 
%     Loading & 0.456 & 8.152 & - & - \\
%     \hline
%     discrete\_cosine\_transform & 0.280 & 8.176 & 10.203 & - \\ 
%     fft\_1D & 0.650 & 9.075 & 11.345 & - \\
%     fft\_2D & 0.311 & 9.500 & 11.929  & 16.433   \\
%     \hline
%     interpolation\_1D & 0.377 & 8.803  & 11.954  & -   \\
%     multivariate\_data &0.626 & 8.648 &  12.711  & - \\
%     bspline & 0.109  & 9.050 & 13.396  & 16.830   \\
%     spline\_1D & 0.111 & 8.618 & 10.960   & - \\
%     spline\_2D & 0.655 & 8.649 & 19.424  & - \\
%     \hline
%     broyden\_fletcher\_goldfarb\_shanno &0.475 & 8.292 & 11.033 & 13.685 \\
%     global\_optimization &1.257 & 55.916  & 93.322  & - \\
%     slsqp & 0.469 & 8.192 & 10.795 & 12.439  \\
%     least\_square\_minimization & 0.505 & 10.151  & 13.888 & -  \\
%     nelder\_mead\_simplex & 0.471 & 8.418  & 11.382  & 12.792  \\
%     newton\_conjugate\_gradient & 0.487 & 8.548  & 11.956  & 14.281 \\
%     root\_findings & 0.462 & 8.198  & 10.787   & - \\
%     root\_finding\_large & 0.801 & 11.525 & 41.493   &  82.701  \\
%     root\_finding\_large\_preconditioned & 0.702 & 10.580 & 33.199   & 57.579  \\
%     trust\_region\_exact & 0.482 & 8.170 & 10.400  & 12.524 \\
%     trust\_region\_truncated\_generalized\_lanczos & 0.459 & 8.457 & 11.216 & 13.245  \\
%     trust\_region\_newton\_conjugate\_gradient & 0.479 & 8.538 & 11.179 & 13.488 \\
%     \hline
%     Adaboost &1.388 &14.553 &  20.999 & - \\
%     Bayesian Ridge Regression & 1.086 & 10.344 & 90.299 & -  \\
%     Online classifier comparison & 15.964 & 72.691 & 2104.114 & -  \\
%     K-Means Clustering & 1.621 & 17.614  & 34.005 & 52.902   \\
%     Covariance estimation & 1.357 & 11.755 & 23.964  & -   \\
%     Decision Tree Regression & 1.185 & 11.194 & 15.760 & 15.918  \\
%     Digits Classification & 1.672 & 12.725 & 40.634 & - \\
%     Face Recognition & 32.159  & 74.356 & - & - \\
%     L1 Penalty & 1.568 & 12.591 & 65.547 & 128.220 \\
%     Lasso and elastic net & 1.057 & 9.903 & 17.078 & -  \\
%     Separating hyperplan & 1.376 & 11.532  & 22.333 & 25.036 \\
%     MNIST classification & 19.650 & 37.838 & 402.389  & - \\
%     Multitask Lasso & 1.090 &10.196 & 15.025  & - \\
%     Orthogonal matching & 1.161 & 12.129 & 21.779   & 31.135 \\
%     PCA decomposition & 1.239 & 10.642 & 15.520 & 17.952    \\
%     Robust Linear Regression & 1.370 & 11.144  & 14.280  & - \\
%     Segmentation toy &1.679 & 22.385 & 87.811  & - \\
%     SVM & 1.403 & 11.737 & 23.203 & 24.126   \\
%     Tomography & 11.108 & 40.177  & 1376.134 & -  \\
%     Weighted samples & 1.358 & 11.768 & 21.255 & 20.541   \\
%     \hline
%     \end{tabular}
% \end{subfigure}
%     \caption{\pytracer tracing overhead in seconds. \tristan{what is the unit in this table?}}
%     \label{tab:pytracer_overhead}
% \end{table}

% \begin{table}[]
%     \centering
%     \begin{tabular}{c|c|c|c|c}
%          Pytracer & Verificarlo Instrumentation & Backend & Time (hours:minutes:seconds) & Overhead \\
%          \hline 
%          No & No & - & 00:30:51 & 1 \\
%          No & Yes & IEEE & 01:32:56 & 3 \\
%          No & Yes & RR & 16:52:08 & 32 \\
%          Yes & No & - & 03:11:07 & 6  \\
%          Yes & Yes & IEEE & 11:14:33 & 22 \\
%          Yes & Yes & RR & 25:05:54 & 48 \\
%     \end{tabular}
%     \caption{Overheads of \pytracer on PyAFQ experiment with and without fuzzy.}
%     \label{tab:pytracer_overhead}
% \end{table}

% \begin{table}[]
%     \centering
%     \begin{tabular}{c|c|c|c|c}
%          Pytracer & Verificarlo Instrumentation & Backend & Time (hours:minutes:seconds) & Overhead \\
%          \hline 
%          No & No & - & 15 & 1 \\
%          No & Yes & IEEE & 49 & 3 \\
%          No & Yes & RR &  & 441 & 29 \\
%          Yes & No & - & 03:11:07 & 6  \\
%          Yes & Yes & IEEE & 11:14:33 & 22 \\
%          Yes & Yes & RR & 25:05:54 & 48 \\
%     \end{tabular}
%     \caption{Caption}
%     \label{tab:pytracer_overhead}
% \end{table}

Figure~\ref{fig:performance_parsing} shows the postprocessing time and trace
sizes produced by \pytracer. The linear correlation between the log of the size
and the log of the time shows a significant correlation (p-value $\leq 10^{-8}$
\remove{for RR and p-value $\leq 10^{-3}$ for MCA}).
% \tristan{Mention linear correlation with p-value here.}

In conclusion, while \pytracer's overhead is substantial, it remains tractable
for real-life examples.

\section{Discussion}

% Numpy   2509 (2403 not tests) (1848) functions, 1196 (1154) (817) classes (pure Python)
% Scipy   4031 (2187) (506) functions, 1532 (472) (109) classes (pure Python)
% Sklearn 4807 (1010) (81) functions, 546 (417) (41) classes (pure Python)

% \pytracer is a tool for visualizing the numerical quality of function inputs
% and outputs along the execution of a Python program. It generates numerical
% traces and post-processes them to measure numerical discrepancies between
% executions. \pytracer does not require any particular noise model to estimate
% uncertainties. We tested it with MCA through the fuzzy ecosystem on two
% broadly used Python libraries.

Through our analysis of SciPy and scikit-learn, we demonstrated \pytracer's
scalability, automaticity, and applicability to a wide range of Python program.
The relatively low instrumentation overhead of 3.5$\times$ on average enables
the analysis of real-scale applications. \pytracer's instrumentation is fully
automatic and does not require manual intervention from the user. Furthermore,
\pytracer works on any Python libraries, with any data format supported by
pickle and any numerical noise model.

\pytracer is the first tool to enable a systematic numerical analysis of Python
codes. Existing tools focus on traditional languages of the HPC community, such
as C, C++, and FORTRAN. However, Python's high-level abstraction allows for
quick application implementations while maintaining high-performance thanks to
specialized libraries such as NumPy. Therefore, \pytracer meets the numerical
analysis needs of an ever-increasing number of applications.

\pytracer allows users to pinpoint the root causes of numerical instabilities by
working at the function level. Numerical stability evaluations are complex since
calculations are strongly interdependent. Function-level instrumentation limits
the size of the traces since only inputs and outputs of a function are dumped.
Moreover, it decreases instrumentation overhead and helps the user identify
instabilities quickly. Working at the function level is thus a suitable
compromise between analysis conducted at the floating-point operation level and
user debug prints. Indeed,  fine-grained analyses are costly and hard to
interpret as they overwhelm users with information. In contrast, looking at
intermediate results through print statements is lightweight but not reliable
since developers may not systematically instrument unstable functions. Finally,
the SciPy environment wraps many computation kernels written in C, such as BLAS
or LAPACK. By working at the function level, \pytracer allows the user to
evaluate the numerical precision of such kernels and their impact on the code
without getting deep into technical implementation aspects.

Integrating \pytracer at the early stages of the development cycle would improve
the overall numerical quality of the developed code. \pytracer's automaticity
facilitates its usage and enables its integration in unit testing. For example,
\pytracer could be used in regression tests to ensure that updates do not
degrade the numerical quality of a calculation.

\pytracer can also be useful for selecting hyperparameters, as is the case in
machine learning. Indeed, it is common for data scientists to explore a range of
hyperparameters and select the ones giving the best scores. With \pytracer, the
user has a practical way to quantify numerical quality and understand why
hyperparameters produce inconsistent results instead of tweaking hyperparameters
to make solvers converge. For instance, the Face Recognition example
(Figure~\ref{fig:face_recognition_svd}) shows how numerical precision could
inform the selection of the number of components in dimensionality reduction.

The visualization provides a global view of the code and interactions between
computations. In addition, the Plotly dashboard helps navigate through the code
easily, without the need for ad-hoc scripts to visualize data. Finally, the
dashboard can be extended to support other data formats to visualize. At this
stage, \pytracer currently supports n-dimensional NumPy arrays as well as native
Python data.

\remove{
    As we showed in Section~\ref{sec:impact_mca_modes}, using Full MCA mode leads to
    runtime errors that have a considerable impact on the proper functioning of the
    execution, which is unfortunate since these errors do not reflect actual
    numerical errors but are related to perturbations that should not occur.
    Conversely, RR mode is far more conservative since it preserves exact operations
    and is easier to use even though it does not simulate all perturbations.
    Therefore, from our experiments, we recommend the use of RR over Full MCA. More
    research is required to address the issues encountered with Full MCA. 
}

Finally, the wide range of results precision obtained from our experiments
demonstrate that looking at the numerical variability is essential and that such
analyses would benefit from being conducted systematically. We showed that
well-known scientific libraries are prone to numerical error even though they
have been well tested for decades. The omnipresence of these libraries in
scientific codes should thus impel users to integrate numerical analysis in
their development workflow in addition to testing their implementation.

\section{Related work}

Several tools have been developed to assess numerical quality, and can be
divided into static and dynamic approaches. Static approaches typically analyze
source code, while dynamic approaches trace floating-point computations
throughout executions. A detailed review of these approaches is presented
in~\cite{cherubin2020tools}. 
% \pytracer adopts the dynamic approach due to its scalability to large code
% bases. Dynamic approaches typically trace floating-point computations, detect
% instabilities in traces, and visualize summary statistics about instabilities.
% This section reviews existing approaches for each of these steps.
% \label{sec:soa} \paragraph{Numerical tracing techniques}
Numerous tools exists to trace floating-point computations for C, C++, or
FORTRAN programs due to the prevalence of these languages in High-Performance
Computing (HPC). The main tracing techniques are \textit{source-to-source},
\textit{compiler-based transformations}, and \textit{dynamic binary
    instrumentation}.

% \gkmod{}{Note: this paragraph just reads like a list with no real
% interpretation/commentary. I think it would be better to state the strengths
% of this technique, cite examples, and then state limitations/why it isn't the
% only technique used.}
The source-to-source technique requires a rewriting of the application to modify
floating-point types and provides a fine-grained control on the analyzed code.
CADNA~\cite{jezequel2008cadna} is a library for C, C++, and Fortran implementing
CESTAC~\cite{vignes1993stochastic} stochastic arithmetic.
Shaman~\cite{demeure_phd} is a C++ library that uses a first-order error model
to propagate numerical error. MCAlib~\cite{frechtling2015mcalib} is a C library
that implements the Monte Carlo Arithmetic (MCA)~\cite{parker1997monte} using
the MPFR~\cite{fousse2007mpfr} library. \add{ADAPT \mbox{~\cite{menon2018adapt}} uses
    Automatic Differentiation to compute a sensitivity profile for each variable of
    given program}. Finally, the work in~\cite{tang2016software} proposed a
source-to-source framework for executing targeted code in infinite and fixed
precision with and without stochastic arithmetic. The main drawback of the
source-to-source approach is its lacks of scalability since rewriting large
codes can be a tedious task.

% \gkmod{}{Note: similar to the above, this is just a list rather than a
% highlight of how the compiler-based approach can be more scalable than
% source-to-source because it requires less manual intervention into a codebase
% (e.g.), then have your list, and finally say why it is still limited/not a
% global fix to this problem.}

The compiler-based approach instruments floating-point expressions at compile
time and so allows the user to automatically instrument large codebases.
Verificarlo~\cite{verificarlo} is a compiler that supports different types of
arithmetic instrumentations, including MCA. The work in~\cite{bao2013fly}
modified the GNU Compiler Collection (GCC) to track local floating-point errors
across executions. pLiner~\cite{guo2020pliner} is a root-cause analyzer based on
the Clang compiler that detects floating-point operations responsible for result
variability using a source-to-source transformation at the Abstract Syntax Tree
(AST) level to rewrite parts of code with higher precision.
PFPSanitizer~\cite{chowdhary2020debugging,chowdhary2021parallel} \add{and
    NSan~\mbox{\cite{courbet2021nsan}} are compiler based approaches}
that use shadow execution to detect numerical issues using higher
precision.
FLiT~\cite{sawaya2017flit} is a framework to detect variability
induced by compilers and their optimizations. The work
in~\cite{wang2012development} proposes a numerical debugger based on
GDB~\cite{stallman1988debugging} for Discrete Stochastic Arithmetic (DSA) on
FPGA as an Eclipse plugging. Similarly, Cadtrace~\cite{jezequel2008cadna} and
Shaman propose a GDB-based tool to use the CADNA and Shaman libraries with GDB,
respectively. The main limitation of the compiler-based approach is that one
must have access to the source code.

Dynamic Binary Instrumentation (DBI) operates directly on executables, without
the need for recompilation or manual changes. Therefore, it is applicable to any
programming language. CraftHPC~\cite{lam2013dynamic} uses DBI to detect
catastrophic cancellations at runtime. Verrou~\cite{fevotte2016verrou} is a
Valgrind~\cite{nethercote2007valgrind} based tool that dynamically replaces
floating-point operations with their stochastic arithmetic counterparts.
FPDebug~\cite{benz2012dynamic} uses DBI to detect numerical inaccuracies by
using shadow computations with higher precision.
Herbgrin\remove{g}\add{d}~\cite{sanchez2017finding} is a Valgrind-based tool to
detect numerical instabilities. It uses a shadow memory to detect precision
losses by comparison with results obtained \remove{at} \add{in} higher
precision. It is combined with symbolic computation to backtrack the root of the
error. We can note that all methods based on a high-precision oracle suppose
that computing with a large number of bits is sufficient to obtain an accurate
result, which is not always true (see for example, the famous Muller's
sequence~\cite{bajard1996introduction}). Although versatile, the DBI looses
high-level information useful to understand the source-code logic.

% \gkmod{}{Note: I like this comparison here, but do think that the comments
% above about providing some more info when each is mentioned would still help
% this section.}
Being agnostic to the programming language is a serious advantage of the DBI
method compared to source-to-source and compiler-bases methods. However, working
at the binary level makes it difficult to access high-level information needed
to debug and understand the source-code logic. Conversely, the source-to-source
approach provides a fine-grained control on the analyzed code, but it lacks
scalability, as rewriting large codes can be a tedious task. Finally, the
compiler-based approach takes advantage of the best of both by being automatic
and having access to high-level information. However, like source-to-source, the
compiler-based method is not suitable for analyzing closed-source libraries.

To the best of our knowledge, \pytracer is the first tool dedicated to the
numerical analysis of Python code. Existing tools for tracing Python code focus
on performance profiling for time (cProfile) or memory consumption (memprofile).
Anteater~\cite{faust2019anteater} is a Python tracer tool to debug Python
applications. It performs source transformations at the AST level but only deals
with native numeric Python types. Moreover, the user needs to manually tag each
variable to trace. Finally, according to the authors, Anteater does not scale to
large traces. cProfile, memprofile, and Anteater use Python decorators as the
primary instrumentation technique, a Python mechanism to instrument a function
by adding a line over a function declaration. While this method is appropriate
when targeting specific functions, it is not feasible for large codebases where
potentially unstable code sections are unknown.

% \label{sec:detecting-instabilities} \paragraph{Dynamically detecting numerical
% instabilities} \label{sec:mca}
% :
%  \Yohan{Either detail this section with UA/SA or remove it since MCA is
%  introduced in 'noise injection' subsection. } \tristan{I agree, let's just
%  remove it, the core of the contribution is on the tracing, not in the
%  detection of instabilities.}

% % We review here the existing theoretical framework that might be used to
% detect numerical instabilities. % Our experiments rely on the Monte Carlo
% Arithmetic (MCA), a stochastic arithmetic allowing % to detect numerical
% errors by adding random perturbations in the floating-point computations. %
% However, Pytracer do not rely on any assumption about the perturbation model
% used and thus leaves the user free to use % an alternative model. As such, we
% describe two others options that we thought might be beneficial and
% complementary to MCA: % input data uncertainty analysis and random seed
% sensitivity.

% % \subsubsection{Stochastic Arithmetic}

% % The Monte Carlo Arithmetic is a variant of so-called stochastic arithmetic,
% whose purpose is to model round-off errors by a random variable.

% Three main approaches exist to detect numerical instabilities dynamically:
% stochastic arithmetic, uncertainty or sensitivity
% analysis~\cite{helton2006survey}, and random seed
% analysis~\cite{click2011quality}. Stochastic arithmetic leverages randomness
% to estimate numerical instabilities coming from the use of floating-point
% representations. The main idea is to treat round-off errors as random
% variables and characterize them statistically. The first published complete
% stochastic arithmetic framework was CESTAC~\cite{vignes1993stochastic}. In
% CESTAC, each floating-point computation is performed multiple times with
% different rounding modes: round-up or round-down. From these stochastic
% samples, one can then derive the number of significant digits in the computed
% result. Discrete Stochastic Arithmetic~\cite{vignes2004discrete} extends
% CESTAC with comparison operators implemented in the CADNA library. Monte Carlo
% Arithmetic (MCA)~\cite{parker1997monte} uses the same principle but introduces
% two differences: (i) a virtual precision parameter, allowing to simulate
% reduced working precision, and (ii) different perturbation modes to introduce
% perturbations on function inputs or output. Stochastic arithmetic quantifies
% numerical error using the number of significant digits $s$ estimated among the
% sampled values. A common formula to determine this number from MCA samples is
% presented in~\cite{parker1997monte}. Sohier et al.~\cite{sohier2018confidence}
% recently provided a generalization of this formula to include confidence
% intervals.

% % \subsubsection{Uncertainty, sensitivity, and random seed analysis} %
% \tristan{I'm wondering if we could remove this sub-section, as it doesn't seem
% to be critical to the understanding of the paper, wdyt?} % Uncertainty
% analysis (UA) is a method to estimate the output distribution given an input
% distribution. % Sensitivity analysis (SA) is a variation of UA that measures
% the local response to a small perturbation~\cite{loucks2017water}. % Several
% approaches for UA and SA exist, among them differential analysis, response
% surface methodology, Monte Carlo analysis, and variance decomposition
% (see~\cite{helton2006survey} for a review). In particular, Monte Carlo
% analysis is a sampling-based technique that samples inputs space to estimate
% output response. Monte Carlo analysis shares similarities with Monte Carlo
% Arithmetic. For instance, the PRECISE framework~\cite{chaitin1996lectures}
% shares with MCA the idea of varying virtual precision to estimate output
% sensitivity.

% % Furthermore, the quality of a Random Number Generator is a crucial
% characteristic when dealing with stochastic model and differ from cryptography
% requirements. A good RNG provides a tradeoff between performance, period,
% correlation, or uniformity~\cite{james2020review} for numerical simulation.  
% On the other hand, poor-quality RNGs may have a substantial impact on
% results~\cite{click2011quality}. First, measuring variation induced by seed is
% a good indicator of program stability that we can compare against variability
% induced by floating-point arithmetic.

% \paragraph{Visualization of numerical data} \tristan{now that it's here, I'm
% not sure if this section really adds to the paper. It could also go.}

% Floating-point values approximate real number in the computer arithmetic but
% data manipulated by scientific applications include a large variety of
% mathematical objects beyond scalar values. Indeed, scientific visualization
% \gkmod{}{Note: You kind of jump right in to this section. Visualization of
% what in particular? Why do we want (need?) to visualize this sort of data?}
% covers a wide range of numerical simulation domains, from general-purpose
% numerical visualization with tools such as VTK~\cite{schroeder2000visualizing}
% to topological data analysis~\cite{tierny2018topological}, including
% matrix~\cite{wu2008matrix} or tensor~\cite{kindlmann2006diffusion}
% visualization. To be able to visualize those objects is crucial for the
% practitioner as information about the structure of the object may be as
% important as the basic elements that constitute it (i.e. vectors or matrices).
% Despite the large number of existing tools to identify numerically unstable
% sections of source code, few allow the user to have a global view of the
% numerical precision of their computations or to compare several executions of
% the same program under various perturbations.

% Veritracer~\cite{chatelain2018veritracer} was the first tool to propose an
% embedded visualizer of numerical stability over the execution of C, C++, and
% Fortran code. Veritracer extends Verificarlo~\cite{verificarlo} with
% contextual information about traced variables, which allows the user to
% identify instabilities quickly. However, Veritracer neither offers a dynamic
% interface nor includes information about high-level structures such as
% \texttt{struct} or \texttt{array}. FPAvizual~\cite{gu2014fpavisual} visualizes
% the effects of floating-point arithmetic but only for educational purposes and
% so does not target computational intensive applications\gkmod{}{Q: What does
% this mean?}. Anteater~\cite{faust2019anteater} proposes a visualizer for
% Python code, but as mentioned previously, it only deals with native Python
% types, and the user needs to select variables to trace manually. Anteater
% provides an insightful reflection around visual debugging beyond
% floating-point analysis, which has inspired the design of \pytracer's
% visualization interface.

% Through our analysis of SciPy and Scikit-learn, we demonstrated \pytracer's
% scalability, automaticity, and applicability to any Python program. The
% relatively low instrumentation overhead of $3.5\times$ on average enables the
% analysis of real applications. \pytracer's instrumentation is fully automatic
% and does not require manual intervention from the user. \pytracer works on any
% Python libraries, with any data format supported by Pickle and any numerical
% noise model.

% \pytracer is the first tool to enable a systematic numerical analysis of
% Python codes. Existing tools focus on traditional languages of the HPC
% community such as C, C++, and FORTRAN. However, Python's high-level
% abstraction allows for quickly developing applications while maintaining high
% performance thanks to specialized libraries such as NumPy. Therefore,
% \pytracer meets the numerical analysis needs of an ever-increasing number of
% applications.

% Furthermore, \pytracer allows users to pinpoint numerical instabilities root
% causes precisely by working at the function level. Indeed, evaluating the
% numerical stability is complex since calculations are strongly correlated
% between them. The instrumentation at the function level limits the size of the
% traces since only inputs and outputs of a function are dumped. It moreover
% decreases the overhead of the instrumentation and helps the user to identify
% instabilities quickly. Working at function level is thus a suitable compromise
% between (1) analyzing at the floating-point operation level and (2) user debug
% prints. (1) Indeed, the fine-grained analysis is costly and hard to analyze
% due to the large amount of computations evolved in HPC codes. It overwhelms
% the user with too much information, making analysis more difficult but can be
% interesting at the final step of the understanding. (2) Looking at
% intermediate results through print statements focuses on valuable information
% (like residuals or norms) but may miss numerically unstable functions
% unsuspected by the developer. Finally, the SciPy environment wraps a lot of
% computation kernels written in C, like BLAS or LAPACK. By working at function
% level, \pytracer allows the user to evaluate the numerical precision of such
% kernels and their impact on the code without getting deep into technical
% implementation aspects.

% Python is a popular language to prototype proof-of-concept. Integrating
% \pytracer at early stages of the development cycle allows identifying
% numerical instabilities to write code of better quality. \pytracer's
% automaticity facilitates the usage and enables its integration in unit
% testing. It can be for example used in a non-regression test to ensure that
% the numerical quality of a calculation is not degraded by updates. Moreover,
% the Plotly dashboard offers a quick visualization to analyze more finely the
% potential issues.

% \pytracer can also be useful to select the hyperparameters of a code, as it is
% the case in machine learning. Indeed, we find many cases in the literature
% where users explore a range of hyperparameters to select the ones giving the
% best scores. With \pytracer, the user has a practical way to quantify the
% numerical quality and to understand why hyperparameters lead to unstable
% outputs, instead of tweaking hyperparameters to make solvers converge,

% The visualization permits a global view of the code and to understand
% interactions between computations. The Plotly dashboard help to easily
% navigate through the code. It moreover removes the need for ad-hoc scripts to
% visualize data. Finally, the dashboard can be extended to support other data
% format to visualize. \pytracer currently supports n-dimensional NumPy arrays
% as well as native Python data.

% As we showed in Section~\ref{sec:impact_mca_modes}, using Full MCA mode leads
% to runtime errors that have a considerable impact on the proper functioning of
% the execution. This is unfortunate since these errors do not reflect actual
% numerical errors but errors related to perturbations that should not occur.
% Conversely, RR mode is far more conservative since it preserves exact
% operations and is easier to use even though it does not simulate all
% perturbations. From our experiments, we do not recommend the Full MCA mode for
% now and more researches are needed to help fixing issues we encountered.

% Fixing this issue would enable the study of catastrophic cancellations, which
% are common sources of numerical errors. In addition to the runtime errors it
% raises, mismatching shapes is also a problem when aggregating traces as
% \pytracer will not compute statistics for a set of arrays with different
% shapes \tristan{So do you conclude that Full MCA is not usable while RR is? If
% so, you should recommend to use Pytracer with RR only.} \tristan{are you still
% talking about Full MCA here? if not, make another paragraph. If so, make it
% clearer.}. Two solutions exist to overcome this issue: the first is not to
% instrument the \texttt{PyArray\_Arange} function with MCA. The advantage is
% that it is easy to set up since it requires one extra line in the exclusion
% file. However, the same kind of problem may appear somewhere else because, as
% we have seen, the evolved operations are elementary. Therefore, the chances
% that developers reproduce the same pattern are very probable. Therefore, the
% second solution is to ensure that the Full MCA mode preserves the exact
% operations. This solution has the advantage of not worrying about which
% functions should be instrumented or not with MCA, but it is hard to achieve it
% in practice. Indeed, we rely on the MCA implementation provided by Verificarlo
% that uses an internal precision to add the random perturbation that is higher
% than the working precision for applying the random noise. Verificarlo then
% rounds the result to the working precision before returning it, so any
% residual noise lower than the working precision is lost. Keeping track of that
% noise would require recording each intermediate computation in a shadow memory
% or keeping a tag about their exactness. This option would require more extra
% works than the first one but would ensure more consistent results. \tristan{I
% find this paragraph too verbose, you should contract it to a few key
% sentences.}

% \tristan{The following paragraphs are future work, not discussion. I'd remove
% them and flesh out the discussion from the ideas mentioned in the current 3rd
% paragraph}

% Also, we would like to extend the post-process analysis to aggregate traces
% where the execution flow path diverges. Therefore, a posteriori reconstruction
% is challenging as one must precisely identify the flow path taken and find
% synchronization points to pursue the analysis. We retrieve, for example, this
% divergence behavior in two general classes of numerical schemes: iterative
% methods and optimization processes. One solution is to integrate methods used
% in code coverage analysis to keep track of flow paths taken and the
% backtraces. However, this would increase the complexity of the postprocess
% analysis as well as an extra slowdown factor. However, by coupling the Gantt
% chart with static analysis, \pytracer would catch some divergence patterns
% occurring within a function. Indeed, by instrumenting at the function level,
% \pytracer views the function's callees (i.e., the functions called within a
% function) and thus should partially detect flow divergences.

% By design principle, \pytracer is flexible on the noise model used, although
% we only tested it with the MCA model in our experiments. Nevertheless, we
% would like to compare different models against MCA in future works to see the
% resulting differences. Furthermore, we would like to test \pytracer on
% parallel codes.

% Finally, we would extend \pytracer to instrument elementary arithmetic
% operations which can be helpful to more profound analyses. We nevertheless
% believe that instrumentation at the function level presents two advantages: it
% first limits the size of the traces since only inputs and outputs of a
% function are saved instead of a full report which, on the one hand, decreases
% the overhead of the instrumentation and, on the other hand, help the user to
% identify instabilities quickly. Indeed, full reporting can overwhelm the user
% with too much information, making analysis more difficult at first. Second,
% the analysis at fine granularity is interesting at the final step of the
% understanding. Such an analysis would require a different instrumentation
% approach by, for example, working at the Abstract Syntax Tree (AST) level
% since it is trivial to get it in Python. Moreover, working at the AST level
% allows doing dependency analysis. Another extension would be to extract and
% replay a function, in the same vein as the CERE tool does but for Python, and
% to use an AST instrumentation to analyze the numerical quality of elementary
% operations step by step within the fuzzy environment.

% \pytracer is a tool for visualizing the numerical quality of function
% inputs/outputs over time. It generates numerical traces that are aggregated in
% postprocessing to measure numerical discrepancies between executions.
% \pytracer does not assume the noise model used to estimate uncertainties but
% has been tested with the MCA through the fuzzy ecosystem on two broadly used
% scientific libraries for Python with partial success.

% We demonstrated the Pytracer's capabilities to analyze intensively used Python
% libraries such as NumPy or Scikit-learn, such as 
% 1) Scalability. The low instrumentation overhead of 3.5x enables analysis of
%    real applications and permits having a global representation of the
%    numerical quality of an application. 
% 2) Automaticity. Pytracer instrumentation is fully automatic and does not
%    require manual intervention from the user, allowing the instrumentation of
%    the NumPy, SciPy and Scikit-learn libraries. // that cumulate 5600 and 2043
%    pure Python functions and classes. 
% 3) Agnosticity. Pytracer works on any python libraries, with any data format
%    supported by pickle, and any numerical noise model flavor.

% Pytracer is the first tool to bring numerical analysis of Python codes
% available to everyone. Indeed, existing tools developed in the last years
% focused on traditional HPC languages community such as C,C++, and FORTRAN.
% However, the Python high-level abstraction allows you to quickly develop
% applications while maintaining high performance with the help of specialized
% libraries like NumPy. Therefore, it is quite natural to see more fields
% systematically using Python, as is the case in neuroimaging. Thus, Pytracer
% meets the numerical analysis needs of an ever increasing number of
% applications writing in Python.

% By working at function level, Pytracer allows users to precisely pinpoint
% numerical instabilities root causes. Furthermore, the automaticity facilitates
% the usage and its integration in unit testing. Hence, instead of tweaking
% hyperparameters to make solvers converging, the user has a practical way with
% Pytracer to quantify the numerical quality of its computations.

% As we showed in section~\ref{sec:impact_mca_modes}, using the Full MCA mode
% leads to runtime errors that have a considerable impact on the proper
% functioning of the execution, which is unfortunate since these errors do not
% reflect actual numerical errors but errors related to perturbations that
% should not occur. Fixing this misbehavior would greatly help use Full MCA mode
% since this mode allows simulating catastrophic cancellations, which are
% familiar sources of numerical errors. By preserving exact operations, RR mode
% is far more conservative and so easy to use even though it does not simulate
% all perturbations that might be encountered.

% Also, using a post-process analysis to aggregate traces, \pytracer falls in
% the same pitfalls as veritracer and is limited to aggregate traces where the
% execution flow path extremely diverges. Therefore, a posteriori reconstruction
% is challenging as one must precisely identify the flow path taken and find
% synchronization points to pursue the analysis. We retrieve, for example, this
% divergence behavior in two general classes of numerical schemes: iterative
% methods and optimization processes. One solution to overcome this issue is to
% integrate methods used in code coverage analysis to keep track of flow paths
% taken and the backtraces. However, this would increase the complexity of the
% postprocess analysis as well as an extra slowdown factor. However, by coupling
% the Gantt chart with static analysis, \pytracer would catch some divergence
% patterns occurring within a function. Indeed, by instrumenting at the function
% level, \pytracer views the function's callees (i.e., the functions called
% within a function) and thus should partially detect flow divergences.

% By design principle, \pytracer is flexible on the noise model used, although
% we only tested it with the MCA model in our experiments. Nevertheless, we
% would like to compare different models against MCA in future works to see the
% resulting differences. In particular, although \pytracer only works on a
% sequential application for the moment, it can be used to analyze parallel
% sections internal to a function. The only restriction is that the section does
% not call other functions traced by \pytracer; we would fall into the previous
% issue of divergence if it would not be the case. Since \pytracer only cares
% about inputs and outputs of a function, the order of internal operations is
% transparent for it and thus should be able to handle those cases.

% Finally, \pytracer does not instrument elementary arithmetic operations, and
%it can be a barrier to more profound analysis. Such an analysis would require a
%different instrumentation approach by, for example, working at the Abstract
%Syntax Tree (AST) level since it is trivial to get it in Python. An ideal way
%to do that would be to extract and replay a function, in the same vein as the
%CERE~\cite{castro2015cere} tool does but for Python, and use an AST
%instrumentation to analyze the numerical quality of elementary operations step
%by step within the fuzzy environment. Lastly, AST analysis can serve as a basis
%for finding relationships between variables, valuable information for the user.

% In addition to the runtime errors it raises, mismatching shapes is also a
% problem when aggregating traces as PyTracer will not compute statistics for a
% set of arrays with different shapes. Two solutions exist to overcome this
% issue: the first is not to instrument the \texttt{PyArray\_Arange} function
% with MCA. The advantage is that it is easy to set up since it requires one
% extra line in the exclusion file. However, the same kind of problem may appear
% somewhere else because, as we have seen, the operations being evolved are
% elementary. Therefore, the chances that developers reproduce the same pattern
% are very probable. The second solution is to ensure that the Full MCA mode
% preserves the exact operations. This solution has the advantage of not
% worrying about which functions should be instrumented or not with MCA, but it
% is hard to achieve it in practice. Indeed, we rely on the MCA implementation
% provided by Verificarlo that uses an internal precision to add the random
% perturbation that is higher than the working precision for applying the random
% noise. Verificarlo then rounds the result to the working precision before
% returning it, so any residual noise lower than the working precision is lost.
% Keeping track of that noise would require recording each intermediate
% computation in a shadow memory or keeping a tag about their exactness.
% \Yohan{This idea emerged by discussing with Verificarlo's team. I must find a
% way to express that and not give the impression that we discovered or suggest
% it.} This option would require more extra works than the first one but would
% ensure more consistent results.

\section{Conclusion}

% \tristan{I moved this from the opening of Section 4, as this reads more like a
% summary of \pytracer's features than an experiment description. It should be
% refactored in 1-2 paragraphs summarizing the main properties of \\pytracer.}

Evaluating the numerical stability of scientific Python programs is a tedious
task that requires substantial expertise. Automatic tools to help users
understand the behavior of their code are therefore crucial. In this
perspective, \pytracer is the first tool to offer a complete framework for
analyzing and visualizing the numerical stability of Python codes. It works by
automatically instrumenting Python libraries to produce numerical traces
visualized through an interactive Plotly Dash server. \pytracer does not require
manual code modification, and we tested it with major scientific libraries such
as NumPy, SciPy, and Scikit-learn. Our results show that \pytracer is scalable,
with an average instrumentation slowdown of $3.5\times$. Moreover, \pytracer's
visualization interface remains responsive even with traces of up to 1TB. In our
experiments, \pytracer enabled the systematic evaluation of 40 examples from the
SciPy and scikit-learn libraries.

In future work, we plan to extend the post-processing analysis to aggregate
traces when the execution flow path diverges by leveraging techniques used in
code coverage analysis. Moreover, we would like to extend \pytracer to
instrument elementary arithmetic operations which can be helpful to more
profound analyses. Finally, although \pytracer supports multiple noise models,
we only tested it with Monte Carlo Arithmetic. Therefore, we would like to
compare different models against MCA. \pytracer is available at
\href{https://github.com/big-data-lab-team/pytracer}{github://big-data-lab-team/pytracer}
under MIT License.

% % use section* for acknowledgment
% \ifCLASSOPTIONcompsoc
%     % The Computer Society usually uses the plural form
%     \section*{Acknowledgments}
% \else
%     % regular IEEE prefers the singular form
%     \section*{Acknowledgment}
% \fi

% The authors would like to thank...

% Can use something like this to put references on a page
% by themselves when using endfloat and the captionsoff option.
\ifCLASSOPTIONcaptionsoff
    \newpage
\fi

\bibliographystyle{IEEEtran}
\bibliography{paper-annotated}

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
%% \begin{thebibliography}{1}

%% \bibitem{IEEEhowto:kopka}
%% H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
%%   0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.

%% \end{thebibliography}

% biography section
% 
% If you have an EPS/PDF photo (graphicx package needed) extra braces are
% needed around the contents of the optional argument to biography to prevent
% the LaTeX parser from getting confused when it sees the complicated
% \includegraphics command within an optional argument. (You could create
% your own custom macro containing the \includegraphics command to make things
% simpler here.)
%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
% or if you just want to reserve a space for a photo:

\vskip -2\baselineskip plus -1fil
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{photos/yohan.jpg}}]{Yohan
        Chatelain} received his Ph.D. degree from the University of Paris-Saclay
    (UVSQ), Versailles, France in 2019. He is a postdoctoral researcher in the
    Big Data Infrastructures for Neuroinformatics lab at the University of
    Concordia, Montreal, Canada. His research topics include computer
    arithmetic, high-performance computing, and reproducibility. Yohan aims at
    democratizing the use of the analysis of stability for scientific computing
    codes through automatic tools to improve numerical quality. 
\end{IEEEbiography}

\vskip -2\baselineskip plus -1fil
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{photos/NigelPhoto.png}}]{Nigel
        Yong Sao Young} was born in Mauritius Island. He is a recent Computer
    Science graduate at Concordia University. He is currently working on his
    software startup, based in the city of Montreal.
\end{IEEEbiography}

%% % if you will not have a photo at all:
\vskip -2\baselineskip plus -1fil
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{photos/greg_kiar.png}}]{Gregory
        Kiar} is a research scientist at the Child Mind Institute. Throughout his
    degrees in Biomedical Engineering, Greg has developed techniques to study
    biosignal data, a turn-key tool that allows researchers to generate maps of
    brain connectivity from diffusion MRI data, and techniques to assess and
    improve the stability of research in neuroscience. Greg’s research bridges
    the fields of numerical analysis and brain imaging to evaluate and improve
    the trustworthiness of techniques used to study the brain. 
\end{IEEEbiography}

\vskip -2\baselineskip plus -1fil
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{photos/glatard_tristan.jpg}}]{Tristan
        Glatard} is Associate Professor in the Department of Computer Science and
    Software Engineering at Concordia University in Montreal, and Canada
    Research Chair (Tier II) on Big Data Infrastructures for Neuroinformatics.
    Before that, he was research scientist at the French National Centre for
    Scientific Research and Visiting Scholar at McGill University. 
\end{IEEEbiography}

%% % insert where needed to balance the two columns on the last page with
%% % biographies
%% %\newpage

%% \begin{IEEEbiographynophoto}{Jane Doe}
%% Biography text here.
%% \end{IEEEbiographynophoto}

% You can push biographies down or up by placing
% a \vfill before or after them. The appropriate
% use of \vfill depends on what kind of text is
% on the last page and whether or not the columns
% are being equalized.

%\vfill

% Can be used to pull up biographies so that the bottom of the last one
% is flush with the other column.
%\enlargethispage{-5in}

\appendix

\include{annex}

% that's all folks
\end{document}