### 2022-11-21 IEEEtran/paper.tex
Final edit before publication

%% bare_jrnl_compsoc.tex % V1.4b % 2015/08/26 % by Michael Shell % See: %
%http://www.michaelshell.org/ % for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls % (requires
%IEEEtran.cls version 1.8b or later) with an IEEE % Computer Society journal
%paper.
%%
%% Support sites: % http://www.michaelshell.org/tex/ieeetran/ %
%http://www.ctan.org/pkg/ieeetran % and % http://www.ieee.org/

%%*************************************************************************
%% Legal Notice: % This code is offered as-is without any warranty either
%expressed or % implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! % User assumes all risk. % In no event
%shall the IEEE or any contributor to this code be liable for % any damages or
%losses, including, but not limited to, incidental, % consequential, or any
%other damages, resulting from the use or misuse % of any information contained
%here.
%%
%% All comments are the opinions of their respective authors and are not %
%necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL) % (
%http://www.latex-project.org/ ) version 1.3, and may be freely used, %
%distributed and modified. A copy of the LPPL, version 1.3, is included % in the
%base LaTeX documentation of all distributions of LaTeX released % 2003/12/01 or
%later. % Retain all contribution notices and credits. % ** Modified files
%should be clearly indicated as such, including  ** % ** renaming them and
%changing author support contact information. **
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. The IEEE's font choices and paper sizes can   ***
% *** trigger bugs that do not appear when using other class files.       ***
% *** The testflow support page is at: http://www.michaelshell.org/tex/testflow/

\documentclass[10pt,journal,compsoc]{IEEEtran}

\usepackage{todonotes}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{color}
\usepackage{hyperref}
\usepackage{multirow}
\usepackage{pifont}
\usepackage{xcolor,colortbl}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=blue!50!black,citecolor=blue!75!black,filecolor=black,urlcolor=blue!75!black}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{xspace}
\usepackage{makecell}
\usepackage{ulem} % provides sout
\usepackage[frozencache=true,cachedir=minted-cache]{minted}
\usepackage[htt]{hyphenat}
\usepackage{subcaption}
\usepackage{graphicx,import}
\usepackage{booktabs}
\usepackage{soul}
%\usepackage{cancel}
%
% If IEEEtran.cls has not been installed into the LaTeX system files, manually
% specify the path to it like:
% \documentclass[10pt,journal,compsoc]{../sty/IEEEtran}

% Some very useful LaTeX packages include: (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf} Heiko Oberdiek's ifpdf.sty is very useful if you need
% conditional compilation based on whether the output is pdf or dvi. usage:
% \ifpdf % pdf code \else % dvi code \fi The latest version of ifpdf.sty can be
% obtained from: http://www.ctan.org/pkg/ifpdf Also, note that IEEEtran.cls V1.7
% and later provides a builtin \ifCLASSINFOpdf conditional that works the same
% way. When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
\ifCLASSOPTIONcompsoc
  % IEEE Computer Society needs nocompress option requires cite.sty v4.0 or
  % later (November 2003)
  \usepackage[nocompress]{cite}
\else
  % normal IEEE
  \usepackage{cite}
\fi
% cite.sty was written by Donald Arseneau V1.6 and later of IEEEtran pre-defines
% the format of the cite.sty package \cite{} output to follow that of the IEEE.
% Loading the cite package will result in citation numbers being automatically
% sorted and properly "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6]
% without using cite.sty will become [1], [2], [5]--[7], [9] using cite.sty.
% cite.sty's \cite will automatically add leading space, if needed. Use
% cite.sty's noadjust option (cite.sty V3.8 and later) if you want to turn this
% off such as if a citation ever needs to be enclosed in parenthesis. cite.sty
% is already installed on most LaTeX systems. Be sure and use version 5.0
% (2009-03-20) and later if using hyperref.sty. The latest version can be
% obtained at: http://www.ctan.org/pkg/cite The documentation is contained in
% the cite.sty file itself.
%
% Note that some packages require special options to format as the Computer
% Society requires. In particular, Computer Society  papers do not use
% compressed citation ranges as is done in typical IEEE papers (e.g., [1]-[4]).
% Instead, they list every citation separately in order (e.g., [1], [2], [3],
% [4]). To get the latter we need to load the cite package with the nocompress
% option which is supported by cite.sty v4.0 and later. Note also the use of a
% CLASSOPTION conditional provided by IEEEtran.cls V1.7 and later.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx} declare the path(s) where your graphic files
  % are \graphicspath{{../pdf/}{../jpeg/}} and their extensions so you won't
  % have to specify these with every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx will
  % default to the driver specified in the system graphics.cfg if no driver is
  % specified. \usepackage[dvips]{graphicx} declare the path(s) where your
  % graphic files are \graphicspath{{../eps/}} and their extensions so you won't
  % have to specify these with every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is required if
% you want graphics, photos, etc. graphicx.sty is already installed on most
% LaTeX systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/graphicx Another good source of documentation is
% "Using Imported Graphics in LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated postscript
% (.eps) format. pdflatex in pdf mode supports graphics in .pdf, .jpeg, .png and
% .mps (metapost) formats. Users should ensure that all non-photo figures use a
% vector format (.eps, .pdf, .mps) and not a bitmapped formats (.jpeg, .png).
% The IEEE frowns on bitmapped formats which can result in "jaggedy"/blurry
% rendering of lines and letters as well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
%\usepackage{amsmath} A popular package from the American Mathematical Society
% that provides many useful and powerful commands for dealing with mathematics.
%
% Note that the amsmath package sets \interdisplaylinepenalty to 10000 thus
% preventing page breaks from occurring within multiline equations. Use:
% \interdisplaylinepenalty=2500 after loading amsmath to restore such page
% breaks as IEEEtran.cls normally does. amsmath.sty is already installed on most
% LaTeX systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/amsmath

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic} algorithmic.sty was written by Peter Williams and
% Rogerio Brito. This package provides an algorithmic environment fo describing
% algorithms. You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide correct
% IEEE style captions. The latest version and documentation of algorithmic.sty
% can be obtained at: http://www.ctan.org/pkg/algorithms Also of interest may be
% the (relatively newer and more customizable) algorithmicx.sty package by Szasz
% Janos: http://www.ctan.org/pkg/algorithmicx

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array} Frank Mittelbach's and David Carlisle's array.sty patches
% and improves the standard LaTeX2e array and tabular environments to provide
% better appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with respect to
% the quality of the end results, all users are strongly advised to use an
% enhanced (at the very least that provided by array.sty) set of table tools.
% array.sty is already installed on most systems. The latest version and
% documentation can be obtained at: http://www.ctan.org/pkg/array

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES *** \ifCLASSOPTIONcompsoc
%\usepackage[caption=false,font=footnotesize,labelfont=sf,textfont=sf]{subfig}
%\else \usepackage[caption=false,font=footnotesize]{subfig} \fi subfig.sty,
%written by Steven Douglas Cochran, is the modern replacement for subfigure.sty,
%the latter of which is no longer maintained and is incompatible with some LaTeX
%packages including fixltx2e. However, subfig.sty requires and automatically
%loads Axel Sommerfeldt's caption.sty which will override IEEEtran.cls' handling
%of captions and this will result in non-IEEE style figure/table captions. To
%prevent this problem, be sure and invoke subfig.sty's "caption=false" package
%option (available since subfig.sty version 1.3, 2005/06/28) as this is will
%preserve IEEEtran.cls handling of captions. Note that the Computer Society
%format requires a sans serif font rather than the serif font used in
%traditional IEEE formatting and thus the need to invoke different subfig.sty
%package options depending on whether compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/pkg/subfig

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e} fixltx2e, the successor to the earlier fix2col.sty, was
% written by Frank Mittelbach and David Carlisle. This package corrects a few
% problems in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a single
% column figure to be placed prior to an earlier double column figure. Be aware
% that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's corrections
% already built into the system in which case a warning will be issued if an
% attempt is made to load fixltx2e.sty as it is no longer needed. The latest
% version and documentation can be found at: http://www.ctan.org/pkg/fixltx2e

%\usepackage{stfloats} stfloats.sty was written by Sigitas Tolusis. This package
% gives LaTeX2e the ability to do double column floats at the bottom of the page
% as well as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command: \fnbelowfloat to enable the placement of
% footnotes below bottom floats (the standard LaTeX2e kernel puts them above
% bottom floats). This is an invasive package which rewrites many portions of
% the LaTeX2e float routines. It may not work with other packages that modify
% the LaTeX2e float routines. The latest version and documentation can be
% obtained at: http://www.ctan.org/pkg/stfloats Do not use the stfloats
% baselinefloat ability as the IEEE does not allow \baselineskip to stretch.
% Authors submitting work to the IEEE should note that the IEEE rarely uses
% double column equations and that authors should try to avoid such use. Do not
% be tempted to use the cuted.sty or midfloat.sty packages (also by Sigitas
% Tolusis) as the IEEE does not format its papers in such ways. Do not attempt
% to use stfloats with fixltx2e as they are incompatible. Instead, use Morten
% Hogholm'a dblfloatfix which combines the features of both fixltx2e and
% stfloats:
%
% \usepackage{dblfloatfix} The latest version can be found at:
% http://www.ctan.org/pkg/dblfloatfix

%\ifCLASSOPTIONcaptionsoff \usepackage[nomarkers]{endfloat}
%  \let\MYoriglatexcaption\caption
%  \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}} \fi
%  endfloat.sty was written by James Darrell McCauley, Jeff Goldberg and Axel
%  Sommerfeldt. This package may be useful when used in conjunction with
%  IEEEtran.cls'  captionsoff option. Some IEEE journals/societies require that
%  submissions have lists of figures/tables at the end of the paper and that
%  figures/tables without any captions are placed on a page by themselves at the
%  end of the document. If needed, the draftcls IEEEtran class option or
%  \CLASSINPUTbaselinestretch interface can be used to increase the line spacing
%  as well. Be sure and use the nomarkers option of endfloat to prevent endfloat
%  from "marking" where the figures would have been placed in the text. The two
%  hack lines of code above are a slight modification of that suggested by in
%  the endfloat docs (section 8.4.1) to ensure that the full captions always
%  appear in the list of figures/tables - even if the user used the short
%  optional argument of \caption[]{}. IEEE papers do not typically make use of
%  \caption[]'s optional argument, so this should not be an issue. A similar
%  trick can be used to disable captions of packages such as subfig.sty that
%  lack options to turn off the subcaptions: For subfig.sty:
%  \let\MYorigsubfloat\subfloat
%  \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}} However, the above
%  trick will not work if both optional arguments of the \subfloat command are
%  used. Furthermore, there needs to be a description of each subfigure
%  *somewhere* and endfloat does not add subfigure captions to its list of
%  figures. Thus, the best approach is to avoid the use of subfigure captions
%  (many IEEE journals avoid them anyway) and instead reference/explain all the
%  subfigures within the main caption. The latest version of endfloat.sty and
%  its documentation can obtained at: http://www.ctan.org/pkg/endfloat
%
% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used later in
% the document, say, to conditionally put the References on a page by
% themselves.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url} url.sty was written by Donald Arseneau. It provides better
% support for handling and breaking URLs. url.sty is already installed on most
% LaTeX systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. *** *** Do
% not use packages that alter fonts (such as pslatex).         *** There should
% be no need to do such things with IEEEtran.cls V1.6 and later. (Unless
% specifically asked to do so by the journal or conference you plan to submit
% to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\newcommand{\cross}[0]{\cellcolor{red!65}\ding{53}}
\newcommand{\valid}[0]{\cellcolor{green!75!black}\ding{51}}
\newcommand{\warn}[0]{\cellcolor{orange!75}?}
\newcommand{\na}[0]{\cellcolor{gray!25}}
\newcommand{\s}[1]{\cellcolor{cyan!25}#1} \newcommand{\scross}[0]{\ding{53}~}
\newcommand{\svalid}[0]{\ding{51}~}
\newcommand{\swarn}[0]{?~}
\newcommand{\pytracer}[0]{PyTracer\xspace}

\definecolor{lightred}{rgb}{0.6,0.0,0.0}
\definecolor{lightgreen}{rgb}{0.8,1.0,0.8}

\usepackage{ifthen}
\newif\ifANNOTATED \ifthenelse{\equal{\detokenize{annotated}}{\jobname}}{
\ANNOTATEDtrue }{ \ANNOTATEDfalse }

\ifANNOTATED
\definecolor{ao(english)}{rgb}{0.0, 0.5, 0.0}
\DeclareRobustCommand{\add}[1]{\textcolor{ao(english)}{#1}}%\sethlcolor{lightgreen}\hl{#1}}
\DeclareRobustCommand{\remove}[1]{}
\DeclareRobustCommand{\addCite}[1]{\textcolor{lightgreen}{\mbox{#1}}}
\newcommand{\TG}[1]{\color{red}\textsc{From Tristan}: #1\xspace\color{black}}
\else
\DeclareRobustCommand{\add}[1]{#1} \DeclareRobustCommand{\remove}[1]{}
\DeclareRobustCommand{\addCite}[1]{#1}
\newcommand{\TG}[1]{}
\fi

\setlength{\arrayrulewidth}{0.1pt}

\include{href}

\makeatletter
\newcommand{\thickhline}{%
    \noalign {\ifnum 0=`}\fi \hrule height 1pt \futurelet \reserved@a \@xhline }
    \newcolumntype{"}{@{\hskip\tabcolsep\vrule width 1pt\hskip\tabcolsep}}
\makeatother

\lstdefinestyle{customPython}{ belowcaptionskip=1\baselineskip, breaklines=true,
  xleftmargin=\parindent, language=Python, showstringspaces=false,
  basicstyle=\scriptsize\ttfamily, keywordstyle=\bfseries\color[rgb]{0.580,
  0.000, 0.827},
  %{purple!40!lightgray},
  commentstyle=\itshape\color{green!40!black},
  identifierstyle=\bfseries\color{cyan!75!black}, stringstyle=\color{orange},
  deletekeywords={double,float}, classoffset=1, % starting new class
  otherkeywords={double,float}, morekeywords={double,float},
  keywordstyle=\bfseries\color{green!55!black}, classoffset=0 }

\lstdefinestyle{customC}{ belowcaptionskip=1\baselineskip, breaklines=true,
  xleftmargin=\parindent, language=C, showstringspaces=false,
  basicstyle=\scriptsize\ttfamily, keywordstyle=\bfseries\color[rgb]{0.580,
  0.000, 0.827},
  %{purple!40!lightgray},
  commentstyle=\itshape\color{green!40!black},
  identifierstyle=\bfseries\color{cyan!75!black}, stringstyle=\color{orange},
  deletekeywords={double,float}, classoffset=1, % starting new class
  otherkeywords={double,float}, morekeywords={double,float},
  keywordstyle=\bfseries\color{green!55!black}, classoffset=0 }

\begin{document}

\makeatletter
\let\orig@lstnumber=\thelstnumber
\newcommand\lstsetnumber[1]{\gdef\thelstnumber{#1}}
\newcommand\lstresetnumber{\global\let\thelstnumber=\orig@lstnumber}
\makeatother

%
% paper title Titles are generally capitalized except for words such as a, an,
% and, as, at, but, by, for, in, nor, of, on, or, the, to and up, which are
% usually not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired. Do not
% put math or special symbols in the title.
\title{PyTracer: Automatically Profiling Numerical Instabilities in Python}
%
%
% author names and IEEE memberships note positions of commas and nonbreaking
% spaces ( ~ ) LaTeX will not break a structure at a ~ so this keeps an author's
% name from being broken across two lines. use \thanks{} to gain access to the
% first footnote area a separate \thanks must be used for each paragraph as
% LaTeX2e's \thanks was not built to handle multiple paragraphs
%
%
%\IEEEcompsocitemizethanks is a special \thanks that produces the bulleted lists
% the Computer Society journals use for "first footnote" author affiliations.
% Use \IEEEcompsocthanksitem which works much like \item for each affiliation
% group. When not in compsoc mode, \IEEEcompsocitemizethanks becomes like
% \thanks and \IEEEcompsocthanksitem becomes a line break with idention. This
% facilitates dual compilation, although admittedly the differences in the
% desired content of \author between the different types of papers makes a
% one-size-fits-all approach a daunting prospect. For instance, compsoc journal
% papers have the author affiliations above the "Manuscript received ..."  text
% while in non-compsoc journals this is reversed. Sigh.

\author{Yohan Chatelain$^1$, Nigel Yong Sao Young$^1$,  Gregory Kiar$^2$,
    Tristan Glatard$^1$\\
    $^1$Department of Computer Science and Software Engineering, Concordia
    University, Montreal, Canada\\
    $^2$Center for the Developing Brain, Child Mind Institute, New York, NY,
    USA}
%% \author{Michael~Shell,~\IEEEmembership{Member,~IEEE,} %
%John~Doe,~\IEEEmembership{Fellow,~OSA,} %
%and~Jane~Doe,~\IEEEmembership{Life~Fellow,~IEEE}% <-this % stops a space %
%\IEEEcompsocitemizethanks{\IEEEcompsocthanksitem M. Shell was with the
%Department % of Electrical and Computer Engineering, Georgia Institute of
%Technology, Atlanta, % GA, 30332.\protect\\
%% % note need leading \protect in front of \\ to get a newline within \thanks
%as % % \\ is fragile and will error, could use \hfil\break instead. % E-mail:
%see http://www.michaelshell.org/contact.html % \IEEEcompsocthanksitem J. Doe
%and J. Doe are with Anonymous University.}% <-this % stops an unwanted space %
%\thanks{Manuscript received April 19, 2005; revised August 26, 2015.}}

% note the % following the last \IEEEmembership and also \thanks - these prevent
% an unwanted space from occurring between the last author name and the end of
% the author line. i.e., if you had this:
%
% \author{....lastname \thanks{...} \thanks{...} }
%                     ^------------^------------^----Do not want these spaces!
%
% a space would be appended to the last name and could cause every name on that
% line to be shifted left slightly. This is one of those "LaTeX things". For
% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get "AB"
% then you have to do: "\textbf{A}\textbf{B}" \thanks is no different in this
% regard, so shield the last } of each \thanks that ends a line with a % and do
% not let a space in before the next \thanks. Spaces after \IEEEmembership other
% than the last one are OK (and needed) as you are supposed to have spaces
% between the names. For what it is worth, this is a minor point as most people
% would not even notice if the said evil space somehow managed to creep in.

% The paper headers
\markboth{Journal of \LaTeX\ Class Files,~Vol.~14, No.~8, August~2015}%
{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Computer
    Society Journals}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
%
% *** Note that you probably will NOT want to include the author's *** *** name
% in the headers of peer review papers.                   *** You can use
% \ifCLASSOPTIONpeerreview for conditional compilation here if you desire.

% The publisher's ID mark at the bottom of the page is less important with
% Computer Society journal papers as those publications place the marks outside
% of the main text columns and, therefore, unlike regular IEEE journals, the
% available text space is not reduced by their presence. If you want to put a
% publisher's ID mark on the page you can do it like this:
% \IEEEpubid{0000--0000/00\$00.00~\copyright~2015 IEEE} or like this to get the
% Computer Society new two part style. \IEEEpubid{\makebox[\columnwidth]{\hfill
% 0000--0000/00/\$00.00~\copyright~2015 IEEE}%
% \hspace{\columnsep}\makebox[\columnwidth]{Published by the IEEE Computer
% Society\hfill}} Remember, if you use this you must call \IEEEpubidadjcol in
% the second column for its text to clear the IEEEpubid mark (Computer Society
% jorunal papers don't need this extra clearance.)

% use for special paper notices \IEEEspecialpapernotice{(Invited Paper)}

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran command as
% these need to go into the title area created by \maketitle. As a general rule,
% do not put math, special symbols or citations in the abstract or keywords.
\IEEEtitleabstractindextext{%
    \begin{abstract}
        Numerical stability is a crucial requirement of reliable scientific
        computing. However, despite the pervasiveness of Python in data science,
        analyzing large Python programs remains challenging due to the lack of
        scalable numerical analysis tools available for this language. To fill
        this gap, we developed \pytracer, a profiler to quantify numerical
        instability in Python applications. \pytracer transparently instruments
        Python code to produce numerical traces and visualize them interactively
        in a Plotly dashboard. We designed \pytracer to be agnostic to numerical
        noise model, allowing for numerical profiling through Monte-Carlo
        Arithmetic, random rounding, random data perturbation, or structured
        noise for a particular application. We illustrate \pytracer's
        capabilities by testing the numerical stability of key functions in both
        SciPy and Scikit-learn, two dominant Python libraries for mathematical
        modeling. Through these evaluations, we demonstrate \pytracer as a
        scalable, automated, and generic framework for numerical profiling in
        Python.
    \end{abstract}

    % Note that keywords are not normally used for peerreview papers.
    \begin{IEEEkeywords}
        Computer arithmetic, Quality analysis and evaluation, Visualization
        techniques and methodologies, Tracing,  Stochastic analysis
    \end{IEEEkeywords}
}

% make the title area
\maketitle

% To allow for easy dual compilation without having to reenter the
% abstract/keywords data, the \IEEEtitleabstractindextext text will not be used
% in maketitle, but will appear (i.e., to be "transported") here as
% \IEEEdisplaynontitleabstractindextext when the compsoc or transmag modes are
% not selected <OR> if conference mode is selected
% - because all conference papers position the abstract like regular papers do.
\IEEEdisplaynontitleabstractindextext
% \IEEEdisplaynontitleabstractindextext has no effect when using compsoc or
% transmag under a non-conference mode.

% For peer review papers, you can put extra information on the cover page as
% needed: \ifCLASSOPTIONpeerreview \begin{center} \bfseries EDICS Category:
% 3-BBND \end{center} \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and creates
% the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\IEEEraisesectionheading{\section{Introduction}\label{sec:introduction}}
% Computer Society journal (but not conference!) papers do something unusual
% with the very first section heading (almost always called "Introduction").
% They place it ABOVE the main text! IEEEtran.cls does not automatically do this
% for you, but you can achieve this effect with the provided
% \IEEEraisesectionheading{} command. Note the need to keep any \label that is
% to refer to the section immediately after \section in the above as
% \IEEEraisesectionheading puts \section within a raised box.

\IEEEPARstart{T}{he} scientific Python ecosystem has become a central component
of data analyses in recent years, owing to its rich offering of data
manipulation, array programming, numerical analysis, and visualization tools. In
particular, libraries such as NumPy~\cite{harris2020array},
SciPy~\cite{virtanen2020scipy}, scikit-learn~\cite{pedregosa2011scikit} or
PyTorch~\cite{paszke2019pytorch} are used in hundreds of publications each year,
and serve as a reference set of high-quality open-source core scientific
libraries. Researchers have built numerous domain-specific software tools from
this ecosystem, leveraging Python's simplicity and flexibility.

Numerical stability is a crucial requirement of reliable scientific data
processing. In unstable analyses, small numerical perturbations introduced by
data noise, software and hardware updates, or parallelization have the potential
to introduce substantial deviations in final results, leading to potentially
different scientific conclusions, ultimately threatening the reliability of
computational analyses. To prevent this, numerical stability analyses need to be
conducted systematically, however, no practical tool currently exists to conduct
such analyses for Python programs.

We present PyTracer, a numerical stability profiling and visualization tool for
Python programs. \pytracer adopts a dynamic approach that evaluates numerical
stability using program execution traces and applies to large programs without
manual intervention. The transparent nature of \pytracer is particularly
valuable in contrast to theoretical approaches based on condition numbers or
backward error analysis~\cite{higham2002accuracy}, which require detailed
modeling of the program and its numerical implementation. Similarly, static code
analysis techniques~\cite{goubault2006static} hardly scale to large codebases,
particularly in dynamically-typed languages such as Python.

\pytracer estimates the variability of floating-point variables by combining
program traces obtained with different numerical perturbations, which requires
(a) generating numerically-perturbed executions, (b) tracing floating-point
computations along with program executions, and (c) visualizing stability
evaluations. PyTracer addresses this work (a) by relying on
Fuzzy~\cite{kiar2020comparing}, an existing collection of scientific computing
libraries instrumented with Monte-Carlo arithmetic through the Verificarlo
tool~\cite{verificarlo}, (b) by dynamically instrumenting Python functions,
modules, and classes through meta-programming, and (c) by visualizing stability
measures throughout program executions with an interactive Plotly dashboard.
\mbox{\pytracer} is available at
\mbox{\href{https://github.com/big-data-lab-team/pytracer}{github.com/big-data-lab-team/pytracer}}
(MIT License).

This manuscript presents the design of \pytracer and highlights its potential in
various use cases. Through stability evaluations of the SciPy and scikit-learn
libraries, we demonstrate \pytracer as a practical solution to review the
numerical stability of extensive code bases.

\section{\pytracer design}

To detect numerical instabilities, \pytracer works in three main steps:
\begin{enumerate}
    \item Instruments Python programs and their module dependencies
          automatically, to save the input and output values of functions in a
          trace file.
    \item Executes the instrumented program several times using Monte-Carlo
          Arithmetic to inject perturbations in function outputs.
    \item Aggregates all trace files to compute statistics about the traced
          values, and visualize them in a GUI.
\end{enumerate}
\pytracer adopts the workflow shown in Figure~\ref{fig:workflow}. In Step
1, \pytracer uses monkey patching to replace the functions called by the
application with a wrapper that saves their arguments and returned values in a
trace file. In Step 2, it executes the application multiple times with
Monte-Carlo Arithmetic as implemented in the Fuzzy~\cite{kiar2020comparing}
environment. It then aggregates the resulting trace files to compute summary
statistics on traced variables. In Step 3, a Plotly dashboard provides
interactive visualizations highlighting unstable code sections.

The remainder of this section assumes some knowledge of Python concepts such as
classes, functions, modules, decorators, callables, and iterators. We refer
to~\cite{beazley2013python} for a detailed presentation of these concepts.

\begin{figure*}
    \centering
    \includegraphics[width=0.75\linewidth]{figure/workflow.pdf}
    \caption{Pytracer workflow
    }
    \label{fig:workflow}
\end{figure*}

\subsection{Python code instrumentation}

\pytracer dynamically instruments Python modules without requiring source code
modifications in the application. To do so, it creates a new, instrumented
version of each module that preserves its original attributes so that the
application can transparently use it.
By default, \pytracer instruments all the attributes in a given module except
the special ones of the form \texttt{\_\_*\_\_}, which is common for reserved
Python methods. However, some attributes are not writable. In this case,
\pytracer preserves the original attributes and warns the user. The user can
also restrict the list of traced attributes through an inclusion/exclusion
mechanism. The main advantage of this instrumentation approach is the lack of
required modification in the application source code, in contrast with  a
decorator-based approach~\mbox{\cite{hunt2019decorators}}. In addition, this
technique is scalable and applicable in read-only environments such as
Singularity containers or HPC clusters.

\subsubsection{Intercepting module import}

Python loads modules through an import mechanism triggered by the
\texttt{import} keyword that involves two objects: \texttt{finders} and
\texttt{loaders}. \texttt{Finders} search for packages (or
namespaces\footnote{Python distinguishes between \texttt{regular} and
    \texttt{namespace} packages. A regular package is a directory that contains an
    \_\_init\_\_.py file and potentially subdirectories (sub-packages) that contain
    themselves an \texttt{\_\_init\_\_.py} file and so on recursively. The package
    hierarchy follows those of the directory. Namespace packages introduced in
    Python 3.3 (\href{https://www.python.org/dev/peps/pep-0420/}{PEP 420}) do not
    contain an \_\_init\_\_.py file and allow for flexible directory structures.
    Hence, parts of the package can be located in zip files, on the network, or in
    separated directories. There are no functional differences between both types of
    packages.}) on the local storage --- such as standard packages installed through
the pip package manager --- or on a distant server. \texttt{Finders} do not load
modules. Instead, they return a specification (\texttt{spec} object) that
includes information on where to find and how to load modules. \texttt{Loaders}
create module objects, initialize them, and fill import-related attributes such
as \texttt{\_\_spec\_\_}; then execute modules and populate their namespace.
Finally, modules are bound to their import name in the \texttt{sys.modules}
dictionary.

Python supports custom \texttt{finder} classes registered in the
\texttt{sys.meta\_path} list. When Python encounters an \texttt{import}
statement, it first looks for an existing binding in \texttt{sys.modules} and
then iterates over the \texttt{finder} classes in the \texttt{sys.meta\_path}
list until it finds the module. \pytracer adds a custom \texttt{finder} class at
the head of \texttt{sys.meta\_path} that intercepts the module import and
creates the module with a custom \texttt{loader} class.

\pytracer's \texttt{loader} class first loads the original module and then
copies it as a new instance of the \texttt{ModuleType} class. It then calls the
appropriate instrumentation function for each attribute depending on its type
(function, class, or instances) and replaces the original module in the
\texttt{sys.modules} map with the instrumented version. Finally, it updates all
existing references to the original module contained in the global symbol table.
As a result, the application will transparently call the instrumented module.

\subsubsection{Instrumenting module functions}

For each function to be instrumented, \pytracer's \texttt{loader} class creates
a wrapper function, dynamically compiles it with the \texttt{compile} builtin
function, and substitutes it for the original module function.  Although simple,
this technique does not apply to callable class instances, i.e., class instances
that can be called directly with the () operator, because they are not of type
\texttt{FunctionType}. Indeed, applying the wrapper technique would 1) modify
the type of the class instance to \texttt{FunctionType}, which could cause
syntax and semantic errors, and 2) mask class attributes, leading to
\texttt{AttributeError} exceptions. \pytracer overrides the
\texttt{\_\_call\_\_} attribute with the wrapper function when possible to
overcome this issue. When the \texttt{\_\_call\_\_} attribute is not writable,
Pytracer does not instrument the class and returns a warning. This limitation is
expected to have a minor impact in practice since non-writable classes are
usually called from pure Python classes that can be instrumented. Therefore,
\pytracer reports information about the stability of non-writable classes
through the instrumentation of the pure Python classes that use them.

\subsubsection{Instrumenting classes}

\pytracer instruments classes by wrapping their callable attributes as described
previously. This instrumentation preserves class types, which avoids type
mismatches in the instrumented application. However, classes that describe
particular base types and NumPy's universal functions (class \texttt{ufunc})
contain read-only attributes that cannot be instrumented. By default, \pytracer
returns a warning when it encounters such classes.

\subsubsection{Instrumenting class instances}

\pytracer automatically instruments class instances since it instruments
classes. However, it cannot instrument or instantiate classes that have
read-only attributes --- in particular class \texttt{ufunc}, --- which is
problematic given their pervasiveness in scientific computing. In particular,
NumPy's universal functions include vectorized elementary mathematical and other
prevalent functions. Universal functions are implemented in C, which prevents
their modification.

To overcome this issue, \pytracer wraps the class instance into a transparent
wrap class (\texttt{twc}) that overloads 1) function
\texttt{\_\_getattribute\_\_}, called when an instance attribute is accessed ---
instrumented to return the attributes from the original class, and 2) function
\texttt{\_\_call\_\_}, used when an instance behaves like a function through the
() operator --- instrumented with function \texttt{generic\_wrapper}. The
\texttt{twc} returns the queried attribute of the instance instead of returning
its attributes, allowing transparent access to the instance.  Finally, \pytracer
overloads the builtin \texttt{type} function to preserve the type of the wrapped
instance.

\subsubsection{Instrumenting iterators}

In functional programming, iterators traverse containers lazily, meaning that
the next element in the sequence is only computed when the application uses it.
This technique allows for the manipulation of virtually infinite sequences with
finite memory. However, it implies that no complete mapping of the container
returned by an iterator exists in memory since the application computes each
element on the fly. Hence, iterators are not serializable which implies that
\pytracer cannot save them to output traces. A workaround would be to convert
iterators into explicit containers but this would increase the memory footprint
significantly. \pytracer, therefore, does not instrument iterators.

We do not expect this limitation to have a critical impact on the applicability
of \pytracer. Iterators generally involve simple computations that are unlikely
to be unstable. Besides, we encountered only a few iterators in our experiments
and none of them involved floating-point values. Should instabilities occur in
iterators, \pytracer would detect them through the inputs of the functions
consuming these iterators.

\subsection{Detecting numerical instabilities}

\pytracer detects numerical instabilities by computing summary statistics across
multiple executions perturbed with numerical noise. While \pytracer's
instrumentation, trace aggregation, and visualization work with various types of
numerical noise, we experimented primarily with Monte-Carlo Arithmetic (MCA) as
it is an accurate model for floating-point errors~\cite{verificarlo}.

\subsubsection{Noise injection}
\label{sec:fuzzy}

Three  main  approaches  exist  to  detect  numerical  instabilities: stochastic
arithmetic,  uncertainty  or  sensitivity analysis~\cite{helton2006survey}, and
random seed analysis~\cite{hellekalek1998good}. PyTracer is agnostic to the
noise model used and can therefore be used for all these methods.  For our use
cases, we focus on stochastic arithmetic since it does not make any assumption
about the noise model, in contrast to sensitivity analysis.

Stochastic arithmetic leverages randomness to estimate numerical instabilities
coming from the use of floating-point representations. The main idea is to treat
round-off errors as random variables and characterize them statistically. Monte
Carlo Arithmetic (MCA)~\cite{parker1997monte} uses this principle by introducing
two improvements: (i) a virtual precision parameter, allowing for simulating
reduced working precision, and (ii) different perturbation modes to introduce
perturbations on function inputs or output. MCA replaces each floating-point
number $x$ with a stochastic counterpart computed as:
\[
    inexact(x) =  x + \beta^{e_x - t}\xi
\]
where $\beta$ is the number base, $e_x$ is the magnitude of $x$, $t$ the virtual
precision and $\xi\! \in\! (-\frac{1}{2},\frac{1}{2})$ is a random uniform
variable. Virtual precision allows for the simulation of reduced working
precision. In our experiments, we used MCA in Random Rounding (RR) mode, which
simulates roundoff errors by applying function $inexact$ only to the output
returned by a function. The Precision Bounding (PB) mode
that applies the $inexact$ function on inputs
can identify other types of instabilities, in particular catastrophic
cancellations, however, it is more difficult to apply to large code bases as it
often leads to execution errors that require human intervention.

Stochastic arithmetic quantifies numerical error using the number of significant
digits $s$ estimated among the sampled values. A common formula to determine
this number from MCA samples is presented in~\cite{parker1997monte}:
\begin{equation}
    s = -\log_{\beta}{ \left| \dfrac{\sigma}{\mu} \right|} \label{eq:sig-digits}
\end{equation}
where $\mu$ and $\sigma$ are the sample mean and standard deviation of a
variable sampled with MCA. This can be seen as the ratio of signal over noise.

We enabled MCA in Python programs via Verificarlo~\cite{verificarlo} a
clang-based compiler~\cite{lattner2008llvm} that replaces floating-point
operations with a generic call to a configurable floating-point model. Several
floating-point models are
available~\cite{chatelain2019automatic,chatelain2019outils}, also called
backends. \pytracer leverages Fuzzy~\cite{kiar2020comparing}, a collection of
software tools compiled with Verificarlo. In particular, Fuzzy provides
MCA-instrumented versions of the Python interpreter as well as the BLAS, LAPACK,
NumPy, SciPy, and scikit-learn libraries, which enables MCA for a wide range of
existing scientific Python programs. Fuzzy is available in Verificarlo's GitHub
organization at
\href{https://github.com/verificarlo/fuzzy}{\url{github.com/verificarlo/fuzzy}}.

Monte Carlo Arithmetic samples the floating-point variability of a given
program, and hence the more MCA repetitions we have, the more accurate the
distribution will be. In theory, an infinite number of MCA repetitions should
give the exact distribution. While in practice low sample sizes (3 to 30) can
give good estimates, the issue is to capture ``rare events" making the
computation to strongly diverge from the mean. Determining the number of MCA
samples to use is still an open research question. Recent theoretical
works~\cite{sohier2021confidence} on the estimation of significant digits allow
computing an associated confidence interval depending on the sample size and the
underlying distribution.

\subsubsection{Trace format}

\pytracer stores trace in the pickle format, a binary format to serialize Python
objects. The main advantages of the pickle format compared to other
serialization approaches such as marshal or JSON are its ability to serialize
most Python objects and its native compression.  In addition, pickle writes
Python objects sequentially, which preserves the temporal ordering required by
subsequent trace analyses.

Traces store numerical values at the granularity of function calls. When the
application invokes a function, \pytracer saves a call input object containing
contextual information (function's name, module's function, stacktrace) and the
values of the function arguments. When the function returns, \pytracer saves a
similar call output object containing the returned value.

\subsubsection{Trace aggregation}

Once traces are available, \pytracer sequentially parses them and computes the
mean, standard deviation, and the number of significant digits (as in
Equation~\ref{eq:sig-digits}) of all saved values. \pytracer saves these summary
statistics in an HDF5 file, a hierarchical format that facilitates visual
browsing since only required information is loaded when requested by the user.
This operation critically relies on the ordering of call input and output
objects in the traces. It assumes that the application is single-threaded and
that the control flow does not depend on a random state. Therefore,
\mbox{\pytracer} cannot merge traces resulting from different control flow paths
and will stop the parsing when encountering such a situation.

\begin{figure*}
    \centering
    \includegraphics[width=0.75\linewidth]{figure/pytracer_layout.pdf}
    \caption{Pytracer visualization layout.}
    \label{fig:visu-layout}
\end{figure*}

To aggregate traces, \pytracer successively unpickles call objects in each
trace, checks that the call objects refer to the same function and stacktrace,
and terminates the parsing otherwise. PyTracer stores function arguments or
returned values in a NumPy array for each matching call object to compute
statistics. For compound objects such as tuples, dictionaries, or objects, it
recursively parses the fields until finding numeric types or NumPy arrays of
numeric types.

\subsection{Interactive visualization}
The third main component of \pytracer is its interactive trace visualizer built
with Plotly Dash~\cite{plotly}, a Python framework for web visualization. Plotly
Dash abstracts the low-level details of web applications in a highly
customizable development interface allowing to offload heavy computations to a
server if necessary. Figure~\ref{fig:visu-layout} presents the general layout of
the \pytracer visualizer, divided into five sub-components:
\begin{enumerate}
    \item Timeline view: displays the values of each input and output for a
          given function at a given invocation. By default, \pytracer shows the
          number of significant digits computed across the traces. The mean and
          standard deviation across perturbed samples are also available. The
          visualizer uses an upper triangle for inputs and a lower triangle for
          outputs. \pytracer internally represents all values as NumPy arrays.
          In the case of non-scalar values or types that cannot be trivially
          cast into NumPy arrays, \pytracer splits the value into several
          variables. Hence, \pytracer will transform a class containing $n$
          floating-point parameters into $n$ variables. Each variable is
          prefixed with the original value and postfixed with the name of the
          parameter. Finally, non-scalar values (vectors or matrices) are
          represented by their average and can be inspected with the Zoom view
          (see below).
    \item Gantt chart view: displays the call graph as a Gantt chart
          representing the instrumented functions. Each function call is
          represented with a bar. The time overlaps reflect the caller/callee
          relation. For example, in the figure, \texttt{function1} calls
          \texttt{function2} which itself calls \texttt{function4} and
          \texttt{function5}. This view provides insights into the calling
          context for each function that may influence numerical stability.
    \item  Source code view: displays a local snapshot of the source code file
          of the variable hovered in the timeline view.
    \item Zoom view: provides details about non-scalar values. For example, a
          matrix may have instabilities around its diagonal but not in other
          regions. The visualizer also provides additional metrics such as array
          shape, infinite norm, or condition number.
    \item  Functions list view: allows the user to select the functions plotted
          in the timeline view.
\end{enumerate}

\section{Results}

We evaluated \pytracer on two widely-used Python libraries:
SciPy~\cite{virtanen2020scipy}, the main library for scientific computing in
Python, and  scikit-learn~\cite{pedregosa2011scikit}, a reference
machine-learning framework that implements a wide range of state-of-the-art
algorithms. For each library, we executed the software examples provided in
their respective GitHub repositories. We traced each example five times, having
activated MCA through Fuzzy in the Python interpreter as well as BLAS, LAPACK,
NumPy, SciPy, and scikit-learn. These libraries are written in large parts in a
compiled language other than Python (such as C or Cython) and their execution is
not perturbed by the sole use of Fuzzy Python. We chose to activate MCA in these
libraries to provide an overall picture of numerical uncertainty in the tested
examples. Depending on the use case of interest, one might choose to use MCA
only in a subset of these libraries or even just in the Python interpreter. We
used MCA in Random Rounding (RR) mode, with the default virtual precision in
Verificarlo, namely 24 bits for single-precision floating-point values and 53
bits for double-precision values, which simulates machine error. We used
Python-3.8.5, NumPy-1.19.1, SciPy-1.5.4, and scikit-learn-0.23.2.

\begin{table*}[]
    \centering
    \normalsize
    \begin{subfigure}[t]{\linewidth}
        \centering
        \begin{tabular}{|lll|c|c|c|}
            \hline
            \multicolumn{3}{|c}{ \multirow{2}{*}{Application} }
             & \multicolumn{3}{|c|}{Random Rounding (RR)}

            \\
            \cline{4-6}
             & 
             & 
             & 
            trace
             & 
            parse
             & 
            sigbits
            \\
            \hline
            \multicolumn{1}{|c|}{ \multirow{20}{2em}{ \rotatebox{90}{SciPy} } }
             & \multicolumn{1}{c|}{ \multirow{3}{2em}{ \rotatebox{90}{FFT} }}

             & \discreteCosineRF
             & \valid                                                         & \valid
             & \s{52}                                                                  \\
            \multicolumn{1}{|c|}{}
             & \multicolumn{1}{c|}{}

             & \fftOneDRf                                                     & \valid
             & \valid                                                         & 41     \\
            \multicolumn{1}{|c|}{}
             & \multicolumn{1}{c|}{}

             & \fftTwoDRf                                                     & \valid
             & \valid                                                         & \s{52} \\
            \cline{2-6}
            \multicolumn{1}{|c|}{}
             & \multicolumn{1}{c|}{  \multirow{5}{*}{

                    \rotatebox{90}{Interpolation} } }

             & 
            \interOneDRf
             & 
            \valid
             & 
            \valid
             & 
            51
            \\
            \multicolumn{1}{|c|}{}
             & \multicolumn{1}{c|}{}

             & \multiRf                                                       & \valid
             & \valid                                                         & \s{51} \\
            \multicolumn{1}{|c|}{}
             & \multicolumn{1}{c|}{}

             & \bsplineRf                                                     & \valid
             & \valid                                                         & 10     \\
            \multicolumn{1}{|c|}{}
             & \multicolumn{1}{c|}{}

             & \splineOneDRf                                                  & \valid
             & \cross                                                         & \s{52} \\
            \multicolumn{1}{|c|}{}
             & \multicolumn{1}{c|}{}

             & \splineTwoDRf                                                  & \valid
             & \valid                                                         & 10     \\
            \cline{2-6}
            \multicolumn{1}{|c|}{}
             & \multicolumn{1}{c|}{ \multirow{12}{2em}{

                    \rotatebox{90}{Optimization} } }

             & 
            \bfgsRf
             & 
            \valid
             & 
            \valid
             & 
            \s{48}
            \\
            \cline{3-6}
            \multicolumn{1}{|c|}{}
             & \multicolumn{1}{c|}{}

             & \globalRf                                                      & \valid
             & \cross                                                         & 17     \\
            \multicolumn{1}{|c|}{}
             & \multicolumn{1}{c|}{}

             & | SHGO\footnote{Simplicial Homology Global Optimization}       & \na
             & \na                                                            & \s{25} \\
            \multicolumn{1}{|c|}{}
             & \multicolumn{1}{c|}{}

             & | Dual Annealing                                               & \na
             & \na                                                            & 4      \\
            \multicolumn{1}{|c|}{}
             & \multicolumn{1}{c|}{}

             & | Differential Evolution                                       & \na
             & \na                                                            & \s{0}  \\
            \multicolumn{1}{|c|}{}
             & \multicolumn{1}{c|}{}

             & | Basin Hopping                                                & \na
             & \na                                                            & 40     \\
            \multicolumn{1}{|c|}{}
             & \multicolumn{1}{c|}{}

             & | SHGO Sobol                                                   & \na
             & \na                                                            & \s{45} \\
            \cline{3-6}
            \multicolumn{1}{|c|}{}
             & \multicolumn{1}{c|}{}

             & \slsqpRf                                                       & \valid
             & \valid                                                         & 45     \\
            \multicolumn{1}{|c|}{}
             & \multicolumn{1}{c|}{}

             & \lsmRf                                                         & \valid
             & \valid                                                         & \s{48} \\
            \multicolumn{1}{|c|}{}
             & \multicolumn{1}{c|}{}

             & \nelderRf                                                      & \valid
             & \valid                                                         & 44     \\
            \multicolumn{1}{|c|}{}
             & \multicolumn{1}{c|}{}

             & \ncgRf                                                         & \valid
             & \valid                                                         & \s{48} \\
            \multicolumn{1}{|c|}{}
             & \multicolumn{1}{c|}{}

             & \rootRf                                                        & \valid
             & \cross                                                         & 52     \\
            \multicolumn{1}{|c|}{}
             & \multicolumn{1}{c|}{}

             & \rootLargRf                                                    & \valid
             & \cross                                                         & \s{29} \\
            \multicolumn{1}{|c|}{}
             & \multicolumn{1}{c|}{}

             & \rootLargePredRf                                               & \valid
             & \cross                                                         & 42     \\
            \multicolumn{1}{|c|}{}
             & \multicolumn{1}{c|}{}

             & \trustExactRf                                                  & \valid
             & \valid                                                         & \s{48} \\
            \multicolumn{1}{|c|}{}
             & \multicolumn{1}{c|}{}

             & \trustTruncRf                                                  & \valid
             & \valid                                                         & 48     \\
            \multicolumn{1}{|c|}{}
             & \multicolumn{1}{c|}{}

             & \trustNCGRf                                                    & \valid
             & \valid                                                         & \s{52} \\
            \hline
            \multicolumn{2}{|c|}{ \multirow{20}{2em}{
                    \rotatebox{90}{Scikit-learn}
                }
            }
             & 
            \AdaboostRf
             & 
            \valid
             & 
            \valid
             & 
            48
            \\
            \cline{3-6}
            \multicolumn{2}{|c|}{}
             & \brrRf
             & \valid                                                         & \valid
             & \s{25}                                                                  \\
            \multicolumn{2}{|c|}{}
             & | $1^{st}$ dataset
             & \na                                                            & \na
             & 25                                                                      \\
            \multicolumn{2}{|c|}{}
             & | $2^{nd}$ dataset
             & \na                                                            & \na
             & \s{nan}                                                                 \\
            \cline{3-6}
            \multicolumn{2}{|c|}{}
             & \onlineClassifierComparisonRf
             & \valid                                                         & \valid
             & 20                                                                      \\
            \multicolumn{2}{|c|}{}
             & \kmeansRf
             & \valid                                                         & \valid
             & \s{52}                                                                  \\
            \multicolumn{2}{|c|}{}
             & \covarianceRf
             & \valid                                                         & \cross
             & 48                                                                      \\
            \multicolumn{2}{|c|}{}
             & \decisionRf
             & \valid                                                         & \valid
             & \s{50}                                                                  \\
            \multicolumn{2}{|c|}{}
             & \digitsRf
             & \valid                                                         & \valid
             & 52                                                                      \\
            \multicolumn{2}{|c|}{}
             & \faceRf
             & \cross                                                         & \valid
             & \s{-}                                                                   \\
            \multicolumn{2}{|c|}{}
             & \penaltyRf
             & \valid                                                         & \valid
             & 52                                                                      \\
            \multicolumn{2}{|c|}{}
             & \lassoRf
             & \valid                                                         & \valid
             & \s{48}                                                                  \\
            \multicolumn{2}{|c|}{}
             & \hyperplaneRf
             & \valid                                                         & \valid
             & 3                                                                       \\
            \multicolumn{2}{|c|}{}
             & \mnistRf
             & \valid                                                         & \valid
             & \s{43}                                                                  \\
            \multicolumn{2}{|c|}{}
             & \multitaskRf
             & \valid                                                         & \valid
             & 50                                                                      \\
            \multicolumn{2}{|c|}{}
             & \ompRf
             & \valid                                                         & \valid
             & \s{52}                                                                  \\
            \multicolumn{2}{|c|}{}
             & \pcaRf
             & \valid                                                         & \valid
             & 48                                                                      \\
            \cline{3-6}
            \multicolumn{2}{|c|}{}
             & \robustRf
             & \valid                                                         & \valid
             & \s{50}                                                                  \\
            \multicolumn{2}{|c|}{}
             & | Linear
             & \na                                                            & \na
             & 48                                                                      \\
            \multicolumn{2}{|c|}{}
             & | RANSAC
             & \na                                                            & \na
             & \s{50}                                                                  \\
            \cline{3-6}
            \multicolumn{2}{|c|}{}
             & \toyRf
             & \valid                                                         & \valid
             & 35                                                                      \\
            \cline{3-6}
            \multicolumn{2}{|c|}{}
             & \svmRf
             & \valid                                                         & \valid
             & \s{43.5}                                                                \\
            \multicolumn{2}{|c|}{}
             & | Non-weighted
             & \na                                                            & \na
             & 43                                                                      \\
            \multicolumn{2}{|c|}{}
             & | Weighted
             & \na                                                            & \na
             & \s{44}                                                                  \\
            \cline{3-6}
            \multicolumn{2}{|c|}{}
             & \tomographyRf
             & \valid                                                         & \cross
             & -                                                                       \\
            \multicolumn{2}{|c|}{}
             & \weightedRf
             & \valid                                                         & \valid
             & \s{44}                                                                  \\
            \hline
        \end{tabular}
    \end{subfigure}
    \caption{\pytracer execution summary on SciPy and Scikit-learn examples.
        \emph{trace}: outcome of \pytracer tracing, \emph{parse}: outcome of
        \pytracer trace parsing and aggregation, \emph{result}: result precision
        in number of significant bits. \valid: successful run, \cross: numerical
        error (tracing) or different control flow paths (parsing). \textit{nan}
        stands for (\texttt{Not-A-Number}) meaning that the output result
        comes from an invalid arithmetic operation. Example names are linked to
        their source code.}
    \label{tab:pytracer_results_summary}
\end{table*}

\subsection{\pytracer is applicable to reference scientific computing libraries}

Table~\ref{tab:pytracer_results_summary} summarizes \pytracer's outcome on SciPy
and scikit-learn examples. Columns \textit{trace} and \textit{parse} report on
the tracing and aggregation steps with two possible outcomes: green
\valid\xspace for success and red \cross\xspace for error. Tracing errors occur
in the application as a result of numerical perturbations, whereas parsing
errors occur when traces result from different control flow paths (CFP).
Therefore, an application that consistently fails at the same function call
will produce traces with equivalent Control Flow Paths which can be parsed by
\pytracer. In our experiments, this was the case for the Face Recognition
shows.

Overall, the tracing and parsing of SciPy and scikit-learn examples in Random
Rounding mode was entirely successful for 32/40 examples.

Column \textit{sigbits} in Table~\ref{tab:pytracer_results_summary} summarizes
the result precision for the main outputs of SciPy and scikit-learn examples. We
use the term ``result precision" to denote the statistical variability measured
in the number of significant bits, in contrast with the other meanings of the
word precision used in the expressions ``virtual precision" or ``single/double
precision'' in this paper. We measured the result precision in significant bits
on the final primary example output by taking the element-wise mean for
non-scalar values and, if the example produced multiple outputs, the average
value among them. Significant bits were computed on the final program output and
could therefore be computed even when parsing failed. The numerical quality
across all examples varied from 0 to 52, with an average of 44 significant bits.
We note that only one example failed due to a numerical error
(\texttt{face\_recognition} for RR and \texttt{Bayesian Ridge Regression} for
full MCA).

\subsection{\pytracer supports the profiling of compound data types}

PyTracer supports native floating-point formats as well as compound data types,
such as complex floating-point values, N-dimensional arrays, and arbitrary
Python classes. Figures~\ref{fig:fft1D_inputs} and~\ref{fig:fft1D_outputs} show
the visualization of the \texttt{fft\_1D} input and output values that are
complex values encoded in double precision. PyTracer allows visualizing the
result precision of real (blue points) and imaginary (orange points) parts. The
top row in Figures~\ref{fig:fft1D_inputs} and~\ref{fig:fft1D_outputs} show the
input and output mean value while the bottom row shows the standard
deviation. Even though Random Rounding (RR) was used, function inputs were
perturbed since they result from multiple intermediate computations perturbed
with RR. The points with low magnitude in Figure~\ref{fig:fft1D_inputs}
correspond to inputs near 0 when the input of $\sin$ or $\exp$ is close to
$k\pi$, $k \in \mathbb{Z}$. Indeed, since MCA introduces noise, the result is
not exactly 0. We can see in Figure~\ref{fig:fft1D_inputs} that the maximal
noise introduced by MCA on the inputs is in the order of $10^{-14}$, two orders
of magnitude higher than the $ulp \simeq 10^-16$ for double precision. This
slight noise introduced can be interpreted as white noise on the input signal.
We can see on the bottom row of Figure~\ref{fig:fft1D_outputs} that the
frequency noise is of the same order of magnitude as the input noise, which is
expected. The two peaks at x=38 and x=562 correspond to the actual frequencies
of the input signal.

Figure~\ref{fig:spline2d_rr} shows the results of the spline evaluation with
\texttt{bisplev} from the \texttt{spline\_2D} example. It computes the 2D spline
interpolation for function $z=(x+y)e^{-6(x^2+y^2)}$, with $(x,y) \in
    [-1,1]\times[-1,1]$. The result is a 2D array that is natively supported by
PyTracer. We observe a significant loss of result precision in some regions,
which follows an interesting pattern that could be further investigated.
This example shows that \pytracer works on 2D inputs and the importance of
visualizing complex data types in addition to summary statistics.

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figure/FFT/fft_x.pdf}
    \caption{Absolute value of the mean and standard deviation values for fft 1D
        inputs within RR mode (log scale). The real part is shown in blue and
        the imaginary part in orange.}
    \label{fig:fft1D_inputs}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figure/FFT/fft_y.pdf}
    \caption{Absolute value of the mean and standard deviation values for fft 1D
        outputs within RR mode (log scale).}
    \label{fig:fft1D_outputs}
\end{figure}

\begin{figure*}
    \centering
    \begin{subfigure}{.3\linewidth}
        \centering
        \includegraphics[width=\linewidth]{figure/spline_2d/bisplev_mean.pdf}
        \caption{Mean value}
        \label{fig:bisplev_mean}
    \end{subfigure}
    \begin{subfigure}{.3\linewidth}
        \centering
        \includegraphics[width=\linewidth]{figure/spline_2d/bisplev_mean_log.pdf}
        \caption{Absolute mean value (log)}
        \label{fig:bisplev_mean_log}
    \end{subfigure}
    \begin{subfigure}{.3\linewidth}
        \centering
        \includegraphics[width=\linewidth]{figure/spline_2d/bisplev_std_log.pdf}
        \caption{Standard deviation (log)}
        \label{fig:bisplev_std_log}
    \end{subfigure}
    \caption{2D Spline interpolation results. Figure~\ref{fig:bisplev_mean}
        shows the mean result of the 2D-spline interpolation.
        Figures~\ref{fig:bisplev_mean_log} and~\ref{fig:bisplev_std_log} show
        respectively in the logarithmic scale the absolute mean value and the
        standard deviation of the 5 samples run within RR mode. }
    \label{fig:spline2d_rr}
\end{figure*}

\subsection{\pytracer allows visualizing numerical stability at a fine-grained level}

We illustrate on two examples how PyTracer is effective for analyzing numerical
stability in complex codes at a fine-grained level: Face recognition and
Bayesian Ridge Regression. The face recognition example uses eigenfaces
(eigenvectors) and Support Vector Machines (SVMs) to classify faces from the
Labeled Faces in the Wild dataset. This example uses a Principal Component
Analysis (PCA) to extract the eigenfaces and then trains an SVM classifier to
predict the label associated with a face. When executing in PyTracer, this
example crashed because the Singular Value Decomposition (SVD) did not converge.
This example is the only one to raise an error due to the non-convergence of the
SVD performed in the PCA. The PCA uses the Randomized
SVD~\cite{halko2011finding} method that first computes a low-rank approximation
of the input matrix using a stochastic algorithm and then uses this
approximation as input for the SVD. When walking through the code with PyTracer,
we observed that the \texttt{randomized\_svd} scikit-learn function uses the
Divide \& Conquer SVD method \texttt{gesdd}. The LAPACK library has two main
methods for computing the SVD: \texttt{gesdd} and \texttt{gesvd}. The former
uses a Divide \& Conquer approach, while the latter uses a QR decomposition.
While both methods are expected to have the same
accuracy~\cite{nakatsukasa2013stable}, the Divide \& Conquer does not converge
with MCA.

Replacing \texttt{gesdd} with \texttt{gesvd} solved the SVD convergence
issue. \pytracer allows us to quickly visualize the new result precision.
Figure~\ref{fig:face_recognition_svd} shows the result precision of the
eigenvalues computed with \texttt{gesvd} in single precision. The result
precision varies between 15 and 20 significant bits out of the 24 available in
single precision. The classifier metrics (precision, recall, F1-score, and
support) are identical to the IEEE execution for all MCA samples. We can note
that even with the \texttt{gesvd} method, the result precision of the singular
values and eigenvectors is not uniform. Indeed, we can see that the result
precision of the lowest singular values and eigenvectors decreases until
reaching 0 significant bits.

The second example compares a Bayesian Ridge Regression (BRR) to the Ordinary
Least Squares (OLS) estimator on a synthetic dataset and for one-dimensional
regression using polynomial feature expansion. Although the example does not
raise a runtime error, the regression coefficients computed from the fitting are
not significant. \pytracer traces reveal that the SVD solver used is once again
the root cause for this error. Indeed, the BayesianRidge model first computes an
SVD before iteratively updating the coefficients until reaching the convergence
criterion. However, PyTracer shows an issue in the singular values computed. We
can see in Figure~\ref{fig:BRR_svd} that the last singular value is \texttt{NaN}
due to numerical errors in the SVD, although the remaining singular values
exhibit high result precision (between 44 and 50 bits). Surprisingly, this
\texttt{NaN} value does not raise an exception and is silently propagated.
Indeed, the stopping criterion is never verified and the optimization stops
after reaching the maximal number of iterations. Figure~\ref{fig:BRR_svd_fixed}
shows that using the \texttt{gesvd} instead of the \texttt{gesdd} method removes
the \texttt{NaN} value although the last singular values still have a low result
precision (below 0). According to Equation~\ref{eq:sig-digits}, a number of
significant bits less than 0 means that the variability ($\sigma$) is higher
than the average value ($\mu$).

Similarly, for OLS, PyTracer traces show that the regression coefficients fitted
are \texttt{NaN} values. By looking at the callgraph, we can see that the scipy
calls the \texttt{lstsq} method from the \texttt{scipy.linalg} package which
uses the \texttt{gesdd} LAPACK function. Figure~\ref{fig:OLS_svd} shows that one
singular value is a \texttt{NaN} value, which corrupted the calculation of the
coefficients. When using the \texttt{gesvd} method, the \texttt{NaN} value
disappears (Fig.~\ref{fig:OLS_svd_fixed}), resulting in a result precision of 42
bits on average in the coefficients. We can see that having access to all the
values computed by the application with PyTracer allows the user to efficiently
identify numerical issues.

In conclusion, \pytracer and the MCA model were helpful to detect the
instability of the \texttt{gesdd} method mentioned in several GitHub issues such
as
\href{https://github.com/google/TensorNetwork/pull/962}{github.com/google/TensorNetwork/pull/962}
or
\href{https://github.com/scipy/scipy/issues/15024}{github.com/scipy/scipy/issues/15024}
including the LAPACK repository. Furthermore, it is worth mentioning that a Pull
Request at
\href{https://github.com/scikit-learn/scikit-learn/pull/20617}{github.com/scikit-learn/scikit-learn/pull/20617}
has recently been submitted in the scikit-learn repository to let the user
select the SVD method and avoid this kind of issue. Thus, \pytracer can detect
instabilities that affect fundamental scientific computing components such as
the LAPACK library. The unified view of all traces allows the user to quickly
navigate through the result precision of different variables.

\begin{figure*}
    \centering
    \begin{subfigure}{0.3\linewidth}
        \includegraphics[width=\linewidth]{figure/face_recognition/randomized_svd_ret_U_sig.pdf}
        \caption{Left matrix $U$ of the SVD decomposition}
        \label{fig:randomized_svd_U}
    \end{subfigure}
    \begin{subfigure}{0.3\linewidth}
        \includegraphics[width=\linewidth]{figure/face_recognition/svd_ret_S_sig.pdf}
        \caption{Singular values $\Sigma$ of the SVD decomposition}
        \label{fig:randomized_svd_S}
    \end{subfigure}
    \begin{subfigure}{0.3\linewidth}
        \includegraphics[width=\linewidth]{figure/face_recognition/randomized_svd_ret_V_sig.pdf}
        \caption{Right matrix $V$ of the SVD decomposition}
        \label{fig:randomized_svd_V}
    \end{subfigure}
    \caption{Precision of the \texttt{randomized\_svd} function in the
        \texttt{face\_recognition} example using the \texttt{gesvd} SVD method
        and RR. Figures~\ref{fig:randomized_svd_U},~\ref{fig:randomized_svd_S},
        and~\ref{fig:randomized_svd_V} show respectively the values $U$,
        $\Sigma$ and $V$. Result precision is lower for smaller singular values
        and this loss of precision translates to singular vectors. It would be
        interesting to investigate the use of numerical precision for
        dimensionality reduction. }
    \label{fig:face_recognition_svd}
\end{figure*}

\begin{figure}
    \centering
    \begin{subfigure}{0.49\linewidth}
        \includegraphics[width=\linewidth]{figure/BRR_singular_values.pdf}
        \caption{BRR (\texttt{gesdd})}
        \label{fig:BRR_svd}
    \end{subfigure}
    \begin{subfigure}{0.49\linewidth}
        \includegraphics[width=\linewidth]{figure/OLS_singular_values.pdf}
        \caption{OLS (\texttt{gesdd})}
        \label{fig:OLS_svd}
    \end{subfigure}
    \begin{subfigure}{0.49\linewidth}
        \includegraphics[width=\linewidth]{figure/BRR_singular_values_fixed.pdf}
        \caption{BRR (\texttt{gesvd})}
        \label{fig:BRR_svd_fixed}
    \end{subfigure}
    \begin{subfigure}{0.49\linewidth}
        \includegraphics[width=\linewidth]{figure/OLS_singular_values_fixed.pdf}
        \caption{OLS (\texttt{gesvd})}
        \label{fig:OLS_svd_fixed}
    \end{subfigure}
    \caption{Result precision of the singular values of the SVD from the
        Bayesian Ridge model fitting (left) and from the Ordinary Least Square
        model fitting (right) in the Bayesian Ridge Regression example. Red
        stars represent \texttt{NaN} values. Switching from Divide \& Conquer
        (\texttt{gesdd}) to QR (\texttt{gesvd}) factorization improves numerical
        precision. PyTracer helps visualizing improvements.}
\end{figure}

\subsection{Overhead remains tractable and evolves linearly with the trace size}

\pytracer's tracing overhead consists of (1) module instrumentation time
(initialization overhead) and (2) data write time during the execution (runtime
overhead). The initialization overhead is proportional to the number of
functions and classes traced. For example, we measured an average initialization
time of 8 seconds during our experiments with  NumPy, SciPy, and Scikit-learn.
This overhead can be reduced by excluding intensively used functions such as
\texttt{numpy.array} from the instrumentation.
Figure~\ref{fig:performance_tracing} summarizes \pytracer's tracing overhead for
SciPy and scikit-learn examples. \textit{Instrumentation and initialization}
represents the slowdown of \pytracer without fuzzy, measured by applying
PyTracer to Python programs without Monte-Carlo Arithmetic. The slowdown is
measured as the ratio between the actual execution time (without \pytracer) and
the execution time using \pytracer. The \textit{Total with RR MCA} shows the
slowdown within the Fuzzy environment. Without Fuzzy, \pytracer introduces an
average slowdown of $17\times$ for the examples running less than 8 seconds
because \pytracer spends most of its time in the initialization step. Since this
step is a constant factor of the number of functions instrumented, we also
measured the instrumentation cost itself (\textit{Instrumentation only})
calculated by subtracting the initialization time to \textit{Instrumentation}.
We then obtain an average slowdown of $3.5\times$. Finally, using MCA operations
(through Fuzzy) instead of the native ones increased the average slowdown to
$37\times$. As shown in Figure~\ref{fig:performance_parsing}, the log
aggregation time is linearly correlated with the log trace size (F-score=68,
p-value $\leq 10^{-8}$, F-test for linear regression).

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figure/performance.pdf}
    \caption{ \pytracer tracing overheads. \textit{Slowdown} represents the
        slowdown of \pytracer without Fuzzy measured as the ratio between the
        actual execution time (without \pytracer) and the execution time using
        \pytracer. \textit{Slowdown (Instrumentation)} show the actual
        instrumentation cost computed by subtracting the initialization step.
        The \textit{Slowdown (RR)} is the slowdown of \pytracer using the Fuzzy
        environment with the RR mode. The last four examples measure the
        initialization steps depending on the modules instrumented.
    }
    \label{fig:performance_tracing}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figure/performance_parsing.pdf}

    \caption{\pytracer aggregation performance.
    }
    \label{fig:performance_parsing}
\end{figure}

In conclusion, while \pytracer's overhead is substantial, it remains tractable
for real-life examples.

\section{Discussion}

Through our analysis of SciPy and scikit-learn, we demonstrated \pytracer's
scalability, automaticity, and applicability to a wide range of Python programs.
The relatively low instrumentation overhead of 3.5$\times$ on average enables
the analysis of real-scale applications. \pytracer's instrumentation is fully
automatic and does not require manual intervention from the user. Furthermore,
\pytracer works with any Python libraries, with any data format supported by
pickle and any numerical noise model.

\pytracer is the first tool to enable a systematic numerical analysis of Python
codes. Existing tools focus on traditional languages of the HPC community, such
as C, C++, and FORTRAN. However, Python's high-level abstraction allows for
quick application implementations while maintaining high-performance thanks to
specialized libraries such as NumPy. Therefore, \pytracer meets the numerical
analysis needs of an ever-increasing number of applications.

\pytracer allows users to pinpoint the root causes of numerical instabilities by
working at the level of Python functions. Numerical stability evaluations are
complex since calculations are strongly interdependent. Function-level
instrumentation limits the size of the traces since only inputs and outputs of a
function are dumped. Moreover, it decreases instrumentation overhead and helps
the user identify instabilities quickly. Working at the function level is thus a
suitable compromise between analyses conducted at the floating-point operation
level and user debug prints. Indeed, operation-level analyses are costly and
hard to interpret as they overwhelm users with information. In contrast, looking
at intermediate results through print statements is lightweight but not reliable
since developers may not systematically instrument unstable functions.

\pytracer can be used with different types of numerical perturbations, including
random seeds, data noise, or stochastic arithmetic. Using Verificarlo to
implement stochastic arithmetic through Fuzzy enables us to study the
impact numerical perturbations introduced in non-Python dependencies
while preserving the high-level view of Python variables and functions. For
instance, the SciPy environment wraps many computation kernels written in C,
such as BLAS or LAPACK, which have a considerable impact on the stability of
applications using them.

Enabling Monte-Carlo arithmetic in such non-Python dependencies requires their
recompilation with Verificarlo, which usually requires human intervention. The
Fuzzy environment provides a pre-instrumented version of NumPy, BLAS, LAPACK,
GNU libmath, SciPy, and Scikit-learn, which, given the prevalence of these
libraries in scientific Python codes, should cover a large ground of non-Python
dependencies encountered in scientific Python programs.

Integrating \pytracer at the early stages of the development cycle would improve
the overall numerical quality of the developed code. \pytracer's automaticity
facilitates its usage and enables its integration in unit testing. For example,
\pytracer could be used in regression tests to ensure that updates do not
degrade the numerical quality.

\pytracer might also be useful for selecting hyperparameters, as is the case in
machine learning. Indeed, it is common for data scientists to explore a range of
hyperparameters and select the ones giving the best scores. With \pytracer, the
user has a practical way to quantify numerical quality and understand why
hyperparameters produce inconsistent results instead of tweaking hyperparameters
to make solvers converge. For instance, the Face Recognition example
(Figure~\ref{fig:face_recognition_svd}) shows how numerical precision could
inform the selection of the number of components in dimensionality reduction.

\pytracer's visualization provides a global view of the code and interactions
between computations. In addition, the Plotly dashboard helps navigate through
the code easily, without the need for ad-hoc scripts to visualize data. The
dashboard can be extended to support other data formats to visualize. At this
stage, \pytracer currently supports n-dimensional NumPy arrays as well as native
Python data.

Finally, the wide range of result precision observed in our experiments suggests
that such analyses would benefit from being conducted systematically. We showed
that well-known scientific libraries are prone to numerical errors even though
they have been well tested for decades. The prevalence of these libraries in
scientific codes should thus impel users to integrate numerical analysis in
their development workflow in addition to testing their implementation.

\section{Related work}

Several tools have been developed to assess the numerical quality and can be
divided into static and dynamic approaches. Static approaches typically analyze
source code, while dynamic approaches trace floating-point computations
throughout executions. A detailed review of these approaches is presented
in~\cite{cherubin2020tools}.
Numerous tools exist to trace floating-point computations for C, C++, or FORTRAN
programs due to the prevalence of these languages in High-Performance Computing
(HPC). The main tracing techniques are \textit{source-to-source},
\textit{compiler-based transformations}, and \textit{dynamic binary
    instrumentation}.

The source-to-source technique requires a rewriting of the application to modify
floating-point types and provides fine-grained control of the analyzed code.
CADNA~\cite{jezequel2008cadna} is a library for C, C++, and Fortran implementing
CESTAC~\cite{vignes1993stochastic} stochastic arithmetic.
Shaman~\cite{demeure_phd} is a C++ library that uses a first-order error model
to propagate numerical errors. MCAlib~\cite{frechtling2015mcalib} is a C library
that implements the Monte Carlo Arithmetic (MCA)~\cite{parker1997monte} using
the MPFR~\cite{fousse2007mpfr} library. ADAPT \mbox{~\cite{menon2018adapt}} uses
Automatic Differentiation to compute a sensitivity profile for each variable in
a program. Finally, the work in~\cite{tang2016software} proposes a
source-to-source framework for executing targeted code in infinite and fixed
precision with and without stochastic arithmetic. The main drawback of the
source-to-source approach is its lack of scalability since rewriting large codes
can be a tedious task.

The compiler-based approach instruments floating-point expressions at compile
time and so allows the user to automatically instrument large codebases.
Verificarlo~\cite{verificarlo} is a compiler that supports different types of
arithmetic instrumentations, including MCA. The work in~\cite{bao2013fly}
modified the GNU Compiler Collection (GCC) to track local floating-point errors
across executions. pLiner~\cite{guo2020pliner} is a root-cause analyzer based on
the Clang compiler that detects floating-point operations responsible for result
variability using a source-to-source transformation at the Abstract Syntax Tree
(AST) level to rewrite parts of code with higher precision.
FPSanitizer~\cite{chowdhary2020debugging,chowdhary2021parallel} and
NSan~\mbox{\cite{courbet2021nsan}} are compiler-based approaches that use shadow
execution to detect numerical issues using higher precision.
FLiT~\cite{sawaya2017flit} is a framework to detect variability induced by
compilers and their optimizations. The work in~\cite{wang2012development}
proposes a numerical debugger based on GDB~\cite{stallman1988debugging} for
Discrete Stochastic Arithmetic (DSA) on FPGA as an Eclipse plugging. Similarly,
Cadtrace~\cite{jezequel2008cadna} and Shaman propose a GDB-based tool to use the
CADNA and Shaman libraries with GDB, respectively. The main limitation of the
compiler-based approach is that one must have access to the source code.

Dynamic Binary Instrumentation (DBI) operates directly on executables, without
the need for recompilation or manual changes. Therefore, it applies to any
programming language. CraftHPC~\cite{lam2013dynamic} uses DBI to detect
catastrophic cancellations at runtime. Verrou~\cite{fevotte2016verrou} is a
Valgrind~\cite{nethercote2007valgrind} based tool that dynamically replaces
floating-point operations with their stochastic arithmetic counterparts.
FPDebug~\cite{benz2012dynamic} uses DBI to detect numerical inaccuracies by
using shadow computations with higher precision.
Herbgrind~\cite{sanchez2017finding} is a Valgrind-based tool to detect numerical
instabilities. It uses a shadow memory to detect  precision losses by comparison
with results obtained in higher precision. It is combined with symbolic
computation to backtrack the root of the error. We can note that all methods
based on a high-precision oracle suppose that computing with a large number of
bits is sufficient to obtain an accurate result, which is not always true (see,
for example, the famous Muller's sequence~\cite{bajard1996introduction}).
Although versatile, the DBI loses high-level information useful to understand
the source-code logic.

Being agnostic to the programming language is a serious advantage of the DBI
method compared to source-to-source and compiler-bases methods. However, working
at the binary level makes it difficult to access high-level information needed
to debug and understand the source-code logic. Conversely, the source-to-source
approach provides fine-grained control of the analyzed code, but it lacks
scalability, as rewriting large codes can be a tedious task. Finally, the
compiler-based approach takes advantage of the best of both by being automatic
and having access to high-level information. However, like source-to-source, the
compiler-based method is not suitable for analyzing closed-source libraries.

To the best of our knowledge, \pytracer is the first tool dedicated to the
numerical analysis of Python code. Existing tools for tracing Python code focus
on performance profiling for time (cProfile) or memory consumption (memprofile).
Anteater~\cite{faust2019anteater} is a Python tracer tool to debug Python
applications. It performs source transformations at the AST level but only deals
with native numeric Python types. Moreover, the user needs to manually tag each
variable to trace. Finally, according to the authors, Anteater does not scale to
large traces. cProfile, memprofile, and Anteater use Python decorators as the
primary instrumentation technique, a Python mechanism to instrument a function
by adding a line over a function declaration. While this method is appropriate
when targeting specific functions, it is not feasible for large codebases where
potentially unstable code sections are unknown.

\section{Conclusion}

Evaluating the numerical stability of scientific Python programs is a tedious
task that requires substantial expertise. Automatic tools to help users
understand the behavior of their code are therefore crucial. \pytracer is the
first tool to offer a complete framework for analyzing and visualizing the
numerical stability of Python codes. It works by automatically instrumenting
Python programs to produce traces of numerically-perturbed executions and
visualize these traces through an interactive Plotly Dash server. \pytracer does
not require manual code modification. We tested it with Python programs
involving major scientific libraries such as NumPy, SciPy, and Scikit-learn.
\pytracer enabled the systematic evaluation of 40 examples from the SciPy and
scikit-learn libraries. Our results show that \pytracer is scalable, with an
average instrumentation slowdown of $3.5\times$. Moreover, \pytracer's
visualization interface remains responsive even with traces of up to 1TB.

In the future, we plan to extend the post-processing analysis to aggregate
traces obtained with different control flow paths, by leveraging techniques used
in code coverage analysis. This would also enable larger support for
multi-threaded applications. Moreover, we would like to extend \pytracer to
instrument elementary arithmetic operations (+,-,*,/), to allow for more
detailed analyses.

\ifCLASSOPTIONcaptionsoff
    \newpage
\fi

\bibliographystyle{IEEEtran}
\bibliography{paper}

\vskip -2\baselineskip plus -1fil
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{photos/yohan.jpg}}]{Yohan
        Chatelain} received his Ph.D. degree from the University of Paris-Saclay
    (UVSQ), Versailles, France in 2019. He is a postdoctoral researcher in
    the Big Data Infrastructures for Neuroinformatics lab at the University
    of Concordia, Montreal, Canada. His research topics include computer
    arithmetic, high-performance computing, and reproducibility. Yohan aims
    at democratizing the use of the analysis of stability for scientific
    computing codes through automatic tools to improve numerical quality.
\end{IEEEbiography}

\vskip -2\baselineskip plus -1fil
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{photos/NigelPhoto.png}}]{Nigel
        Yong Sao Young} was born in Mauritius Island. He is a recent Computer
    Science graduate at Concordia University. He is currently working on his
    software startup, based in the city of Montreal.
\end{IEEEbiography}

%% % if you will not have a photo at all:
\vskip -2\baselineskip plus -1fil
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{photos/greg_kiar.png}}]{Gregory
        Kiar} is a research scientist at the Child Mind Institute. Throughout
    his degrees in Biomedical Engineering, Greg has developed techniques to
    study biosignal data, a turn-key tool that allows researchers to
    generate maps of brain connectivity from diffusion MRI data, and
    techniques to assess and improve the stability of research in
    neuroscience. Greg’s research bridges the fields of numerical analysis
    and brain imaging to evaluate and improve the trustworthiness of
    techniques used to study the brain.
\end{IEEEbiography}

\vskip -2\baselineskip plus -1fil
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{photos/glatard_tristan.jpg}}]{Tristan
        Glatard} is Associate Professor in the Department of Computer Science
    and Software Engineering at Concordia University in Montreal, and Canada
    Research Chair (Tier II) on Big Data Infrastructures for
    Neuroinformatics. Before that, he was research scientist at the French
    National Centre for Scientific Research and Visiting Scholar at McGill
    University.
\end{IEEEbiography}

\end{document}